### 2019-07-11 manuscript/toeplitz_manuscript_rev.tex
manuscript folder

\documentclass[paper]{geophysics}
%\documentclass[paper,revised]{geophysics}
\usepackage{listings}
% An example of defining macros
\newcommand{\rs}[1]{\mathstrut\mbox{\scriptsize\rm #1}}
\newcommand{\rr}[1]{\mbox{\rm #1}}

\begin{document}

\title{Fast equivalent layer technique for gravity data processing with Block-Toeplitz Toeplitz-Block matrix systems}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\ms{GEO-Example} % paper number

\address{
\footnotemark[1]Observat\'orio Nacional, \\
77 General Jos\'e Cristino Av., \\
Rio de Janeiro, RJ, 20921400 \\
}
\author{Diego Takahashi\footnotemark[1], Vanderlei C. Oliveira Jr.\footnotemark[1] and Val\'eria
	C. F. Barbosa\footnotemark[1]}

%\footer{Example}
\lefthead{Takahashi, Oliveria Jr. \& Barbosa}
\righthead{\emph{Geophysics}}

\maketitle

\begin{abstract}
We present a new approach of the fast equivalent layer technique for gravity data processing, modifying \cite{siqueira2017fast}'s work, with the potential to use very large datasets at low computational cost. Taking advantage of the properties related to the symmetric Block-Toeplitz Toeplitz-block (BTTB) and Block-Circulant Circulant-Block (BCCB) matrices, that raises when regular grids of observation points and equivalent sources are used with the equivalent layer, we developed an algorithm which greatly reduces the number of flops and memory RAM necessary to complete the process of parameter estimative using this technique. The algorithm is based on the struture of symmetric BTTB matrices, where all its elements are comprised by the first row and can be embedded into a symmetric BCCB matrix, that also only needs its first row to be completed reconstructed. From the first column, the eigenvalues of BCCB matrices can be calculated using the Fast Fourier Transform, which can be used to readily compute matrix-vector products. Using examples, we demonstrate that even small and medium sized grids benefits from this approach and larger the dataset, faster and more efficient this method becomes compared to the usual. Synthetic tests using the equivalent layer with the method presented in this work evaluate this approach, and demonstrate satisfatory results for different gravity processing, as upward and downward continuation. Tests with real data from Caraj\'as, Brazil, shows its applicabillity and potential for field processing.
\end{abstract}

\section{Introduction}

The equivalent layer is a well-known technique for processing potential-field data in applied geophysics since the 60's. It comes from potential theory as a mathematical solution of the Laplace's equation, in the region above the sources, by using the Dirichlet boundary condition \cite[]{kellogg1929}.
This theory states that any potential field produced by an arbitrary 3D physical property distribution can be exactly reproduced by a fictitious layer located at any depth and having a continuos 2D physical property  distribution. In pratical situations, the layer is approximated by a finite set of sources (e.g., point masses or dipoles) and their physical properties are estimated by solving a linear system to fit the observed potential field. These fictitious sources are called equivalent sources.

Many previous works have used the equivalent layer as a processing technique in potential methods. \cite{dampney1969equivalent} used  for gridding and vertical continuation. \cite{cordell1992, mendonca-silva1994} used for data interpolation and gridding. \cite{emilia1973,hansen-miyazaki1984,li-oldenburg2010} for upward continuation. \cite{silva1986,leao-silva1989,guspi-novara2009,oliveirajr-etal2013} to reduction to the pole of magnetic data. \cite{boggs-dransfield2004} for combination of multiple data sets. \cite{barnes-lumley2011} for gradiente data processing.

The classic equivalent layer formulation consists on the multiplication between the property distribution and the kernel of the function. For gravity data, this kernel is a harmonic funtion related to the inverse of the distance between the observation point and the equivalent source. When these observation points and equivalent sources are regularly spaced, a Toeplitz system arises. Toeplitz systems are well-known in many fields of science as mathematics - numerical partial and ordinary diferential equations \cite[]{lin2003strang} -, image processing \cite[]{chan1999cosine} and neural networks \cite[]{wray1994calculation}. \cite{chan2007introduction} and \cite{jin2003developments} give many examples of applications for Toeplitz systems.

In potential methods the Toeplitz system properties was used for downward continuation \cite[]{zhang2016bttb} and for 3-D gravity field inversion using a 2-D multilayer model \cite[]{zhang2015bttb}. In the particular case of gravity data, the kernel generates a linear system with a matrix known as symmetric Block-Toeplitz Toeplitz-Block (BTTB).

Because of the importance and vast occurency of Toeplitz systems, many authors studied methods for solving them. Direct methods were conceived by \cite{levinson1946wiener} and by \cite{trench1964algorithm}. Currently the conjugate gradient is used in most cases. In \cite{grenander1984szeg}, Szeg\"{o} noticed that a circulant matrix can be diagonalized by taking the fast Fourier trasform of its first column, making it possible to calculate the matrix-vector product and solve the system with low computational cost \cite[]{strang1986introduction,olkin1986linear}. \cite{chan2007introduction} shows some preconditioners to embedd the Toeplitz and BTTB matrices into circulant matrices and Block-Circulant Circulant-Block (BCCB), respectively, solving the system applying the conjugate gradient method.

Although the use of the equivalent layer technique increased over the last decades, one of the its biggest problem stills its high computational cost for processing large data sets. \cite{siqueira2017fast} developed a computationally efficient scheme for processing gravity data. This scheme does not solve a linear system, instead uses an iterative process that corrects the property distribution of the layer by proportionally adding mass to the gravity residual. Although of its efficiency, at each iteration the forward problem must be solved to evaluate the convergence of the estimative. This processes accounts for most of the computational cost of this method.

We propose the use of BTTB and BCCB matrices properties to solve the forward problem of \cite{siqueira2017fast}'s method in a more efficient way, resulting in faster parameter estimation and the possibility to use very large datasets. We show how the system memory RAM usage can be drastically decreased by calculating only the first row of the BTTB matrix and embedding into a BCCB matrix. Using the Szeg\"{o} theorema  combined with \cite{strang1986introduction} the matrix-vector product can be accomplished with very low cost, reducing in some orders of magnitude the number of operations required to complete the process. We present synthetic tests to validate our proposal and real field data from Caraj\'as, Brazil to demonstrate its applicability.

\section*{Methodology}
\subsection{Equivalent layer theory for gravity data}
In applied geophysics, the observed gravity disturbance $\delta g(x, y, z)$ \citep{heiskanen1967physical} is commomly approximated as the vertical component $z$ of the gravitational attraction produced by gravity sources, as follows:

\begin{equation}
\delta g(x, y, z) = c_{g} \, G \, \int\int\limits_{v}\int \rho(x^{\prime}, y^{\prime}, z^{\prime}) \frac{(z^{\prime} - z) \quad dx^{\prime}dy^{\prime}dz^{\prime}}{[(x - x^{\prime})^{2} + (y - y^{\prime})^{2} + (z - z^{\prime})^{2}]^{\frac{3}{2}}}\: ,
\label{eq:gravity-disturbance}
\end{equation}
where $c_{g} = 10^5$ is a constant transforming from $m/s^2$ to $mGal$, $G$ is the Newton's gravitational constant $(m^3/kg \, s^2)$ and $\rho(x^{\prime}, y^{\prime}, z^{\prime})$ is the density at the point $(x^{\prime}, y^{\prime}, z^{\prime})$ inside the volume $V$ of the source. This integral is defined in a Cartesian coordinate system with axis $x$ pointing to north, $y$ pointing to east and $z$ pointing downward.

According to the classic upward continuation integral \citep{henderson1960comprehensive, henderson1970validity}, it is possible to compute the gravity disturbance $\delta g(x_{i}, y_{i}, z_{i})$, at a point $(x_i, y_i, z_i)$, from the gravity disturbance $\delta g(x, y, z_{0})$ at the constant plane $z_{0}$:

\begin{equation}
\delta g(x_i, y_i, z_i) = \frac{1}{2\pi}  \int\limits_{-\infty}^{+\infty} \int\limits_{-\infty}^{+\infty} \,
\frac{\delta g(x, y, z_0) (z_{0} - z_{i})}{\left[(x_{i} - x)^{2} +
	(y_{i} - y)^{2} +
	(z_{i} - z_{0})^{2} \right]^{\frac{3}{2}}} \, 
dx \, dy,
\label{eq:upward-continuation-integral}
\end{equation}
where, $z_0 > z_i$.
Equation \ref{eq:upward-continuation-integral} shows that the gravity disturbance $\delta g(x_{i}, y_{i}, z_{i})$ is a convolution between $\delta g(x, y, z_{0})$ and another harmonic function on the horizontal plane $z=z_0$.
Multiplying and dividing equation \ref{eq:upward-continuation-integral} by $G$ and discretizing the integral, we obtain:

\begin{equation}
\delta g(x_{i}, y_{i}, z_{i}) = \sum_{j=1}^{N} p_j a_{ij} \; ,
\label{eq:integral-sum}
\end{equation}
where $a_{ij}$ is given by:

\begin{equation}
a_{ij}= c_{g} \, G \, \frac{(z_{j} - z_{i})}{\left[(x_{i} - x_{j})^{2} +	(y_{i} - y_{j})^{2} +	(z_{i} - z_{j})^{2} \right]^{\frac{3}{2}}}
\label{eq:aij}
\end{equation}

\noindent and $p_j$ is the coefficient representing the physical property of the $j$-th equivalent source. In this case, the harmonic function $a_{ij}$ represents the vertical component of the gravitational attraction exerted, at the point $(x_{i}, y_{i}, z_{i})$, by a point mass located at the point $(x_{j}, y_{j}, z_{j})$, with mass $p_{j}$.
Equation \ref{eq:integral-sum} can be expressed in matrix form as:

\begin{equation}
\mathbf{d}(\mathbf{p}) = \mathbf{A} \mathbf{p} \: ,
\label{eq:predicted-data-vector}
\end{equation}
where $\mathbf{d}(\mathbf{p})$ is an $N \times 1$ vector, whose i-th element is the predicted gravity disturbance $\delta g(x_{i}, y_{i}, z_{i})$, $\mathbf{p}$ is an $N \times 1$ parameter vector,  whose j-th element is the coefficient $p_{j}$ representing the physical property of the $j$-th equivalent source and $\mathbf{A}$ is  an $N \times N$ sensibility matrix, where each element $a_{ij}$ is given by equation 4 and defines the field produced by  the j-th equivalent source at the point $(x_{i}, y_{i}, z_{i})$.
The solution of the parameters \textbf{p} can be solved by a linear inversion minimizing the function:

\begin{equation}
\Psi(\mathbf{p}) = \theta_g(\mathbf{p}) + \mu \, \theta_m(\mathbf{p}) \: ,
\label{eq:goal-function}
\end{equation}
where $\theta_g(\mathbf{p})$ is the misfit function given by:

\begin{equation}
\theta_g(\mathbf{p}) = ||\mathbf{d}^{o}-\mathbf{d(p)}||_2^2 \: ,
\label{eq:goal-function_d}
\end{equation}
which is the euclidian norm of the residual between the observed data $\mathbf{d}^{o}$ and the predicted data $\mathbf{d(p)}$.

The function $\theta_m(\mathbf{p})$ is the regularization, for example, the zeroth-order Tikhonov:

\begin{equation}
\theta_m(\mathbf{p}) = ||\mathbf{p}||_2^2 \: ,
\label{eq:tikhonov-function}
\end{equation}
which is the euclidian norm of the parameters $\mathbf{p}$. The variable $\mu$ is a real positive number regularizing the parameter.

Derivating equation \ref{eq:goal-function} in relation to $\mathbf{p}$ and making equal to zero, it is possible to estimate the parameters:

\begin{equation}
\mathbf{p}^{\ast} = \left( \mathbf{A}^{\top}\mathbf{A} + 
\mu \, \mathbf{I} \right)^{-1}
\mathbf{A}^{\top} \mathbf{d}^{o} \,.
\label{eq:p-ast-parameter-space}
\end{equation}

\subsection{Fast equivalent layer technique}
In \cite{siqueira2017fast} the authors developed an iterative least-squares method to estimate the mass distribution of the equivalent layer based on the excess of mass and the positive correlation between the observed gravity data and the equivalent sources. This scheme is proven to have a better computational efficiency than the classical equivalent layer approach with data sets of at least 500 observation points, even using a large number of iterations.
This work also proves that the excess of mass of a body is proportional to the surface integration of the gravity data. Considering each equivalent source directly beneath the observation points, a initial approximation of mass distibution is made:

\begin{equation}
\mathbf{p}^0 = \tilde{\mathbf{A}}^{-1} \mathbf{d}^{o} \: ,
\label{eq:initial_m}
\end{equation}
where $\mathbf{d}^o$ is the gravity data and $\tilde{\mathbf{A}}^{-1} = \Delta s/(2 \pi \, G \, c_g)$ with $\Delta s$ being an element of area located at the vertical coordinate $z_i$ and centered at the horizontal coordinates $(x_i,y_i), \, i = 1,..., N$. At each $k$-th iteration a mass correction vector $\mathbf{\Delta m^k}$ for every source is calculated by minimizing the function:

\begin{equation}
\phi(\mathbf{\Delta p^k}) = ||\mathbf{d}^{o} - \mathbf{A}\hat{\mathbf{p}}^k - \tilde{\mathbf{A}}\mathbf{\Delta p^k}||_2^2
\label{eq:goal-function_fast}
\end{equation}

\noindent and the mass distribution of the equivalent sources is updated as:

\begin{equation}
\mathbf{\Delta p^{k+1}} = \mathbf{p^{k}} + \mathbf{\Delta p^{k}} \: .
\label{eq:update_m}
\end{equation}

\lstset{language=Python}
\lstset{frame=lines}
\lstset{caption={A \textit{Python} algorithm for \cite{siqueira2017fast}'s fast equivalent layer.}}
\lstset{label={lst:code_direct}}
\lstset{basicstyle=\footnotesize}
\begin{lstlisting}
diagonal_A = G*(10**5)*2*pi/(dx*dy)
for i in range (itmax):
	res = (data - A.dot(p))
	delta_p = res/diagonal_A
	p += delta_p
\end{lstlisting}

Each iteration of this algorithm, the matrix-vector product $\tensor{A} \mathbf{p}$ must be calculated to get a new residual $\mathbf{d^0} - \tensor{A} \mathbf{p}$. While it is true that for a small number of observation points this represents no computational effort, for very large data sets it is costful and can be overwhelming in terms of memory RAM to maintain such operation.

\subsection{Structure of sensibility matrix A}
If a gravity disturbance  data $\delta g(x_i, y_i, z_i)$ (equation \ref{eq:integral-sum}) is placed on a regular grid at the constant vertical coordinate $z_{i} = z_{1}$, $i = 1, \dots, N$, and there is an equivalent source located directly below each point of this grid at a constant depth $z_{0}$, the elements $a_{ij}$ (equation \ref{eq:aij}) can be rewritten as follows:

\begin{equation}
a_{ij} = c_{g} G 
\frac{\Delta z}{\left[(r_{ij})^{2} + (\Delta z)^{2}\right]^{\frac{3}{2}}} \: ,
\label{eq:a_ij_r_ij}
\end{equation}
where $\Delta z = z_{0} - z_{1}$ and $r_{ij} = \sqrt{(x_{i} - x_{j})^{2} + (y_{i} - y_{j})^{2}}$ represents the  relative horizontal distance between the $i$-th observation point and the $j$-th equivalent source. It is worth noting that: (i) $r_{ij} = r_{ji}$ for any pair $ij$ and (ii) $r_{ij}$ computed for different pairs $ij$ may assume the same value. As a consequence of these two properties, the matrix $\mathbf{A}$ (equation \ref{eq:predicted-data-vector}) has the structure of a symmetric Toeplitz by blocks, where each block is also a symmetric Toeplitz matrix (BTTB) \cite[]{chan2007introduction, golub2013}. 
For example, consider a regular grid of $N_{x} \times N_{y}$ points, where $N_{x} = 4$ and $N_{y} = 3$, comprising a total number of $N = N_{x} \times N_{y} = 12$ points. In this case, matrix $\mathbf{A}$ is $N_{x} \times N_{x}$ grid of matrices given by:

\begin{equation}
\tensor{A} =
\pmatrix{
	\tensor{A_{0}} & \tensor{A_{1}} & \tensor{A_{2}} & \tensor{A_{3}} \cr
	\tensor{A_{1}} & \tensor{A_{0}} & \tensor{A_{1}} & \tensor{A_{2}} \cr
	\tensor{A_{2}} & \tensor{A_{1}} & \tensor{A_{0}} & \tensor{A_{1}} \cr
	\tensor{A_{3}} & \tensor{A_{2}} & \tensor{A_{1}} & \tensor{A_{0}}
},
\label{eq:toeplitz_block_T}
\end{equation}
where each block is a $N_{y} \times N_{y}$ matrix given by:

\begin{equation}
\tensor{A}_{l} =
\pmatrix{
	a_{0}^{\ell} & a_{1}^{\ell} & a_{2}^{\ell} \cr
	a_{1}^{\ell} & a_{0}^{\ell} & a_{1}^{\ell} \cr
	a_{2}^{\ell} & a_{1}^{\ell} & a_{0}^{\ell}
},
\label{eq:toeplitz_element}
\end{equation}
where the elements $a^{l}_{k}$, $l = 0, \dots, N_{x} - 1$, $k = 0, \dots, N_{y} - 1$ are computed with equation \ref{eq:a_ij_r_ij}. In this case, the complete matrix $\mathbf{A}$ can be obtained by computing only its first row or column.

\subsection{BCCB matrix-vector product}
As previous discussed in the  Fast equivalent layer technique, the matrix-vector product accounts for most of the computational cost (Listing \ref{lst:code_direct}). When large data sets are used, this operation can take some time and even be prohibited by memory RAM shortage. In order to lessen this problem we transform the BTTB matrix $\mathbf{A}$ into a Block-Circulating Circulating-Block (BCCB) matrix $\mathbf{C}$ and use its eigenvalues to carry the product of $\mathbf{A}$ and an arbitrary vector $\mathbf{p}$. This strategy has been successfully applied by Zhang and Wong (2015) and Zhang et al. (2016) for optmizing the computational cost of 3D gravity inversion and downward continuation of potential field, respectively. Here, we use this strategy for improving the computational efficiency of the Fast equivalent layer method proposed by Siqueira et al. (2017).

Following the example of our BTTB matrix $\mathbf{A}$ (equation \ref{eq:toeplitz_block_T}), its transformation into a BCCB matrix $\mathbf{C}$ is given by:

\begin{equation}
\tensor{C} =
\pmatrix{
	\tensor{C_{0}} & \tensor{C_{1}} & \tensor{C_{2}} & \tensor{C_{3}} & \tensor{0} & \tensor{C_{3}} & \tensor{C_{2}} & \tensor{C_{1}} \cr
	\tensor{C_{1}} & \tensor{C_{0}} & \tensor{C_{1}} & \tensor{C_{2}} & \tensor{C_{3}} & \tensor{0} & \tensor{C_{3}} & \tensor{C_{2}} \cr
	\tensor{C_{2}} & \tensor{C_{1}} & \tensor{C_{0}} & \tensor{C_{1}} & \tensor{C_{2}} & \tensor{C_{3}} & \tensor{0} & \tensor{C_{3}} \cr
	\tensor{C_{3}} & \tensor{C_{2}} & \tensor{C_{1}} & \tensor{C_{0}} & \tensor{C_{1}} & \tensor{C_{2}} & \tensor{C_{3}} & \tensor{0} \cr
	\tensor{0} & \tensor{C_{3}} & \tensor{C_{2}} & \tensor{C_{1}} & \tensor{C_{0}} & \tensor{C_{1}} & \tensor{C_{2}} & \tensor{C_{3}} \cr
	\tensor{C_{3}} & \tensor{0} & \tensor{C_{3}} & \tensor{C_{2}} & \tensor{C_{1}} & \tensor{C_{0}} & \tensor{C_{1}} & \tensor{C_{2}} \cr
	\tensor{C_{2}} & \tensor{C_{3}} & \tensor{0} & \tensor{C_{3}} & \tensor{C_{2}} & \tensor{C_{1}} & \tensor{C_{0}} & \tensor{C_{1}} \cr
	\tensor{C_{1}} & \tensor{C_{2}} & \tensor{C_{3}} & \tensor{0} & \tensor{C_{3}} & \tensor{C_{2}} & \tensor{C_{1}} & \tensor{C_{0}}
},
\label{eq:circulant_elements}
\end{equation}
where each block $\mathbf{C}_{l}$ is:

\begin{equation}
\tensor{C}_{l} =
\pmatrix{
	\tensor{A}_{l} & \times \cr
	\times & \tensor{A}_{l}
} =
\pmatrix{
	a_{0}^{\ell} & a_{1}^{\ell} & a_{2}^{\ell} & 0^{\ell} & a_{2}^{\ell} & a_{1}^{\ell} \cr
	a_{1}^{\ell} & a_{0}^{\ell} & a_{1}^{\ell} & a_{2}^{\ell} & 0^{\ell} & a_{2}^{\ell} \cr
	a_{2}^{\ell} & a_{1}^{\ell} & a_{0}^{\ell} & a_{1}^{\ell} & a_{2}^{\ell} & 0^{\ell} \cr
	0^{\ell} & a_{2}^{\ell} & a_{1}^{\ell} & a_{0}^{\ell} & a_{1}^{\ell} & a_{2}^{\ell} \cr
	a_{2}^{\ell} & 0^{\ell} & a_{2}^{\ell} & a_{0}^{\ell} & a_{0}^{\ell} & a_{1}^{\ell} \cr
	a_{1}^{\ell} & a_{2}^{\ell} & 0^{\ell} & a_{2}^{\ell} & a_{1}^{\ell} & a_{0}^{\ell}
}.
\label{eq:toeplitz_circulant_element}
\end{equation}

Matrix $\mathbf{C}$ is a $4N_x N_y \times 4N_x N_y$ Block-Circulant formed by Circulant-Blocks matrix. This matrix is formed by a grid of $2N_x \times 2N_x$ blocks, where each block is a $2N_y \times 2N_y$ matrix.

The product $\mathbf{d} = \mathbf{A}\mathbf{p}$ (equation \ref{eq:predicted-data-vector}), then becomes:

\begin{equation}
\mathbf{C} \mathbf{v} = \mathbf{q} \: ,
\label{eq:BCCB_vector_product}
\end{equation}
where $\mathbf{v}$ and $\mathbf{q}$ are $4N_x N_y \times 1$ vectors given by:

\begin{equation}
\mathbf{v} =
\pmatrix{
	\mathbf{v}_{0} \cr
	\mathbf{v}_{1} \cr
	\vdots \cr
	\mathbf{v}_{N_x- 1} \cr
	\mathbf{0}_{(2 N_x N_y)}
}
\end{equation}

\noindent and

\begin{equation}
\mathbf{q} =
\pmatrix{
	\mathbf{q}_{0} \cr
	\mathbf{q}_{1} \cr
	\vdots \cr
	\mathbf{q}_{N_x - 1} \cr
	\mathbf{0}_{(2 N_x N_y)}
} \: ,
\end{equation}
where $\mathbf{0}_{(2 N_x N_y)}$ is a $2 N_x N_y \times 1$ vetor of zeros, and $\mathbf{v}_{\ell}$ and $\mathbf{q}_{\ell}$, $\ell = 0, \dots, N_x-1$ are:

\begin{equation}
\mathbf{v}_{\ell} =
\pmatrix{
	\mathbf{p}_{\ell} \cr
	\mathbf{0}_{(N_y)}
}
\end{equation}

\noindent and

\begin{equation}
\mathbf{q}_{\ell} =
\pmatrix{
	\mathbf{d}_{\ell} \cr
	\mathbf{0}_{(N_y)}
} \: ,
\end{equation}
where $\mathbf{0}_{(N_y)}$ is a $N_y \times 1$ vetor of zeros.

By using the Kronecker product properties, the auxiliary matrix-vetor product can be rewritten as follows:

\begin{equation}
\mathbf{F}_{(2 N_x)}^{\ast} \left[ \mathbf{L} \circ \left( \mathbf{F}_{(2N_x)} \mathbf{V} \, \mathbf{F}_{(2N_y)} \right) \right] \mathbf{F}_{(2N_y)}^{\ast} = \mathbf{Q} \: ,
\label{eq:fft_q}
\end{equation}
where ``$\circ$'' denotes the Hadamard product, $\mathbf{L}$ is a $2N_x \times 2N_y$ row-oriented matrix containing the eigenvalues of $\mathbf{C}_{(BCCB)}$, and $\mathbf{V}$ and $\mathbf{Q}$ are $2N_x \times 2N_y$ row-oriented matrices obtained from the vectors $\mathbf{v}$ and $\mathbf{q}$.

One of the properties of BCCB matrices is that its eigenvalues can be calculated by a 2D Discrete Fourier Transform \cite[]{chan2007introduction}. This lead to a fast calculation of $\mathbf{L}$ using a 2D Fast Fourier Transform, making the matrix-vector product a low computational cost process.

Note that in general, the first column of blocks forming a BCCB matrix $\mathbf{C}_{\ell}$ (equation \ref{eq:circulant_elements}) is given by:

\begin{equation}
\left[\mathbf{C} \right]_{(0)} = 
\pmatrix{
\mathbf{C}_{0} \cr
\mathbf{C}_{1} \cr
\vdots \cr
\mathbf{C}_{N_x-2} \cr
\mathbf{C}_{N_x-1} \cr
\mathbf{0} \cr
\mathbf{C}_{N_x-1} \cr
\mathbf{C}_{N_x-2} \cr
\vdots \cr
\mathbf{C}_{1}
} \; ,
\end{equation}
where each block $\mathbf{C}_{\ell}$, $\ell = 0, \dots, N_x-1$, is a $2N_y \times 2N_y$ circulant matrix and $\mathbf{0}$ is a $2N_y \times 2N_y$ matrix with all elements equal to zero. Thus, the first column of a circulant matrix $\mathbf{C}_{\ell}$ is given by:

\begin{equation}
\left[\mathbf{C}_{\ell} \right]_{(0)} =
\pmatrix{
	a^{\ell}_{00} \cr
	a^{\ell}_{10} \cr
	\vdots \cr
	a^{\ell}_{(N_y-2)0} \cr
	a^{\ell}_{N_y-1) 0} \cr
	0 \cr
	a^{\ell}_{N_y-1) 0} \cr
	a^{\ell}_{(N_y-2)0} \cr
	\vdots \cr
	a^{\ell}_{10}
} \:.
\end{equation}

To complete the process, after calculating the inverse of $\mathbf{Q}$ it is necessary to rearrange its rows to obtain the vector $\mathbf{q}$ and also rearrange the elements of $\mathbf{q}$ to obtain the wanted vector $\mathbf{d(p)}$.

%The algorithm for computing the matrix-vector product $\mathbf{A}\mathbf{p}$ and obtaining the resulting vector $\mathbf{d(p)}$ can be summarized in the following steps:

%1. Compute the first column of the embedded $4N_x N_y \times 4N_x N_y$ BCCB matrix $\mathbf{C}$ by using the first column of the $N_x N_y \times N_x N_y$ BTTB matrix $\mathbf{A}$;

%2. Compute the eigenvalues of $\mathbf{C}$ rearranging them in rows of the $2N_x \times 2N_y$ matrix $\mathbf{L}$;

%3. Rearrange the vector $\mathbf{v}$ in the rows of the $2N_x \times 2N_y$ matrix $\mathbf{V}$;

%4. Compute the Hadammard product $\mathbf{L} \circ \tilde{\mathbf{P}}$, where $\tilde{\mathbf{P}} = \mathbf{F}_{(2N_x)} \mathbf{P} \, \mathbf{F}_{(2N_y)}$ is computed by using a fast algorithm for 2D DFT;

%5. Compute the resulting $2N_x \times 2N_y$ matrix $\mathbf{Q} = \mathbf{F}_{(2N_x)}^{\ast} \left[ \mathbf{L} \circ \tilde{\mathbf{P}}\right] \mathbf{F}_{(2N_y)}^{\ast}$ by using a fast algorithm for inverse 2D DFT;

%6. Rearrange the rows of $\mathbf{Q}$ to obtain the vector $\mathbf{q}$;

%7. Rearrange the elements of $\mathbf{q}$ to obtain the vector $\mathbf{d(p)}$.

\subsection{Computational performance}
Equation \ref{eq:goal-function_fast} now can be calculated with the approach presented in the previous section. In a normal procedure of the fast equivalent layer, at each iteration a full matrix $\tensor{A}$ (equation \ref{eq:predicted-data-vector}) is multiplied by the estimated mass distribution parameter vector $\hat{\mathbf{p}}^k$. As pointed in \cite{siqueira2017fast} the number of flops (floating-point operations) necessary to estimate the $N$-dimensional parameter vector inside the iteration loop is:

\begin{equation}
f_0 = N^{it} (3N + 2N^2).
\label{eq:float_classic}
\end{equation}

From equation \ref{eq:float_classic} it is clear that the matrix-vector product ($2N^2$) accounts for most of the computational complexity in this method.

It is well known that FFT takes $N \log_2(N)$ flops \cite[]{brigham1988fast}. Computing the eigenvalues of the BCCB matrix ($4N \times 4N$) and applying 2D-FFT on the parameter vector (equation \ref{eq:fft_q}), takes $4N \log(4N)$ each. The point-multiplication takes $4N$. As it is necessary to compute the inverse FFT another two $4N \log(4N)$ must be taken in account. However, the sensibility matrix does not change during the process, thus, the eigenvalues of BCCB must be calculated only once, outside of the iteration. This lead us to:

\begin{equation}
f_1 = N^{it} (7N + 12N\log(4N)).
\label{eq:float_bccb}
\end{equation}

Another major improvement of this methodology is the exoneration of calculating the full sensibility matrix $\tensor{A}$ (equation \ref{eq:predicted-data-vector}). Each element needs $12$ flops (equation \ref{eq:aij}), totalizing $12N^2$ flops for the full matrix. Calculating only the first row of the BTTB matrix, $12N$ flops is required and the computation of the eigenvalues is $4N \log(4N)$ as mentioned above.
The full flops count of \cite{siqueira2017fast}'s method:

\begin{equation}
f_s = 12N^2 + N^{it} (3N + 2N^2),
\label{eq:float_siqueira}
\end{equation}

is decreased in our method to:

\begin{equation}
f_s = 12N + 4N \log(4N) + N^{it} (7N + 12N\log(4N)).
\label{eq:float_new}
\end{equation}

Figure \ref{fig:float} shows the floating points to estimate the parameter vector using the fast equivalent layer with \cite{siqueira2017fast}'s method (equation \ref{eq:float_classic}) and our approach (equation \ref{eq:float_bccb}) versus the number of observation points varyig from $N = 5000$ to $N = 1000000$ with $50$ iterations. The number of operations is drastically decreased.

Table 1 shows the system memory RAM usage needed to store the full matrix, the BTTB first row and the BCCB eigenvalues ($8$ times the BTTB first row). The quantities were computed for different numbers of data (N) with the same corresponding number of equivalent sources (N). This table considers that each element of the matrix is a double-precision number, which requires 8 bytes of storage, except for the BCCB complex eigenvalues, which requires 16 bytes per element. Notice that $1000000$ observation points requires nearly $7.6$ Terabytes of memory ram to store the whole sensibility matrix of the equivalent layer.

Using a PC with a Intel Core i7 4790@3.6GHz processor and 16 Gb of memory RAM, figure \ref{fig:time_comparison} shows the time necessary to run 50 iterations of the \cite{siqueira2017fast}'s method and the one presented in this work. After $10000$ observations points is clear how this method benefits from the new approach in calculating the forward problem. Because of the memory RAM available in this system, we could not test the comparison with more observations, limited to $22500$. In figure \ref{fig:time_bccb} we show the time necessary to run the equivalent layer technique with 50 iterations using only the new approach, where the RAM is not a limitation factor. We could run up to $25$ million observation points. In comparison, one million observation points took $26.8$ seconds to run, where the maximum $22500$ observation points, with \cite{siqueira2017fast}'s method, took $48.3$ seconds.

\section{Synthetic tests}
The synthetic data presented in this section has the objective to validate this new approach when used jointly with the fast equivalent layer presented in \cite{siqueira2017fast}. We constructed a model with two polygonal prisms, with density contrast of $0.35$ (upper-left body) and $0.4 g/cm^3$ (upper-right body), and a sphere with radius of $1000 m$ with density contrast of $-0.5 g/cm^3$. The vertical component of gravity generated by this bodies were calculated and are shown in figure \ref{fig:synthetic_data} together with their horizontal projections. A gaussian noise was added to the data, with mean of zero and a maximum of $0.5\%$ of the maximum value of the original data. As previous said, only in regular grids the BTTB matrix structures appears. We created $10000$ observation points regularly spaced in a grid of $100 \times 100$, with an uniform $100$ m of height for all the observations.

In figure \ref{fig:classic_fast} we show the fitted data with the fast equivalent layer using \cite{siqueira2017fast}'s work. In comparison we have the figure \ref{fig:bccb_fast} showing the fitted data with the new approach presented in this work by calculating the forward problem using equation \ref{eq:BCCB_vector_product}. In both figures A) is the original contaminated synthetic data, B) is the predict data by the equivalent layer and C) is the residual between the synthetic data and fitted data, with mean of $-8.264e^{-7}$ and standard deviation of $0.0144$. As we can see in the figure \ref{fig:delta_comparison}, there is virtually no difference in the fitted data presented in figures \ref{fig:classic_fast}b and \ref{fig:bccb_fast}c, showing that the result associated by calculating a matrix-vector product of a embedded BTTB into a BCCB matrix using equation \ref{eq:BCCB_vector_product} is the same as a normal matrix-vecotr product. In figure \ref{fig:delta_rho} we have the difference between the mass distribution of the equivalent sources estimated by using the two methods.

In figure \ref{fig:upward_comparison} and figure \ref{fig:downward_comparison} we show two forms of processing a gravity data using the equivalent layer, the upward and the downward continuation, respectively. The upward height is $300 m$ and the downward is at $50 m$. Both figures show: A) the upward or downward in the traditional \cite{siqueira2017fast}'s work B) the upward or downward using this new approach and C) the residuals between the two forms of processing. For the upward continuation the mean of the residuals is $-5.938e^{-18}$ and the standard deviation is $8.701e^{-18}$. For the downward continuation the mean of the residuals is $5.914e^{-18}$ and the standard deviation is $9.014e^{-16}$. With \cite{siqueira2017fast}'s method the upward processing took $7.62026$ seconds and $0.00834$ seconds with the new approach. For the downward processing the times necessary were $7.59654$ seconds and $0.00547$ seconds, respectively.

\section{Real data tests}
Tests with real data are conducted with the gravity data of Caraj\'as provided by CPRM (Companhia de Pesquisa de Recursos Minerais). This area covers the southeast part of the state of Par\'a, Brazil. Aeromagnetic and aerogravimetric data were collected in $113$ flight lines with $3$ km apart from each other with N-S orientation. There was two separated teams for data collection, each responsible for a determined area. For gravity data the sample spacing were $7.65 m$ and $15.21 m$ for each team, totalizing more $4353428$ observation points. The height of the flights were fixed at $900 m$. All $4353428$ million gravity data were gridded into a regularly spaced dataset of $250000$ observation points ($500 \times 500$) for processing (figure \ref{fig:carajas_real_data}).

Figure \ref{fig:carajas_gz_predito} shows the gridded gravity data (A), the fitted data with 50 iterations of the fast equivalent layer at $300$ m depth using this new approach (B) and the residual (C). The mean of the residual was $0.000292$ and standard deviation of $0.105$ which demonstrates a good fit for the predicted data, evaluating this technique to be applied in real field data.

An upward continuation processing were made (Figure \ref{fig:up2000_carajas_500x500}) at $5000$ m over the real data. It shows a reasonable processing, atenuating the short-wave lenghts. The processing of the $250000$ observations points took $0.216$ seconds.

\section{Conclusions}
By exploring the properties related to Block-Toeplitz Toeplitz-block (BTTB) and Block-Circulant Circulant -Block (BCCB) matrices, we show a new approach for calculating the matrix-vector product of the fast equivalent layer technique from \cite{siqueira2017fast}'s work when regular grids of observation points and equivalent sources are employed. This alghorithm greatly reduces the number of flops necessary to complete the estimative of the equivalent layer. For example, when processing one million observation points, the number of flops is reduced in $10^4$ times. When processing such amount of data, the full sensibility matrix takes $7.6$ Terabytes of memory RAM storage, which is impractical, while takes only $61.035$ Megabytes with the method presented in this work.

This approach takes advantage of the symmetric BTTB system that arises when processing a harmonic function as the vertical component of gravity, that depends on the inverse of distance. Symmetric BTTB matrices can be stored by its only first row and can be embedded into a symmetric BCCB matrix that also only needs its first row. Using the Fast Fourier Transform it is possible to calculate the eigenvalues of BCCB matrices which can be used to compute a matrix-vector product in a very low computational cost. The time necessary to process medium sized grids of observational data, for example $22500$ points, is cutted in $10^2$ times.

Synthetic and real data tests were conducted with the equivalent layer using the method presented in this work, showing satisfatory results that evaluate it to process gravity data, as upward and downward continuations, even using very large datasets. In the future, applications of the equivalent layer using great amount of data, as in continental or global scale can be researched.
\newpage

\subsection*{Figures}
\renewcommand{\figdir}{Fig} % figure directory

Figure~\ref{fig:float}

\plot{float}{width=\textwidth}
{floating points to estimate the parameter vector using the fast equivalent layer with \cite{siqueira2017fast}'s method (equation \ref{eq:float_classic}) and our approach (equation \ref{eq:float_bccb}) versus the numbers of observation points varyig from $N = 5000$ to $N = 1000000$ with $50$ iterations. The number of operations is drastically decreased.}
\newpage

Figure~\ref{fig:time_comparison}

\plot{time_comparison}{width=10cm}
{time necessary to run 50 iterations of the \cite{siqueira2017fast}'s method and the one presented in this work. With the limitation of $16$ Gb of memory RAM in our system, we could test only up to $22500$ obervation points.}

Figure~\ref{fig:time_bccb}

\plot{time_bccb}{width=10cm}
{time necessary to run the equivalent layer technique with 50 iterations using only this new approach, where the RAM is not a limitation factor. We could run up to $25$ million observation points. In comparison, $1$ million observation points took $26.8$ seconds to run, where the maximun $22500$ observation points in figure \ref{fig:time_comparison}, with Siqueira's method, took $48.3$ seconds.}
\newpage

Figure~\ref{fig:synthetic_data}

\plot{synthetic_data}{width=13.5cm}
{model with two polygonal prisms, with density contrast of $0.35$ (upper-left body) and $0.4 g/cm^3$ (upper-right body), and a sphere with radius of $1000 m$ with density contrast of $-0.5 g/cm^3$. The vertical component of gravity generated by this bodies were calculated and are shown together with their horizontal projections. A gaussian noise was added to the data with mean of zero and maximum value of $0.5\%$ of the maximum of the original data. As previous said only in regular grids the BTTB matrix structures appears. We created $10000$ observation points regularly spaced in a grid of $100 \times 100$, with a uniform $100 m$ of height for all the observations.}
\newpage

Figure~\ref{fig:classic_fast}

\plot{classic_fast}{width=5cm}
{A) original contaminated synthetic data, B) fitted data using the fast equivalent layer with the \cite{siqueira2017fast}'s work. C) residual between the synthetic data and fitted data, with mean of $-8.264e^{-7}$ and standard deviation of $0.0144$.}
\newpage

Figure~\ref{fig:bccb_fast}

\plot{bccb_fast}{width=5cm}
{A) original contaminated synthetic data, B) fitted data with the new approach presented in this work of calculating the forward problem using equation \ref{eq:BCCB_vector_product} for the equivalent layer. C) residual between the synthetic data and fitted data, with mean of $-8.264e^{-7}$ and standard deviation of $0.0144$.}
\newpage

Figure~\ref{fig:delta_comparison}

\plot{delta_comparison}{width=13.5cm}
{virtually zero difference in the fitted data presented in figures \ref{fig:classic_fast}b and \ref{fig:bccb_fast}c, showing that matrix-vector product of a embedded BTTB into a BCCB matrix using equation \ref{eq:BCCB_vector_product} is pratically equal to a normal matrix-vector product.}
\newpage

Figure~\ref{fig:delta_rho}

\plot{delta_rho}{width=13.5cm}
{small difference in the estimated density between the classic matrix-vector product and the matrix-vector product of a embedded BTTB into a BCCB matrix using equation \ref{eq:BCCB_vector_product}, showing that the estimative using this work's method is pratically the same.}
\newpage

Figure~\ref{fig:upward_comparison}

\plot{upward_comparison}{width=5cm}
{A) the upward continuation conducted at the height of $300 m$ using the traditional \cite{siqueira2017fast}'s work B) the upward continuation conducted at the same height of $300 m$ using this new approach and C) residuals between the two forms of processing with mean of $-5.938e^{-18}$ and standard deviation of $8.701e^{-18}$. With \cite{siqueira2017fast}'s method this processing took $7.62026$ seconds and $0.00834$ seconds with the new approach.}
\newpage

Figure~\ref{fig:downward_comparison}

\plot{downward_comparison}{width=5cm}
{A) the downward continuation conducted at the height of $50 m$ using the traditional \cite{siqueira2017fast}'s work B) the upward continuation conducted at the same height of $50 m$ using this new approach and C) residuals between the two forms of processing with mean of $5.914e^{-18}$ and standard deviation of $9.014e^{-16}$. With \cite{siqueira2017fast}'s method this processing took $7.59654$ seconds and $0.00547$ seconds with the new approach.}
\newpage

Figure~\ref{fig:carajas_real_data}

\plot{carajas_real_data}{width=13.5cm}
{real data gravity data of Caraj\'as gridded into a regularly spaced dataset of $250000$ observation points ($500 \times 500$). This area covers the southeast part of the state of Par\'a, Brazil. Aerogravimetric data was collected in $113$ flight lines with $3 km$ apart from each other and N-S orientation, totalizing more than 4 million observation points. The height of the flights were fixed at $900 m$. All $4 million$ gravity data were gridded for processing.}
\newpage

Figure~\ref{fig:carajas_gz_predito}

\plot{carajas_gz_predito}{width=5cm}
{A) gridded gravity data. B) fitted data with 50 iterations of the fast equivalent layer at $400$ m depth using the new approach of this work. C) residual, defined as the difference between A) and B). The mean of the residual was $0.000292$ and standard deviation of $0.105$ which demonstrates a good fit for the predicted data, evaluating this technique to be applied in real field data.}
\newpage

Figure~\ref{fig:up2000_carajas_500x500}

\plot{up2000_carajas_500x500}{width=13.5cm}
{upward continuation processing at $2000 m$ over the real data of Caraj\'as. It shows a reasonable processing, atenuating the short-wave lenghts. The processing of the $250000$ observations points took $0.216$ seconds.}
\newpage

%\subsubsection{Multiplot} 
%Sometimes it is convenient to put two or more figures from different
%files in an array (see Figure~\ref{fig:exph,exgr}). Individual plots
%are Figures~\ref{fig:exph} and~\ref{fig:exgr}.
%
%\multiplot{2}{exph,exgr}{width=0.4\textwidth}
%{This figure is specified in the document by \newline \texttt{
%    $\backslash$multiplot\{2\}\{exph,exgr\}\{width=0.4$\backslash$textwidth\}\{This caption.\}}.
%}
%
%The first argument of the \texttt{multiplot} command specifies the
%number of plots per row.

\subsection{Tables}

\begin{table}[h]
	\begin{center}
		\begin{tabular}{|l|c|c|c|}
			\hline
			\textbf{$N \times N$} & \textbf{Full RAM (Mb)} & \textbf{BTTB RAM (Mb)}  & \textbf{BCCB RAM (Mb)}\\
			\hline 
			$100 \times 100$ & 0.0763 & 0.0000763 & 0.0006104\\
			\hline
			$400 \times 400$ & 1.22 & 0.0031 & 0.0248\\
			\hline
			$2500 \times 2500$ & 48 & 0.0191 & 0.1528\\
			\hline
			$10000 \times 10000$ & 763 & 0.00763 & 0.6104\\
			\hline
			$40000 \times 40000$ & 12207 & 0.305 & 2.4416 \\
			\hline
			$250000 \times 250000$ & 476837 & 1.907 & 15.3 \\
			\hline
			$500000 \times 500000$ & 1907349 & 3.815 & 30.518 \\
			\hline
			$1000000 \times 1000000$ & 7629395 & 7.629 & 61.035 \\
			\hline
		\end{tabular}
		\caption{Comparison between the system memory RAM usage needed to store the full matrix, the BTTB first row and the BCCB eigenvalues (eight times the BTTB). The quantities were computed for different numbers of data (N) with the same corresponding number of equivalent sources (N). This table considers that each element of the matrix is a double-precision number, which requires 8 bytes of storage, except for the BCCB complex eigenvalues, which requires 16 bytes per element.}
	\end{center}
\end{table}

%The discussion is summarized in Table~\ref{tbl:example}.
%
%\tabl{example}{This table is specified in the document by \texttt{
%    $\backslash$tabl\{example\}\{This caption.\}\{\ldots\}}.
%}{
%  \begin{center}
%    \begin{tabular}{|c|c|c|}
%      \hline
%      \multicolumn{3}{|c|}{Table Example} \\
%      \hline
%      migration\rule[-2ex]{0ex}{5ex} & 
%      $\omega \rightarrow k_z$ & 
%      $k_y^2+k-z^2\cos^2 \psi=4\omega^2/v^2$ \\
%      \hline
%      \parbox{1in}{zero-offset\\diffraction}\rule[-4ex]{0ex}{8ex} &
%      $k_z\rightarrow\omega_0$ &
%      $k_y^2+k_z^2=4\omega_0^2/v^2$ \\
%      \hline
%      DMO+NMO\rule[-2ex]{0in}{5ex} & $\omega\rightarrow\omega_0$ & 
%      $\frac{1}{4}
%      v^2k_y^2\sin^2\psi+\omega_0^2\cos^2\psi=\omega^2$ \\
%      \hline
%      radial DMO\rule[-2ex]{0in}{5ex} & $\omega\rightarrow\omega_s$ &
%      $\frac{1}{4}v^2k_y^2\sin^2\psi+\omega_s^2=\omega^2$\\
%      \hline
%      radial NMO\rule[-2ex]{0in}{5ex} & $\omega_s\rightarrow\omega_0$ &
%      $\omega_0\cos\psi=\omega_s$\\
%      \hline
%    \end{tabular}
%  \end{center}
%}

\section{ACKNOWLEDGMENTS}

This study was financed by the brazilian agencies CAPES (in the form of a scholarship), FAPERJ (grant n.$^{\circ}$ E-26 202.729/2018) and CNPq (grant n.$^{\circ}$ 308945/2017-4).

%\append{Appendix example}

%\append[equations]{Another appendix}

%\sideplot{errgrp}{width=0.8\textwidth}
%{This figure is specified in the document by \texttt{
%    $\backslash$sideplot\{errgrp\}\{width=0.8$\backslash$text\-width\}\{This caption.\}}.

%\append{The source of this document}

%\verbatiminput{geophysics_example.ltx}

%\append{The source of the bibliography}

%\verbatiminput{example.bib}

\newpage

\bibliographystyle{seg}  % style file is seg.bst
\bibliography{references}

\end{document}

### 2019-07-16 manuscript/geophysics_paper.tex

%\documentclass[paper]{geophysics}
\documentclass[paper,revised]{geophysics}

% An example of defining macros
\newcommand{\rs}[1]{\mathstrut\mbox{\scriptsize\rm #1}}
\newcommand{\rr}[1]{\mbox{\rm #1}}

\begin{document}

\title{An example \emph{Geophysics} article, \\ with a two-line title}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\ms{GEO-Example} % paper number

\address{
\footnotemark[1]BP UTG, \\
200 Westlake Park Blvd, \\
Houston, TX, 77079 \\
\footnotemark[2]Bureau of Economic Geology, \\
John A. and Katherine G. Jackson School of Geosciences \\
The University of Texas at Austin \\
University Station, Box X \\
Austin, TX 78713-8924}
\author{Joe Dellinger\footnotemark[1] and Sergey Fomel\footnotemark[2]}

\footer{Example}
\lefthead{Dellinger \& Fomel}
\righthead{\emph{Geophysics} example}

\maketitle

\begin{abstract}
  This is an example of using \textsf{geophysics.cls} for writing
  \emph{Geophysics} papers.
\end{abstract}

\section{Introduction}

This is an introduction. \LaTeX\ is a \new{powerful} document
typesetting system \cite{lamport}. An excellent reference is
\cite[]{kopka}. The new \textsf{geophysics.cls} class complies with
the \LaTeX2e\ standard. \old{We had something else written here but
  decided to get rid of it}.

\section*{Theory}

This is another section.

\subsection{Equations}

Section headings should be capitalized. Subsection headings should
only have the first letter of the first word capitalized.

Here are examples of equations involving vectors and tensors:
\begin{equation}
\tensor{R} = 
\pmatrix{R_{\rs{XX}} & R_{\rs{YX}} \cr R_{\rs{XY}} & R_{\rs{YY}}} 
=
\tensor{P}_{M\rightarrow R} \; \tensor{D} \; \tensor{P}_{S\rightarrow M}
\;\;\; \tensor{S} \ \ \  ,
\label{SVD}
\end{equation}
and
\begin{equation}
R_{j,m}(\omega) =
\sum_{n=1}^{N} \, \,
P_{j}^{(n)}(\mathbf{x}_R) \, \,
D^{(n)}(\omega) \, \,
P_{m}^{(n)}(\mathbf{x}_S) \ \ \ .
\label{SVDray}
\end{equation}

Note that the macro for the \verb#\tensor# command has been changed to
force tensors to be bold uppercase, in compliance with current SEG
submission standards. This is so that documents typeset to the old
standards will print out according to the new ones: e.g., tensor
$\tensor{t}$ (note converted to uppercase).

\subsection*{Figures}
\renewcommand{\figdir}{Fig} % figure directory

Figure~\ref{fig:waves} shows what it is about.

\plot{waves}{width=\textwidth}
{This figure is specified in the document by \texttt{
    $\backslash$plot\{waves\}\{width=$\backslash$textwidth\}\{This caption.\}}.
}

\subsubsection{Multiplot}

Sometimes it is convenient to put two or more figures from different
files in an array (see Figure~\ref{fig:exph,exgr}). Individual plots
are Figures~\ref{fig:exph} and~\ref{fig:exgr}.

\multiplot{2}{exph,exgr}{width=0.4\textwidth}
{This figure is specified in the document by \newline \texttt{
    $\backslash$multiplot\{2\}\{exph,exgr\}\{width=0.4$\backslash$textwidth\}\{This caption.\}}.
}

The first argument of the \texttt{multiplot} command specifies the
number of plots per row.

\subsection{Tables}

The discussion is summarized in Table~\ref{tbl:example}.

\tabl{example}{This table is specified in the document by \texttt{
    $\backslash$tabl\{example\}\{This caption.\}\{\ldots\}}.
}{
  \begin{center}
    \begin{tabular}{|c|c|c|}
      \hline
      \multicolumn{3}{|c|}{Table Example} \\
      \hline
      migration\rule[-2ex]{0ex}{5ex} & 
      $\omega \rightarrow k_z$ & 
      $k_y^2+k-z^2\cos^2 \psi=4\omega^2/v^2$ \\
      \hline
      \parbox{1in}{zero-offset\\diffraction}\rule[-4ex]{0ex}{8ex} &
      $k_z\rightarrow\omega_0$ &
      $k_y^2+k_z^2=4\omega_0^2/v^2$ \\
      \hline
      DMO+NMO\rule[-2ex]{0in}{5ex} & $\omega\rightarrow\omega_0$ & 
      $\frac{1}{4}
      v^2k_y^2\sin^2\psi+\omega_0^2\cos^2\psi=\omega^2$ \\
      \hline
      radial DMO\rule[-2ex]{0in}{5ex} & $\omega\rightarrow\omega_s$ &
      $\frac{1}{4}v^2k_y^2\sin^2\psi+\omega_s^2=\omega^2$\\
      \hline
      radial NMO\rule[-2ex]{0in}{5ex} & $\omega_s\rightarrow\omega_0$ &
      $\omega_0\cos\psi=\omega_s$\\
      \hline
    \end{tabular}
  \end{center}
}

\section{ACKNOWLEDGMENTS}

I wish to thank Ivan P\v{s}en\v{c}\'{\i}k and Fr\'ed\'eric Billette
for having names with non-English letters in them.  I wish to thank
\cite{Cerveny} for providing an example of how to make a bib file that
includes an author whose name begins with a non-English character and
\cite{forgues96} for providing both an example of referencing a Ph.D.
thesis and yet more non-English characters.

\append{Appendix example}
\label{example}

According to the new SEG standard, appendices come before references.

\begin{equation}
\frac{\partial U}{\partial z} = 
\left\{
  \sqrt{\frac{1}{v^2} - \left[\frac{\partial t}{\partial g}\right]^2} +
  \sqrt{\frac{1}{v^2} - \left[\frac{\partial t}{\partial s}\right]^2}
\right\}
\frac{\partial U}{\partial t}
\label{eqn:partial}
\end{equation}
It is important to get equation~\ref{eqn:partial} right. See also
Appendix~\ref{equations}.

\append[equations]{Another appendix}

\begin{equation}
\frac{\partial U}{\partial z} = 
\left\{
  \sqrt{\frac{1}{v^2} - \left[\frac{\partial t}{\partial g}\right]^2} +
  \sqrt{\frac{1}{v^2} - \left[\frac{\partial t}{\partial s}\right]^2}
\right\}
\frac{\partial U}{\partial t}
\label{eqn:partial2}
\end{equation}
Too lazy to type a different equation but note the numeration.

The error comparison is provided in Figure~\ref{fig:errgrp}.

\sideplot{errgrp}{width=0.8\textwidth}
{This figure is specified in the document by \texttt{
    $\backslash$sideplot\{errgrp\}\{width=0.8$\backslash$text\-width\}\{This caption.\}}.
}

\append{The source of this document}

\verbatiminput{geophysics_example.ltx}

\append{The source of the bibliography}

\verbatiminput{example.bib}

\newpage

\bibliographystyle{seg}  % style file is seg.bst
\bibliography{example}

\end{document}

### 2019-07-16 manuscript/toeplitz_manuscript.tex

--- 2019-07-11 manuscript/toeplitz_manuscript_rev.tex
@@
 %\documentclass[paper,revised]{geophysics}
-\usepackage{listings}
+
 % An example of defining macros
@@
 \begin{abstract}
-We present a new approach of the fast equivalent layer technique for gravity data processing, modifying \cite{siqueira2017fast}'s work, with the potential to use very large datasets at low computational cost.
-Taking advantage of the properties related to the symmetric Block-Toeplitz Toeplitz-block (BTTB) and Block-Circulant Circulant-Block (BCCB) matrices, that raises when regular grids of observation points and equivalent sources are used with the equivalent layer, we developed an algorithm which greatly reduces the number of flops and memory RAM necessary to complete the process of parameter estimative using this technique.
-The algorithm is based on the struture of symmetric BTTB matrices, where all its elements are comprised by the first row and can be embedded into a symmetric BCCB matrix, that also only needs its first row to be completed reconstructed.
-From the first column, the eigenvalues of BCCB matrices can be calculated using the Fast Fourier Transform, which can be used to readily compute matrix-vector products.
-Using examples, we demonstrate that even small and medium sized grids benefits from this approach and larger the dataset, faster and more efficient this method becomes compared to the usual.
+We present a new approach to the fast equivalent layer technique for gravity data processing, modifying \cite{siqueira2017fast}'s work, with the potential of using very large datasets at low computational cost.
+Taking advantage of the properties related to the symmetric Block-Toeplitz Toeplitz-block (BTTB) and Block-Circulant Circulant-Block (BCCB) matrices, that raises when regular grids of observation points and equivalent sources are used with the equivalent layer, we developed a algorithm which greatly reduces the number of flops and memory RAM necessary to complete the process of estimative using this technique.
+The algorithm is based on the struture of symmetric BTTB matrices, that all its elements are comprised by the first row and can be embedded into a symmetric BCCB matrix that also only needs its first row to be completed reconstructed.
+From this row, the eigenvalues of BCCB matrices are calculated using the Fast Fourier Transform, which can be used to readily compute matrix-vector products for this type of matrix.
+Using examples, we demonstrate that even small and medium sized grids benefits from this approach and larger the dataset, faster and more efficient this method becomes compared to the usual approach.
 Synthetic tests using the equivalent layer with the method presented in this work evaluate this approach, and demonstrate satisfatory results for different gravity processing, as upward and downward continuation.
@@
 \end{equation}
+
 where $c_{g} = 10^5$ is a constant transforming from $m/s^2$ to $mGal$, $G$ is the Newton's gravitational constant $(m^3/kg \, s^2)$ and $\rho(x^{\prime}, y^{\prime}, z^{\prime})$ is the density at the point $(x^{\prime}, y^{\prime}, z^{\prime})$ inside the volume $V$ of the source.
@@
 \end{equation}
+
 where, $z_0 > z_i$.
@@
 \begin{equation}
-\delta g(x_{i}, y_{i}, z_{i}) = \sum_{j=1}^{N} p_j a_{ij} \; ,
+\delta g(x_{i}, y_{i}, z_{i}) = \sum_{j=1}^{M} p_j a_{ij}
 \label{eq:integral-sum}
 \end{equation}
+
 where $a_{ij}$ is given by:
@@
 \begin{equation}
-a_{ij}= c_{g} \, G \, \frac{(z_{j} - z_{i})}{\left[(x_{i} - x_{j})^{2} +	(y_{i} - y_{j})^{2} +	(z_{i} - z_{j})^{2} \right]^{\frac{3}{2}}}
+a_{ij}= c_{g} \, G \, \frac{(z_{j} - z_{i})}{\left[(x_{i} - x_{j})^{2} +	(y_{i} - y_{j})^{2} +	(z_{i} - z_{j})^{2} \right]^{\frac{3}{2}}},
 \label{eq:aij}
@@
 
-\noindent and $p_j$ is the coefficient representing the physical property of the $j$-th equivalent source.
+and $p_j$ is the coefficient representing the physical property of the $j$-th equivalent source.
 In this case, the harmonic function $a_{ij}$ represents the vertical component of the gravitational attraction exerted, at the point $(x_{i}, y_{i}, z_{i})$, by a point mass located at the point $(x_{j}, y_{j}, z_{j})$, with mass $p_{j}$.
@@
 \end{equation}
-where $\mathbf{d}(\mathbf{p})$ is an $N \times 1$ vector, whose i-th element is the predicted gravity disturbance $\delta g(x_{i}, y_{i}, z_{i})$, $\mathbf{p}$ is an $N \times 1$ parameter vector,  whose j-th element is the coefficient $p_{j}$ representing the physical property of the $j$-th equivalent source and $\mathbf{A}$ is  an $N \times N$ sensibility matrix, where each element $a_{ij}$ is given by equation 4 and defines the field produced by  the j-th equivalent source at the point $(x_{i}, y_{i}, z_{i})$.
+
+where $\mathbf{d}(\mathbf{p})$ is an $N \times 1$ vector, whose i-th element is the predicted gravity disturbance $\delta g(x_{i}, y_{i}, z_{i})$, $\mathbf{p}$ is an $M \times 1$ parameter vector,  whose j-th element is the coefficient $p_{j}$ representing the physical property of the $j$-th equivalent source and $\mathbf{A}$ is  an $N \times M$ sensibility matrix, where each element $a_{ij}$ is given by equation 4 and defines the field produced by  the j-th equivalent source at the point $(x_{i}, y_{i}, z_{i})$.
 The solution of the parameters \textbf{p} can be solved by a linear inversion minimizing the function:
@@
 \end{equation}
+
 where $\theta_g(\mathbf{p})$ is the misfit function given by:
@@
 \begin{equation}
-\theta_g(\mathbf{p}) = ||\mathbf{d}^{o}-\mathbf{d(p)}||_2^2 \: ,
+\theta_g(\mathbf{p}) = ||\mathbf{d^0}-\mathbf{d(p)}||_2^2 \: ,
 \label{eq:goal-function_d}
 \end{equation}
-which is the euclidian norm of the residual between the observed data $\mathbf{d}^{o}$ and the predicted data $\mathbf{d(p)}$.
+
+which is the euclidian norm of the residual between the observed data $\mathbf{d^0}$ and the predicted data $\mathbf{d(p)}$.
 
@@
 \end{equation}
+
 which is the euclidian norm of the parameters $\mathbf{p}$. The variable $\mu$ is a real positive number regularizing the parameter.
 
-Derivating equation \ref{eq:goal-function} in relation to $\mathbf{p}$ and making equal to zero, it is possible to estimate the parameters:
+Deriving equation \ref{eq:goal-function} in relation to $\mathbf{p}$ and making equal to zero, it is possible to estimate the parameters:
 
@@
 \begin{equation}
-\mathbf{p}^0 = \tilde{\mathbf{A}}^{-1} \mathbf{d}^{o} \: ,
+\mathbf{m}^0 = \tilde{\mathbf{A}}^{-1} \mathbf{d} \: ,
 \label{eq:initial_m}
 \end{equation}
-where $\mathbf{d}^o$ is the gravity data and $\tilde{\mathbf{A}}^{-1} = \Delta s/(2 \pi \, G \, c_g)$ with $\Delta s$ being an element of area located at the vertical coordinate $z_i$ and centered at the horizontal coordinates $(x_i,y_i), \, i = 1,..., N$.
+
+where $\mathbf{d}^o$ is the gravity data and $\tilde{\mathbf{A}}^{-1} = \Delta s/(2 \pi \gamma)$ with $\Delta s$ being an element of area located at the vertical coordinate $z_i$ and centered at the horizontal coordinates $(x_i,y_i), \, i = 1,..., N$.
 At each $k$-th iteration a mass correction vector $\mathbf{\Delta m^k}$ for every source is calculated by minimizing the function:
@@
 \begin{equation}
-\phi(\mathbf{\Delta p^k}) = ||\mathbf{d}^{o} - \mathbf{A}\hat{\mathbf{p}}^k - \tilde{\mathbf{A}}\mathbf{\Delta p^k}||_2^2
+\phi(\mathbf{\Delta m^k}) = ||\mathbf{d} - \mathbf{A}\hat{\mathbf{m}}^k - \tilde{\mathbf{A}}\mathbf{\Delta m^k}||_2^2 \: ,
 \label{eq:goal-function_fast}
@@
 
-\noindent and the mass distribution of the equivalent sources is updated as:
+and the mass distribution of the equivalent sources is updated as:
 
 \begin{equation}
-\mathbf{\Delta p^{k+1}} = \mathbf{p^{k}} + \mathbf{\Delta p^{k}} \: .
+\mathbf{\Delta m^{k+1}} = \mathbf{m^{k}} + \mathbf{\Delta m^{k}} \: ,
 \label{eq:update_m}
@@
 
-\lstset{language=Python}
-\lstset{frame=lines}
-\lstset{caption={A \textit{Python} algorithm for \cite{siqueira2017fast}'s fast equivalent layer.}}
-\lstset{label={lst:code_direct}}
-\lstset{basicstyle=\footnotesize}
-\begin{lstlisting}
-diagonal_A = G*(10**5)*2*pi/(dx*dy)
-for i in range (itmax):
-	res = (data - A.dot(p))
-	delta_p = res/diagonal_A
-	p += delta_p
-\end{lstlisting}
-
-Each iteration of this algorithm, the matrix-vector product $\tensor{A} \mathbf{p}$ must be calculated to get a new residual $\mathbf{d^0} - \tensor{A} \mathbf{p}$.
+Each iteration a matrix-vector product must be calculated to get a new residual.
 While it is true that for a small number of observation points this represents no computational effort, for very large data sets it is costful and can be overwhelming in terms of memory RAM to maintain such operation.
 
-\subsection{Structure of sensibility matrix A}
-If a gravity disturbance  data $\delta g(x_i, y_i, z_i)$ (equation \ref{eq:integral-sum}) is placed on a regular grid at the constant vertical coordinate $z_{i} = z_{1}$, $i = 1, \dots, N$, and there is an equivalent source located directly below each point of this grid at a constant depth $z_{0}$, the elements $a_{ij}$ (equation \ref{eq:aij}) can be rewritten as follows:
+\subsection{Toeplitz matrix structure}
+A Toeplitz matrix is one with constant diagonal, i.e., $\tensor{T} \in \tensor{R}^{N \times N} $ is Toeplitz if there exist scalars $r_{-N+1},\dots,r_{N-1}$ such that $r_{j-i}$ for all $i$ and $j$ \cite[]{golub2013,press2007numerical}.
+Only the first row and the first column of  $\tensor{T}$ needs to be known in order to construct the whole matrix.
 
 \begin{equation}
-a_{ij} = c_{g} G 
-\frac{\Delta z}{\left[(r_{ij})^{2} + (\Delta z)^{2}\right]^{\frac{3}{2}}} \: ,
-\label{eq:a_ij_r_ij}
+\tensor{T} =
+\pmatrix{
+r_{0} & r_{1} & \cdots & \cdots & r_{N-1} \cr
+r_{-1} & r_{0} & r_{1} & \cdots & \vdots \cr
+\vdots & r_{-1} & r_{0} & \ddots & \vdots \cr
+\vdots & \cdots & \ddots & \ddots & r_{1} \cr
+r_{1-N} & \cdots & \cdots &  r_{-1} & r_{0}
+}.
+\label{eq:toeplitz_T}
 \end{equation}
-where $\Delta z = z_{0} - z_{1}$ and $r_{ij} = \sqrt{(x_{i} - x_{j})^{2} + (y_{i} - y_{j})^{2}}$ represents the  relative horizontal distance between the $i$-th observation point and the $j$-th equivalent source.
-It is worth noting that: (i) $r_{ij} = r_{ji}$ for any pair $ij$ and (ii) $r_{ij}$ computed for different pairs $ij$ may assume the same value.
-As a consequence of these two properties, the matrix $\mathbf{A}$ (equation \ref{eq:predicted-data-vector}) has the structure of a symmetric Toeplitz by blocks, where each block is also a symmetric Toeplitz matrix (BTTB) \cite[]{chan2007introduction, golub2013}.
-For example, consider a regular grid of $N_{x} \times N_{y}$ points, where $N_{x} = 4$ and $N_{y} = 3$, comprising a total number of $N = N_{x} \times N_{y} = 12$ points.
-In this case, matrix $\mathbf{A}$ is $N_{x} \times N_{x}$ grid of matrices given by:
+
+\subsection{Structure of sensibility matrix A}
+If a gravity disturbance  data $\delta g(x_i, y_i, z_i)$ (equation \ref{eq:integral-sum}) is placed on a regular grid, the matrix $\mathbf{A}$ (equation \ref{eq:predicted-data-vector}) has the structure of a symmetric Toeplitz by blocks, where each block is also a symmetric Toeplitz matrix (BTTB) \cite[]{chan2007introduction}.
+For example, consider a regular grid of $N_{x} \times N_{y}$ points, where $N_{x} = 4$ and $N_{y} = 3$, comprising a total number of $N = N_{x} \, N_{y} = 12$ points.
+Consider also that there is an equivalent source located directly below each point of this grid.
+In this case, matrix $\mathbf{A}$ is $N \times N$ formed by a $N_{x} \times N_{x}$ grid of symmetric Toeplitz matrices, and each block is also a symmetric Toeplitz matrix with $N_{y} \times N_{y}$ elements:
 
@@
 \end{equation}
-where each block is a $N_{y} \times N_{y}$ matrix given by:
+
+the whole matrix is formed by $4 \times 4$ blocks, where each block is a $3 \times 3$ matrix with elements given by equation \ref{eq:aij}:
 
@@
 \pmatrix{
-	a_{0}^{\ell} & a_{1}^{\ell} & a_{2}^{\ell} \cr
-	a_{1}^{\ell} & a_{0}^{\ell} & a_{1}^{\ell} \cr
-	a_{2}^{\ell} & a_{1}^{\ell} & a_{0}^{\ell}
-},
+	a_{0} & a_{1} & a_{2} \cr
+	a_{1} & a_{0} & a_{1} \cr
+	a_{2} & a_{1} & a_{0}
+}.
 \label{eq:toeplitz_element}
 \end{equation}
-where the elements $a^{l}_{k}$, $l = 0, \dots, N_{x} - 1$, $k = 0, \dots, N_{y} - 1$ are computed with equation \ref{eq:a_ij_r_ij}.
-In this case, the complete matrix $\mathbf{A}$ can be obtained by computing only its first row or column.
 
-\subsection{BCCB matrix-vector product}
-As previous discussed in the  Fast equivalent layer technique, the matrix-vector product accounts for most of the computational cost (Listing \ref{lst:code_direct}).
-When large data sets are used, this operation can take some time and even be prohibited by memory RAM shortage.
-In order to lessen this problem we transform the BTTB matrix $\mathbf{A}$ into a Block-Circulating Circulating-Block (BCCB) matrix $\mathbf{C}$ and use its eigenvalues to carry the product of $\mathbf{A}$ and an arbitrary vector $\mathbf{p}$.
-This strategy has been successfully applied by Zhang and Wong (2015) and Zhang et al.
-(2016) for optmizing the computational cost of 3D gravity inversion and downward continuation of potential field, respectively.
-Here, we use this strategy for improving the computational efficiency of the Fast equivalent layer method proposed by Siqueira et al.
-(2017).
+Note that the case where matrix $\mathbf{A}$ is a symmetric BTTB with symmetric blocks, only its first row must be calculated.
 
-Following the example of our BTTB matrix $\mathbf{A}$ (equation \ref{eq:toeplitz_block_T}), its transformation into a BCCB matrix $\mathbf{C}$ is given by:
+\subsection{Circulant matrix structure}
+A circulant matrix is a special case of Toeplitz matrices where each row is a cyclic shift of the row above it as $c_{-k} = c_{n-k}$ for $1 \le k \le n-1$:
 
@@
 \pmatrix{
-	\tensor{C_{0}} & \tensor{C_{1}} & \tensor{C_{2}} & \tensor{C_{3}} & \tensor{0} & \tensor{C_{3}} & \tensor{C_{2}} & \tensor{C_{1}} \cr
-	\tensor{C_{1}} & \tensor{C_{0}} & \tensor{C_{1}} & \tensor{C_{2}} & \tensor{C_{3}} & \tensor{0} & \tensor{C_{3}} & \tensor{C_{2}} \cr
-	\tensor{C_{2}} & \tensor{C_{1}} & \tensor{C_{0}} & \tensor{C_{1}} & \tensor{C_{2}} & \tensor{C_{3}} & \tensor{0} & \tensor{C_{3}} \cr
-	\tensor{C_{3}} & \tensor{C_{2}} & \tensor{C_{1}} & \tensor{C_{0}} & \tensor{C_{1}} & \tensor{C_{2}} & \tensor{C_{3}} & \tensor{0} \cr
-	\tensor{0} & \tensor{C_{3}} & \tensor{C_{2}} & \tensor{C_{1}} & \tensor{C_{0}} & \tensor{C_{1}} & \tensor{C_{2}} & \tensor{C_{3}} \cr
-	\tensor{C_{3}} & \tensor{0} & \tensor{C_{3}} & \tensor{C_{2}} & \tensor{C_{1}} & \tensor{C_{0}} & \tensor{C_{1}} & \tensor{C_{2}} \cr
-	\tensor{C_{2}} & \tensor{C_{3}} & \tensor{0} & \tensor{C_{3}} & \tensor{C_{2}} & \tensor{C_{1}} & \tensor{C_{0}} & \tensor{C_{1}} \cr
-	\tensor{C_{1}} & \tensor{C_{2}} & \tensor{C_{3}} & \tensor{0} & \tensor{C_{3}} & \tensor{C_{2}} & \tensor{C_{1}} & \tensor{C_{0}}
-},
+	c_{0} & c_{n-1} & \cdots & c_{2} & c_{1} \cr
+	c_{1} & c_{0} & c_{n-1} & \cdots & c_{2} \cr
+	\vdots & c_{1} & c_{0} & \ddots & \vdots \cr
+	c_{n-2} & \cdots & \ddots & \ddots & c_{n-1} \cr
+	c_{n-1}  & c_{n-2} & \cdots & c_{1} & c_{0}
+}.
 \label{eq:circulant_elements}
 \end{equation}
-where each block $\mathbf{C}_{l}$ is:
+
+Equally a symmetric Toeplitz, the first row of $\tensor{C}_n$ comprises all the different elements of the matrix.
+An interesting recourse is that any $N \times N$ Toeplitz matrix can be embedded into a $2N \times 2N$ circulant matrix.
+For example, the Toeplitz matrix $\tensor{A}_{l}$ (equation \ref{eq:toeplitz_element}) is part of a circulant matrix:
 
@@
 \pmatrix{
-	a_{0}^{\ell} & a_{1}^{\ell} & a_{2}^{\ell} & 0^{\ell} & a_{2}^{\ell} & a_{1}^{\ell} \cr
-	a_{1}^{\ell} & a_{0}^{\ell} & a_{1}^{\ell} & a_{2}^{\ell} & 0^{\ell} & a_{2}^{\ell} \cr
-	a_{2}^{\ell} & a_{1}^{\ell} & a_{0}^{\ell} & a_{1}^{\ell} & a_{2}^{\ell} & 0^{\ell} \cr
-	0^{\ell} & a_{2}^{\ell} & a_{1}^{\ell} & a_{0}^{\ell} & a_{1}^{\ell} & a_{2}^{\ell} \cr
-	a_{2}^{\ell} & 0^{\ell} & a_{2}^{\ell} & a_{0}^{\ell} & a_{0}^{\ell} & a_{1}^{\ell} \cr
-	a_{1}^{\ell} & a_{2}^{\ell} & 0^{\ell} & a_{2}^{\ell} & a_{1}^{\ell} & a_{0}^{\ell}
+	a_{0} & a_{1} & a_{2} & 0 & a_{2} & a_{1} \cr
+	a_{1} & a_{0} & a_{1} & a_{2} & 0 & a_{2} \cr
+	a_{2} & a_{1} & a_{0} & a_{1} & a_{2} & 0 \cr
+	0 & a_{2} & a_{1} & a_{0} & a_{1} & a_{2} \cr
+	a_{2} & 0 & a_{2} & a_{0} & a_{0} & a_{1} \cr
+	a_{1} & a_{2} & 0 & a_{2} & a_{1} & a_{0}
 }.
@@
 
-Matrix $\mathbf{C}$ is a $4N_x N_y \times 4N_x N_y$ Block-Circulant formed by Circulant-Blocks matrix.
-This matrix is formed by a grid of $2N_x \times 2N_x$ blocks, where each block is a $2N_y \times 2N_y$ matrix.
+This recourse can be extended to a BTTB matrix embedding it into a Block-Circulant Circulant-Block (BCCB), where each block is a circulant matrix as shown in equation \ref{eq:toeplitz_circulant_element} and the whole matrix is block circulant.
+Thus, a $N \times N$ BTTB matrix can be embedded into a $4N \times 4N$ BCCB matrix.
 
-The product $\mathbf{d} = \mathbf{A}\mathbf{p}$ (equation \ref{eq:predicted-data-vector}), then becomes:
+Of many interesting properties of circulant matrix, one in particular will be very useful: the eigenvalues of a circulant matrix comprise the discrete Fourier transform (DFT) of the first column of this matrix \cite[]{grenander1984szeg}.
+Thus, the eigenvalues of a $\tensor{C}$ circulant matrix can be calculated by the fast Fourier transform (FFT) as:
 
 \begin{equation}
-\mathbf{C} \mathbf{v} = \mathbf{q} \: ,
+\Lambda_N = \sum_{j=0}^{N-1} c_j e^{\frac{2\pi \mathrm{i} jk}{N}}, \quad \mathrm{i} \equiv \sqrt{-1}
+\label{eq:eigenvalues}
+\end{equation}
+
+where $j,k = 0,1,...,N-1$ are the entries of the matrix $\tensor{C}$.
+
+\subsection{BCCB matrix-vector product}
+As previous discussed in the  Fast equivalent layer technique, the matrix-vector product accounts for most of the computational cost.
+When large data sets are used, this operation can take some time and even be prohibited by memory RAM shortage .
+In order to lessen this problem we propose the use of the Circulant embedding of a Block-Toeplitz Toeplitz-Block into a Block-Circulating Circulating-Block matrix and computing by FFT the eigenvalues of this matrix to carry the multiplication with the parameter vector \cite[]{strang1986introduction, olkin1986linear}.
+Following the notation presented by \cite{chan2007introduction} the multiplication of a Toeplitz matrix embedded into a circulant matrix and an N-dimensional vector $\mathbf{p}$ is:
+
+\begin{equation}
+\pmatrix{
+\tensor{T} & \times \cr
+\times & \tensor{T}
+}
+\pmatrix{
+\mathbf{p} \cr
+\tensor{0}
+} =
+\pmatrix{
+	\tensor{T} \mathbf{p} \cr
+	\dagger 
+},
+\label{eq:toeplitz_circulant_product}
+\end{equation}
+
+where $\pmatrix{\tensor{T} & \times \cr	\times & \tensor{T}}$ is a $2N \times 2N$ matrix (equation \ref{eq:toeplitz_circulant_element}), $\tensor{0}$ is an $N$-dimensional vector of zeros and $\dagger$ will be drop out.
+Following the property presented in the previous section, one can take the FFT of the first column of this new circulant matrix and make the multiplication with the new zero-padded vector.
+
+For a BTTB matrix-vector product, each Toeplitz block becomes a circulant block as in equation \ref{eq:toeplitz_circulant_product} and its correpondent vector is padded with zeros.
+The whole matrix then becomes a Block-Circulant, again padding the vector with zeros.
+As an example, the product between $N \times N$ matrix $\tensor{A}$ (equation \ref{eq:toeplitz_block_T}) and the $N$-dimensional vector $\mathbf{p}$ becomes:
+
+\begin{equation}
+\pmatrix{
+	\tensor{C}_0 & \tensor{C}_1 & \tensor{C}_2 & \tensor{0}_b & \tensor{C}_2 & \tensor{C}_1 & \cr
+	\tensor{C}_1 & \tensor{C}_0 & \tensor{C}_1 & \tensor{C}_2 & \tensor{0}_b & \tensor{C}_2 \cr
+	\tensor{C}_2 & \tensor{C}_1 & \tensor{C}_0 & \tensor{C}_1 & \tensor{C}_2 & \tensor{0}_b \cr
+	\tensor{0}_b & \tensor{C}_2 & \tensor{C}_1 & \tensor{C}_0 & \tensor{C}_1 & \tensor{C}_2 \cr
+	\tensor{C}_2 & \tensor{0}_b & \tensor{C}_2 &\tensor{C}_1 & \tensor{C}_0 & \tensor{C}_1 \cr
+	\tensor{C}_1 & \tensor{C}_2 & \tensor{0}_b & \tensor{C}_2 & \tensor{C}_1 & \tensor{C}_0
+}
+\pmatrix{
+	\mathbf{q} \cr
+	\tensor{0}
+} =
+\pmatrix{
+	\tensor{C} \mathbf{q} \cr
+	\dagger
+},
 \label{eq:BCCB_vector_product}
 \end{equation}
-where $\mathbf{v}$ and $\mathbf{q}$ are $4N_x N_y \times 1$ vectors given by:
+
+where the matrix BCCB has $4N \times 4N$ dimension, each block $\tensor{C}_l$ is a circulant matrix defined in \ref{eq:toeplitz_circulant_element} and $\tensor{0}_b$ is a matrix of zeros with the same dimension of $\tensor{C}_l$.
+The new vector is $4N$, with $\mathbf{q}$ being a vector $2N$ padded with zeros in correct place, correspondent with the block sizes, and $\tensor{0}$ is a vector of zeros also $2N$.
+Despite this increase in dimension, only the first column of the matrix needs to be calculated.
+As before, the FFT will be applied to compute the eigenvalues of this new circulant matrix (the FFT of the first column of the matrix obtained in equation \ref{eq:BCCB_vector_product}) and also apply on the new parameter vector $\pmatrix{\mathbf{q} \cr \tensor{0}}$.
+To complete the process the inverse FFT must be applied:
 
 \begin{equation}
-\mathbf{v} =
-\pmatrix{
-	\mathbf{v}_{0} \cr
-	\mathbf{v}_{1} \cr
-	\vdots \cr
-	\mathbf{v}_{N_x- 1} \cr
-	\mathbf{0}_{(2 N_x N_y)}
-}
+\pmatrix{\tensor{C} \mathbf{q}\cr\dagger} = \mathcal{F}^{-1}[\Lambda_N * \mathcal{F}(\pmatrix{\mathbf{q} \cr \tensor{0}})],
+\label{eq:fft_product}
 \end{equation}
 
-\noindent and
+where $\Lambda_N$ is the eigenvalues of matrix BCCB defined in equation \ref{eq:eigenvalues}.
+By dropping out $\dagger$ we have the vector $\tensor{C}\mathbf{q}$ with $2N$ dimension.
+By once more dropping the correct leftovers correspondent of each block, the matrix-vector product $\tensor{A}\mathbf{p}$ is complete.
 
-\begin{equation}
-\mathbf{q} =
-\pmatrix{
-	\mathbf{q}_{0} \cr
-	\mathbf{q}_{1} \cr
-	\vdots \cr
-	\mathbf{q}_{N_x - 1} \cr
-	\mathbf{0}_{(2 N_x N_y)}
-} \: ,
-\end{equation}
-where $\mathbf{0}_{(2 N_x N_y)}$ is a $2 N_x N_y \times 1$ vetor of zeros, and $\mathbf{v}_{\ell}$ and $\mathbf{q}_{\ell}$, $\ell = 0, \dots, N_x-1$ are:
-
-\begin{equation}
-\mathbf{v}_{\ell} =
-\pmatrix{
-	\mathbf{p}_{\ell} \cr
-	\mathbf{0}_{(N_y)}
-}
-\end{equation}
-
-\noindent and
-
-\begin{equation}
-\mathbf{q}_{\ell} =
-\pmatrix{
-	\mathbf{d}_{\ell} \cr
-	\mathbf{0}_{(N_y)}
-} \: ,
-\end{equation}
-where $\mathbf{0}_{(N_y)}$ is a $N_y \times 1$ vetor of zeros.
-
-By using the Kronecker product properties, the auxiliary matrix-vetor product can be rewritten as follows:
-
-\begin{equation}
-\mathbf{F}_{(2 N_x)}^{\ast} \left[ \mathbf{L} \circ \left( \mathbf{F}_{(2N_x)} \mathbf{V} \, \mathbf{F}_{(2N_y)} \right) \right] \mathbf{F}_{(2N_y)}^{\ast} = \mathbf{Q} \: ,
-\label{eq:fft_q}
-\end{equation}
-where ``$\circ$'' denotes the Hadamard product, $\mathbf{L}$ is a $2N_x \times 2N_y$ row-oriented matrix containing the eigenvalues of $\mathbf{C}_{(BCCB)}$, and $\mathbf{V}$ and $\mathbf{Q}$ are $2N_x \times 2N_y$ row-oriented matrices obtained from the vectors $\mathbf{v}$ and $\mathbf{q}$.
-
-One of the properties of BCCB matrices is that its eigenvalues can be calculated by a 2D Discrete Fourier Transform \cite[]{chan2007introduction}.
-This lead to a fast calculation of $\mathbf{L}$ using a 2D Fast Fourier Transform, making the matrix-vector product a low computational cost process.
-
-Note that in general, the first column of blocks forming a BCCB matrix $\mathbf{C}_{\ell}$ (equation \ref{eq:circulant_elements}) is given by:
-
-\begin{equation}
-\left[\mathbf{C} \right]_{(0)} = 
-\pmatrix{
-\mathbf{C}_{0} \cr
-\mathbf{C}_{1} \cr
-\vdots \cr
-\mathbf{C}_{N_x-2} \cr
-\mathbf{C}_{N_x-1} \cr
-\mathbf{0} \cr
-\mathbf{C}_{N_x-1} \cr
-\mathbf{C}_{N_x-2} \cr
-\vdots \cr
-\mathbf{C}_{1}
-} \; ,
-\end{equation}
-where each block $\mathbf{C}_{\ell}$, $\ell = 0, \dots, N_x-1$, is a $2N_y \times 2N_y$ circulant matrix and $\mathbf{0}$ is a $2N_y \times 2N_y$ matrix with all elements equal to zero.
-Thus, the first column of a circulant matrix $\mathbf{C}_{\ell}$ is given by:
-
-\begin{equation}
-\left[\mathbf{C}_{\ell} \right]_{(0)} =
-\pmatrix{
-	a^{\ell}_{00} \cr
-	a^{\ell}_{10} \cr
-	\vdots \cr
-	a^{\ell}_{(N_y-2)0} \cr
-	a^{\ell}_{N_y-1) 0} \cr
-	0 \cr
-	a^{\ell}_{N_y-1) 0} \cr
-	a^{\ell}_{(N_y-2)0} \cr
-	\vdots \cr
-	a^{\ell}_{10}
-} \:.
-\end{equation}
-
-To complete the process, after calculating the inverse of $\mathbf{Q}$ it is necessary to rearrange its rows to obtain the vector $\mathbf{q}$ and also rearrange the elements of $\mathbf{q}$ to obtain the wanted vector $\mathbf{d(p)}$.
-
-%The algorithm for computing the matrix-vector product $\mathbf{A}\mathbf{p}$ and obtaining the resulting vector $\mathbf{d(p)}$ can be summarized in the following steps:
-
-%1. Compute the first column of the embedded $4N_x N_y \times 4N_x N_y$ BCCB matrix $\mathbf{C}$ by using the first column of the $N_x N_y \times N_x N_y$ BTTB matrix $\mathbf{A}$;
-
-%2. Compute the eigenvalues of $\mathbf{C}$ rearranging them in rows of the $2N_x \times 2N_y$ matrix $\mathbf{L}$;
-
-%3. Rearrange the vector $\mathbf{v}$ in the rows of the $2N_x \times 2N_y$ matrix $\mathbf{V}$;
-
-%4.
-Compute the Hadammard product $\mathbf{L} \circ \tilde{\mathbf{P}}$, where $\tilde{\mathbf{P}} = \mathbf{F}_{(2N_x)} \mathbf{P} \, \mathbf{F}_{(2N_y)}$ is computed by using a fast algorithm for 2D DFT;
-
-%5.
-Compute the resulting $2N_x \times 2N_y$ matrix $\mathbf{Q} = \mathbf{F}_{(2N_x)}^{\ast} \left[ \mathbf{L} \circ \tilde{\mathbf{P}}\right] \mathbf{F}_{(2N_y)}^{\ast}$ by using a fast algorithm for inverse 2D DFT;
-
-%6. Rearrange the rows of $\mathbf{Q}$ to obtain the vector $\mathbf{q}$;
-
-%7. Rearrange the elements of $\mathbf{q}$ to obtain the vector $\mathbf{d(p)}$.
-
-\subsection{Computational performance}
+\subsection{Computation performance}
 Equation \ref{eq:goal-function_fast} now can be calculated with the approach presented in the previous section.
-In a normal procedure of the fast equivalent layer, at each iteration a full matrix $\tensor{A}$ (equation \ref{eq:predicted-data-vector}) is multiplied by the estimated mass distribution parameter vector $\hat{\mathbf{p}}^k$.
+In a normal procedure of the fast equivalent layer, at each iteration a full matrix $\tensor{A}$ (equation \ref{eq:predicted-data-vector}) is multiplied by the estimated mass distribution parameter vector $\hat{\mathbf{m}}^k$.
 As pointed in \cite{siqueira2017fast} the number of flops (floating-point operations) necessary to estimate the $N$-dimensional parameter vector inside the iteration loop is:
@@
 It is well known that FFT takes $N \log_2(N)$ flops \cite[]{brigham1988fast}.
-Computing the eigenvalues of the BCCB matrix ($4N \times 4N$) and applying 2D-FFT on the parameter vector (equation \ref{eq:fft_q}), takes $4N \log(4N)$ each.
-The point-multiplication takes $4N$.
-As it is necessary to compute the inverse FFT another two $4N \log(4N)$ must be taken in account.
-However, the sensibility matrix does not change during the process, thus, the eigenvalues of BCCB must be calculated only once, outside of the iteration.
+Computing the eigenvalues of the BCCB matrix ($4N \times 4N$) and applying FFT on the parameter vector (equation \ref{eq:BCCB_vector_product}), takes $4N \log(4N)$ each.
+As it is necessary to multiply and compute the inverse FFT (equation \ref{eq:fft_product}) another two $4N \log(4N)$ must be taken in account.
+However, the sensibility matrix does not change during the processes, thus, the eigenvalues of BCCB must be calculated only once, outside of the iteration.
 This lead us to:
@@
 \begin{equation}
-f_1 = N^{it} (7N + 12N\log(4N)).
+f_1 = N^{it} (3N + 12N\log(4N)).
 \label{eq:float_bccb}
@@
 \begin{equation}
-f_s = 12N + 4N \log(4N) + N^{it} (7N + 12N\log(4N)).
+f_s = 12N + 4N \log(4N) + N^{it} (3N + 12N\log(4N)).
 \label{eq:float_new}
@@
 
-Using a PC with a Intel Core i7 4790@3.6GHz processor and 16 Gb of memory RAM, figure \ref{fig:time_comparison} shows the time necessary to run 50 iterations of the \cite{siqueira2017fast}'s method and the one presented in this work.
+Using a PC with a Intel Core i7 4790@3.6GHz processor and 16 Gb of memory ram, figure \ref{fig:time_comparison} shows the time necessary to run 50 iterations of the \cite{siqueira2017fast}'s method and the one presented in this work.
 After $10000$ observations points is clear how this method benefits from the new approach in calculating the forward problem.
-Because of the memory RAM available in this system, we could not test the comparison with more observations, limited to $22500$.
+Because of the memory RAM available in this system, we could not test with more observations, limited to $22500$.
 In figure \ref{fig:time_bccb} we show the time necessary to run the equivalent layer technique with 50 iterations using only the new approach, where the RAM is not a limitation factor.
 We could run up to $25$ million observation points.
-In comparison, one million observation points took $26.8$ seconds to run, where the maximum $22500$ observation points, with \cite{siqueira2017fast}'s method, took $48.3$ seconds.
+In comparison $1$ million observation points took $26.8$ seconds to run, where the maximum $22500$ observation points, with \cite{siqueira2017fast}'s method, took $48.3$ seconds.
 
@@
 A gaussian noise was added to the data, with mean of zero and a maximum of $0.5\%$ of the maximum value of the original data.
-As previous said, only in regular grids the BTTB matrix structures appears.
+As previous said only in regular grids the BTTB matrix structures appears.
 We created $10000$ observation points regularly spaced in a grid of $100 \times 100$, with an uniform $100$ m of height for all the observations.
@@
 
-Synthetic and real data tests were conducted with the equivalent layer using the method presented in this work, showing satisfatory results that evaluate it to process gravity data, as upward and downward continuations, even using very large datasets.
+Synthetic tests and real data were used with the equivalent layer presented in this work, with results that evaluate the technique to be used with large datasets to process gravity data, as upward and downward continuations with satisfatory results.
 In the future, applications of the equivalent layer using great amount of data, as in continental or global scale can be researched.
@@
 \plot{delta_comparison}{width=13.5cm}
-{virtually zero difference in the fitted data presented in figures \ref{fig:classic_fast}b and \ref{fig:bccb_fast}c, showing that matrix-vector product of a embedded BTTB into a BCCB matrix using equation \ref{eq:BCCB_vector_product} is pratically equal to a normal matrix-vector product.}
+{small difference in the fitted data presented in figures \ref{fig:classic_fast}b and \ref{fig:bccb_fast}c, showing that there is error associated in calculating a matrix-vector product of a embedded BTTB into a BCCB matrix using equation \ref{eq:BCCB_vector_product}.
+The highest values are on the borders.}
 \newpage
@@
 \plot{delta_rho}{width=13.5cm}
-{small difference in the estimated density between the classic matrix-vector product and the matrix-vector product of a embedded BTTB into a BCCB matrix using equation \ref{eq:BCCB_vector_product}, showing that the estimative using this work's method is pratically the same.}
+{small difference in the fitted data presented in figures \ref{fig:classic_fast}b and \ref{fig:bccb_fast}c, showing that there is error associated in calculating a matrix-vector product of a embedded BTTB into a BCCB matrix using equation \ref{eq:BCCB_vector_product}.
+The highest values are on the borders.}
 \newpage

### 2019-07-16 manuscript/toeplitz_manuscript_rev_2.tex

\documentclass[paper]{geophysics}
%\documentclass[paper,revised]{geophysics}
\usepackage{listings}
% An example of defining macros
\newcommand{\rs}[1]{\mathstrut\mbox{\scriptsize\rm #1}}
\newcommand{\rr}[1]{\mbox{\rm #1}}

\begin{document}

\title{Fast equivalent layer technique for gravity data processing with Block-Toeplitz Toeplitz-Block matrix systems}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\ms{GEO-Example} % paper number

\address{
\footnotemark[1]Observat\'orio Nacional, \\
77 General Jos\'e Cristino Av., \\
Rio de Janeiro, RJ, 20921400 \\
}
\author{Diego Takahashi\footnotemark[1], Vanderlei C. Oliveira Jr.\footnotemark[1] and Val\'eria
	C. F. Barbosa\footnotemark[1]}

%\footer{Example}
\lefthead{Takahashi, Oliveria Jr. \& Barbosa}
\righthead{\emph{Geophysics}}

\maketitle

\begin{abstract}
We have developed an efficient and very fast equivalent layer technique for gravity data processing by modifying an iterative method grounded on excess mass constraint that does not require the solution of linear systems. Taking advantage of the properties related to the symmetric Block-Toeplitz Toeplitz-block (BTTB) and Block-Circulant Circulant-Block (BCCB) matrices, that raises when regular grids of observation points and equivalent sources (point masses) are used to set up a fictitious equivalent layer, we have developed an algorithm which greatly 
reduces the number of flops and memory RAM necessary to estimate of a 2D mass distribution over the equivalent layer. The algorithm is based on the structure of symmetric BTTB matrix, where all its elements consist of the elements of the first row, which in turn can be embedded into a symmetric BCCB matrix. Likewise, only the first row of the BCCB matrix is needed to reconstruct the full matrix completely. From the first column of BCCB matrix, its eigenvalues can be calculated using the fast Fourier transform, which can be used to readily compute the matrix-vector product. As a result, our method is efficient to process very large datasets using either fine- or mid-grid meshes. The larger the dataset, the faster and more efficient our method becomes compared to the available equivalent-layer techniques. Synthetic tests demonstrate the ability of our method to satisfactorily upward- and downward-continuing the gravity data. Test with real data from Caraj\'as, Brazil, shows its applicability to process very large dataset at low computational cost.
\end{abstract}

\section{Introduction}

The equivalent layer is a well-known technique for processing potential-field data in applied geophysics since 1960. It comes from potential theory as a mathematical solution of the Laplace's equation, in the region above the sources, by using the Dirichlet boundary condition \cite[]{kellogg1929}.
This theory states that any potential-field data produced by an arbitrary 3D physical property distribution can be exactly reproduced by a fictitious layer located at any depth and having a continuous 2D physical property  distribution. In pratical situations, the layer is approximated by a finite set of sources (e.g., point masses or dipoles) and their physical properties are estimated by solving a linear system of equations that yield an acceptable potential-field data fit. These fictitious sources are called equivalent sources.

Many previous works have used the equivalent layer as a processing technique of potential-field data. \cite{dampney1969equivalent} used the equivalent-layer technique for gridding and for computing the upward continuation of the potential-field data. \cite{cordell1992} and \cite{mendonca-silva1994} used it for interpolating and gridding potential-field data. \cite{emilia1973}, \cite{hansen-miyazaki1984} and \cite{li-oldenburg2010} used it for upward continuation of the potential-field data. \cite{silva1986}, \cite{leao-silva1989}, \cite{guspi-novara2009}, and \cite{oliveirajr-etal2013} used it for reducing the magnetic data to the pole. \cite{boggs-dransfield2004} used it for combining multiple data sets and Barnes and Lumley (2011) for gradient-data processing.

The classic equivalent layer formulation consists in estimating the physical-property distribution within a layer, composed by a set of equivalent sources, by solving a linear system of equations formed by harmonic functions (e.g.,the inverse of the distance between the observation point and the equivalent source). When these observation points and equivalent sources are regularly spaced, a Toeplitz system arises. Toeplitz systems are well-known in many branches of science as in (1) mathematics, for solving partial and ordinary diferential equations (e.g., \cite{lin2003strang}); (2) image processing (e.g., \cite{chan1999cosine}) and; (3) computational neuroscience (e.g., \cite[]{wray1994calculation}). \cite{jin2003developments} and \cite{chan2007introduction} give many examples of applications for Toeplitz systems.

In potential-field methods, the properties of Toeplitz system were used for downward continuation \cite[]{zhang2016bttb} and for 3D gravity-data inversion using a 2D multilayer model \cite[]{zhang2015bttb}. In the particular case of gravity data, the kernel generates a linear system with a matrix known as symmetric Block-Toeplitz Toeplitz-Block (BTTB).

A wide variety of applications in mathematics and engineers that fall into Toeplitz systems propelled the development of a large variety of  methods for solving them. Direct methods were conceived by \cite{levinson1946wiener} and by \cite{trench1964algorithm}. Currently the conjugate gradient is used in most cases. In \cite{grenander1984szeg}, Szeg\"{o} noticed that a circulant matrix can be diagonalized by taking the fast Fourier trasform of its first column, making it possible to calculate the matrix-vector product and solve the system with low computational cost \cite[]{strang1986introduction,olkin1986linear}. \cite{chan2007introduction} show some preconditioners to embedd the Toeplitz and BTTB matrices into, respectively, circulant matrices and Block-Circulant Circulant-Block (BCCB) by solving the system applying the conjugate gradient method.

Although the use of the equivalent-layer technique increased over the last decades, one of the biggest problem is still its high computational cost for processing large-data sets. \cite{siqueira2017fast} developed a computationally efficient scheme for processing gravity data. This scheme does not solve a linear system, instead uses an iterative process that corrects the physical-property distribution over the equivalent layer by adding mass corrections that are proportional to the gravity residual. Although efficient, the Siqueira et al.'s (2017) method requires, at each iteration,  the full computation of the forward problem to guarantee the convergence of the algorithm. The time spent on forward modeling accounts for most of the total computation time of Siqueira et al.'s (2017) method.

We propose the use of BTTB and BCCB matrices properties to efficiently  solve the forward modeling in Siqueira et al.'s (2017) method, resulting in faster parameter estimation and the possibility to use very large datasets. Here, we show how the system memory (RAM) usage can be drastically decreased by calculating only the first row of the BTTB matrix and embedding into a BCCB matrix. Using the Szeg\"{o} theorema  combined with \cite{strang1986introduction}, the matrix-vector product can be accomplished with very low cost, reducing in some orders of magnitude the number of operations required to complete the process. We present synthetic tests to validate our proposal and real field data from Caraj\'as, Brazil to demonstrate its applicability.

\section*{Methodology}
\subsection{Equivalent layer theory for gravity data}
In applied geophysics, the observed gravity disturbance $\delta g(x, y, z)$ \citep{heiskanen1967physical} is commomly approximated as the vertical component $z$ of the gravitational attraction produced by gravity sources, as follows:

\begin{equation}
\delta g(x, y, z) = c_{g} \, G \, \int\int\limits_{v}\int \rho(x^{\prime}, y^{\prime}, z^{\prime}) \frac{(z^{\prime} - z) \quad dx^{\prime}dy^{\prime}dz^{\prime}}{[(x - x^{\prime})^{2} + (y - y^{\prime})^{2} + (z - z^{\prime})^{2}]^{\frac{3}{2}}}\: ,
\label{eq:gravity-disturbance}
\end{equation}
where $c_{g} = 10^5$ is a constant transforming from $m/s^2$ to $mGal$, $G$ is the Newton's gravitational constant $(m^3/kg \, s^2)$ and $\rho(x^{\prime}, y^{\prime}, z^{\prime})$ is the density at the point $(x^{\prime}, y^{\prime}, z^{\prime})$ inside the volume $V$ of the source. This integral is defined in a Cartesian coordinate system with axis $x$ pointing to north, $y$ pointing to east and $z$ pointing downward.

According to the classic upward continuation integral \citep{henderson1960comprehensive, henderson1970validity}, it is possible to compute the gravity disturbance $\delta g(x_{i}, y_{i}, z_{h})$, at a point $(x_i, y_i, z_h)$, from the gravity disturbance $\delta g(x_j, y_j, z_{0})$ at the constant plane $z_{0}$:

\begin{equation}
\delta g(x_i, y_i, z_h) = \frac{1}{2\pi}  \int\limits_{-\infty}^{+\infty} \int\limits_{-\infty}^{+\infty} \,
\frac{\delta g(x_j, y_j, z_0) (z_{0} - z_{h})}{\left[(x_{i} - x_j)^{2} +
	(y_{i} - y_j)^{2} +
	(z_{h} - z_{0})^{2} \right]^{\frac{3}{2}}} \, 
dx \, dy,
\label{eq:upward-continuation-integral}
\end{equation}
where, $z_0 > z_h$.
Equation \ref{eq:upward-continuation-integral} shows that the gravity disturbance $\delta g(x_{i}, y_{i}, z_{h})$ is a convolution between $\delta g(x_j, y_j, z_{0})$ and another harmonic function on the horizontal plane $z=z_0$.
Multiplying and dividing equation \ref{eq:upward-continuation-integral} by $G$ and discretizing the integral, we obtain:

\begin{equation}
\delta g(x_{i}, y_{i}, z_{h}) = \sum_{j=1}^{N} p_j a_{ij} \; ,  \quad i,j = 1,\dots,N,
\label{eq:integral-sum}
\end{equation}
where $p_j$ is the coefficient representing the physical property of the $j$th equivalent source:

\begin{equation}
p_j = \frac{\Delta S_j \delta g(x_{j}, y_{j}, z_{0})}{2 \pi c_g G} \; ,
\label{eq:p_j}
\end{equation}

\noindent with $\Delta s$ being an element of area located at the vertical coordinate $z_h$ and centered at the horizontal coordinates $(x_i,y_i)$, and $a_{ij}$ is given by:

\begin{equation}
a_{ij}= c_{g} \, G \, \frac{(z_{0} - z_{h})}{\left[(x_{i} - x_{j})^{2} +	(y_{i} - y_{j})^{2} +	(z_{h} - z_{0})^{2} \right]^{\frac{3}{2}}} \; .
\label{eq:aij}
\end{equation}

In this case, the harmonic function $a_{ij}$ represents the vertical component of the gravitational attraction exerted, at the point $(x_{i}, y_{i}, z_{h})$, by a point mass located at the point $(x_{j}, y_{j}, z_{0})$, with mass $p_{j}$.

Equation \ref{eq:integral-sum} can be expressed in matrix form as:

\begin{equation}
\mathbf{d}(\mathbf{p}) = \mathbf{A} \mathbf{p} \: ,
\label{eq:predicted-data-vector}
\end{equation}
where $\mathbf{d}(\mathbf{p})$ is an $N \times 1$ vector, whose $i$th element is the predicted gravity disturbance $\delta g(x_{i}, y_{i}, z_{h})$, $\mathbf{p}$ is an $N \times 1$ parameter vector,  whose $j$th element is the coefficient $p_{j}$ (equation \ref{eq:p_j}) representing the physical property of the $j$th equivalent source and $\mathbf{A}$ is  an $N \times N$ sensibility matrix, where each element $a_{ij}$ is given by equation \ref{eq:aij} and defines the vertical component of gravity field produced by  the $j$th equivalent source with unitary density at the observation point $(x_{i}, y_{i}, z_{h})$.
The solution of the parameters \textbf{p} can be solved by a linear inversion minimizing the function:

\begin{equation}
\Psi(\mathbf{p}) = \theta_g(\mathbf{p}) + \mu \, \theta_m(\mathbf{p}) \: ,
\label{eq:goal-function}
\end{equation}
where $\theta_g(\mathbf{p})$ is the misfit data function given by:

\begin{equation}
\theta_g(\mathbf{p}) = ||\mathbf{d}^{o}-\mathbf{d(p)}||_2^2 \: ,
\label{eq:goal-function_d}
\end{equation}
which is the Euclidian norm of the residual defined as the difference between the observed data $\mathbf{d}^{o}$ and the predicted data $\mathbf{d(p)}$.

The function $\theta_m(\mathbf{p})$ is the regularization function, for example, the zeroth-order Tikhonov \cite[]{tikhonov-arsenin1977}:

\begin{equation}
\theta_m(\mathbf{p}) = ||\mathbf{p}||_2^2 \: ,
\label{eq:tikhonov-function}
\end{equation}
which is the Euclidian norm of the parameters $\mathbf{p}$. The variable $\mu$ in equation \ref{eq:goal-function} is a real positive number called regularizing parameter.

Derivating equation \ref{eq:goal-function} in relation to $\mathbf{p}$ and making the result equal to zero, it is possible to estimate the parameters as:

\begin{equation}
\hat{\mathbf{p}} = \left( \mathbf{A}^{\top}\mathbf{A} + 
\mu \, \mathbf{I} \right)^{-1}
\mathbf{A}^{\top} \mathbf{d}^{o} \,.
\label{eq:p-ast-parameter-space}
\end{equation}

\subsection{Fast equivalent layer technique}
In \cite{siqueira2017fast}, the authors developed an iterative least-squares method to estimate the mass distribution over the equivalent layer based on the excess of mass and the positive correlation between the observed gravity data and the masses on equivalent sources. This scheme is proven to have a better computational efficiency than the classical equivalent layer approach with data sets of at least 200 observation points, even using a large number of iterations.
This work also proves that the excess of mass of a body is proportional to the surface integration of the gravity data on the horizontal plane $z_0$ where the equivalent sources are distributed. Considering each equivalent source directly beneath the observation points, a initial approximation of mass distibution is given by:

\begin{equation}
\mathbf{p}^0 = \tilde{\mathbf{A}}^{-1} \mathbf{d}^{o} \: ,
\label{eq:initial_m}
\end{equation}
where $\mathbf{d}^o$ is the gravity data and $\tilde{\mathbf{A}}^{-1}$ is an $N \times N$ diagonal matrix whose elements are:

\begin{equation}
\tilde{\mathbf{A}}^{-1} = \frac{\Delta s}{(2 \pi \, G \, c_g)} \: .
\label{eq:a_approx}
\end{equation}

At each $k$th iteration, a mass correction vector $\mathbf{\Delta p^k}$ for every source is calculated by minimizing the function:

\begin{equation}
\phi(\mathbf{\Delta p^k}) = ||\mathbf{d}^{o} - \mathbf{A}\hat{\mathbf{p}}^k - \tilde{\mathbf{A}} \mathbf{\Delta} \hat{\mathbf{p}^k}||_2^2
\label{eq:goal-function_fast}
\end{equation}

\noindent and the mass distribution of the equivalent sources is updated as:

\begin{equation}
\mathbf{\Delta} \hat{\mathbf{p}}^{k+1} = \hat{\mathbf{p}}^{k} + \mathbf{\Delta} \hat{\mathbf{p}}^{k} \: .
\label{eq:update_m}
\end{equation}

%\lstset{language=Python}
%\lstset{frame=lines}
%\lstset{caption={A \textit{Python} algorithm for Siqueira et al.'s (2017) fast equivalent layer.}}
%\lstset{label={lst:code_direct}}
%\lstset{basicstyle=\footnotesize}
%\begin{lstlisting}
%diagonal_A = G*(10**5)*2*pi/(dx*dy)
%for i in range (itmax):
%	res = (data - A.dot(p))
%	delta_p = res/diagonal_A
%	p += delta_p
%\end{lstlisting}

At the $k$th iteration of Siqueira et al.'s (2017) method, the matrix-vector product $\tensor{A} \hat{\mathbf{p}}^{k}$ must be calculated to get a new residual $\mathbf{d^0} - \tensor{A} \hat{\mathbf{p}}^{k}$. While it is true that for a small number of observation points this represents no computational effort, for very large data sets it is costful and can be overwhelming in terms of memory RAM to maintain such operation.

\subsection{Structure of sensibility matrix A}
If a gravity disturbance  data $\delta g(x_i, y_i, z_h)$ (equation \ref{eq:integral-sum}) $i = 1, \dots, N$, is placed on a regular grid at the constant vertical coordinate $z_{h} = z_{1}$ and there is a set of equivalent sources each one located directly below each point of this grid at a constant depth $z_{0}$, the elements $a_{ij}$ (equation \ref{eq:aij}) can be rewritten as follows:

\begin{equation}
a_{ij} = c_{g} G 
\frac{\Delta z}{\left[(r_{ij})^{2} + (\Delta z)^{2}\right]^{\frac{3}{2}}} \: ,
\label{eq:a_ij_r_ij}
\end{equation}
where $\Delta z = z_{0} - z_{1}$ and $r_{ij} = \sqrt{(x_{i} - x_{j})^{2} + (y_{i} - y_{j})^{2}}$ represents the  relative horizontal distance between the $i$th observation point and the $j$th equivalent source. It is worth noting that: (i) $r_{ij} = r_{ji}$ for any pair $ij$ and (ii) $r_{ij}$ computed for different pairs $ij$ may assume the same value. As a consequence of these two properties, the matrix $\mathbf{A}$ (equation \ref{eq:predicted-data-vector}) has the structure of a symmetric Toeplitz by blocks, where each block is also a symmetric Toeplitz matrix (BTTB) \cite[]{chan2007introduction, golub2013}. 
For example, consider a regular grid of $N_{x} \times N_{y}$ points, where $N_{x} = 4$ and $N_{y} = 3$, comprising a total number of $N = N_{x} \times N_{y} = 12$ points. In this case, matrix $\mathbf{A}$ is a partitioned matrix by $N_{x} \times N_{x}$ blocks:

\begin{equation}
\tensor{A} =
\pmatrix{
	\tensor{A_{0}} & \tensor{A_{1}} & \tensor{A_{2}} & \tensor{A_{3}} \cr
	\tensor{A_{1}} & \tensor{A_{0}} & \tensor{A_{1}} & \tensor{A_{2}} \cr
	\tensor{A_{2}} & \tensor{A_{1}} & \tensor{A_{0}} & \tensor{A_{1}} \cr
	\tensor{A_{3}} & \tensor{A_{2}} & \tensor{A_{1}} & \tensor{A_{0}}
},
\label{eq:toeplitz_block_T}
\end{equation}
where each block is a $N_{y} \times N_{y}$ matrix given by:

%\begin{equation}
%\tensor{A}_{l} =
%\pmatrix{
%	a_{0}^{\ell} & a_{1}^{\ell} & a_{2}^{\ell} \cr
%	a_{1}^{\ell} & a_{0}^{\ell} & a_{1}^{\ell} \cr
%	a_{2}^{\ell} & a_{1}^{\ell} & a_{0}^{\ell}
%},
%\label{eq:toeplitz_element}
%\end{equation}

\begin{equation}
\tensor{A}_{0} =
\pmatrix{
	a_{0}^{\ell} & a_{1}^{\ell} & a_{2}^{\ell} \cr
	a_{1}^{\ell} & a_{0}^{\ell} & a_{1}^{\ell} \cr
	a_{2}^{\ell} & a_{1}^{\ell} & a_{0}^{\ell}
},
\label{eq:toeplitz_element}
\end{equation}
where the elements $a^{l}_{k}$, $l = 0, \dots, N_{x} - 1$, $k = 0, \dots, N_{y} - 1$ are computed with equation \ref{eq:a_ij_r_ij}. Here, the most striking feature is that the complete BTTB matrix $\mathbf{A}$ can be obtained by computing only its first row or column.

\subsection{BCCB matrix-vector product}
As previous discussed in the  fast equivalent layer technique section, the matrix-vector product accounts for most of the computational cost. When large data sets are used, this operation can take some time and even be prohibited by memory RAM shortage. In order to lessen this problem we transform the BTTB matrix $\mathbf{A}$ into a Block-Circulating Circulating-Block (BCCB) matrix $\mathbf{C}$ and use its eigenvalues to carry the product of $\mathbf{A}$ and an arbitrary vector $\mathbf{p}$. This strategy has been successfully applied by Zhang and Wong (2015) and Zhang et al. (2016) for optmizing the computational cost of 3D gravity inversion and downward continuation of potential field, respectively. Here, we use this strategy for improving the computational efficiency of the fast equivalent layer method proposed by \cite{siqueira2017fast}.

Following the example of our BTTB matrix $\mathbf{A}$ (equation \ref{eq:toeplitz_block_T}), its transformation into a BCCB matrix $\mathbf{C}$ is given by:

\begin{equation}
\tensor{C} =
\pmatrix{
	\tensor{C_{0}} & \tensor{C_{1}} & \tensor{C_{2}} & \tensor{C_{3}} & \tensor{0} & \tensor{C_{3}} & \tensor{C_{2}} & \tensor{C_{1}} \cr
	\tensor{C_{1}} & \tensor{C_{0}} & \tensor{C_{1}} & \tensor{C_{2}} & \tensor{C_{3}} & \tensor{0} & \tensor{C_{3}} & \tensor{C_{2}} \cr
	\tensor{C_{2}} & \tensor{C_{1}} & \tensor{C_{0}} & \tensor{C_{1}} & \tensor{C_{2}} & \tensor{C_{3}} & \tensor{0} & \tensor{C_{3}} \cr
	\tensor{C_{3}} & \tensor{C_{2}} & \tensor{C_{1}} & \tensor{C_{0}} & \tensor{C_{1}} & \tensor{C_{2}} & \tensor{C_{3}} & \tensor{0} \cr
	\tensor{0} & \tensor{C_{3}} & \tensor{C_{2}} & \tensor{C_{1}} & \tensor{C_{0}} & \tensor{C_{1}} & \tensor{C_{2}} & \tensor{C_{3}} \cr
	\tensor{C_{3}} & \tensor{0} & \tensor{C_{3}} & \tensor{C_{2}} & \tensor{C_{1}} & \tensor{C_{0}} & \tensor{C_{1}} & \tensor{C_{2}} \cr
	\tensor{C_{2}} & \tensor{C_{3}} & \tensor{0} & \tensor{C_{3}} & \tensor{C_{2}} & \tensor{C_{1}} & \tensor{C_{0}} & \tensor{C_{1}} \cr
	\tensor{C_{1}} & \tensor{C_{2}} & \tensor{C_{3}} & \tensor{0} & \tensor{C_{3}} & \tensor{C_{2}} & \tensor{C_{1}} & \tensor{C_{0}}
},
\label{eq:circulant_elements}
\end{equation}
where each block $\mathbf{C}_{l}$ is:

\begin{equation}
\tensor{C}_{l} =
\pmatrix{
	\tensor{A}_{l} & \times \cr
	\times & \tensor{A}_{l}
} =
\pmatrix{
	a_{0}^{\ell} & a_{1}^{\ell} & a_{2}^{\ell} & 0^{\ell} & a_{2}^{\ell} & a_{1}^{\ell} \cr
	a_{1}^{\ell} & a_{0}^{\ell} & a_{1}^{\ell} & a_{2}^{\ell} & 0^{\ell} & a_{2}^{\ell} \cr
	a_{2}^{\ell} & a_{1}^{\ell} & a_{0}^{\ell} & a_{1}^{\ell} & a_{2}^{\ell} & 0^{\ell} \cr
	0^{\ell} & a_{2}^{\ell} & a_{1}^{\ell} & a_{0}^{\ell} & a_{1}^{\ell} & a_{2}^{\ell} \cr
	a_{2}^{\ell} & 0^{\ell} & a_{2}^{\ell} & a_{0}^{\ell} & a_{0}^{\ell} & a_{1}^{\ell} \cr
	a_{1}^{\ell} & a_{2}^{\ell} & 0^{\ell} & a_{2}^{\ell} & a_{1}^{\ell} & a_{0}^{\ell}
}.
\label{eq:toeplitz_circulant_element}
\end{equation}

The matrix $\mathbf{C}$ is a $4N_x N_y \times 4N_x N_y$ Block-Circulant matrix formed by Circulant-Blocks matrices. This matrix is formed by a grid of $2N_x \times 2N_x$ blocks, where each block is a $2N_y \times 2N_y$ matrix.

Instead of calculating the product $\mathbf{d} = \mathbf{A} \mathbf{p}$ (equation \ref{eq:predicted-data-vector}), we now carry the following multiplication:

\begin{equation}
\mathbf{C} \mathbf{v} = \mathbf{q} \: ,
\label{eq:BCCB_vector_product}
\end{equation}
where $\mathbf{v}$ and $\mathbf{q}$ are $4N_x N_y \times 1$ vectors given, respectively, by:

\begin{equation}
\mathbf{v} =
\pmatrix{
	\mathbf{v}_{0} \cr
	\mathbf{v}_{1} \cr
	\vdots \cr
	\mathbf{v}_{N_x- 1} \cr
	\mathbf{0}_{(2 N_x N_y)}
}
\end{equation}

\noindent and

\begin{equation}
\mathbf{q} =
\pmatrix{
	\mathbf{q}_{0} \cr
	\mathbf{q}_{1} \cr
	\vdots \cr
	\mathbf{q}_{N_x - 1} \cr
	\mathbf{\dagger}_{(2 N_x N_y)}
} \: ,
\end{equation}
where $\mathbf{0}_{(2 N_x N_y)}$ is a $2 N_x N_y \times 1$ vector of zeros, $\mathbf{\dagger}_{(2 N_x N_y)}$ is $2 N_x N_y \times 1$ vector of ``garbage" (ACHAR UM JEITO MELHOR DE DESCREVER, CHAN E ZHANG DESCREVEM APENAS COMO ``A PARTE A SER JOGADA FORA") and $\mathbf{v}_{\ell}$ and $\mathbf{q}_{\ell}$, $\ell = 0, \dots, N_x-1$ are:

\begin{equation}
\mathbf{v}_{\ell} =
\pmatrix{
	\mathbf{p}_{\ell} \cr
	\mathbf{0}_{(N_y)}
}
\end{equation}

\noindent and

\begin{equation}
\mathbf{q}_{\ell} =
\pmatrix{
	\mathbf{d}_{\ell} \cr
	\mathbf{\dagger}_{(N_y)}
} \: ,
\end{equation}
where $\mathbf{0}_{(N_y)}$ is a $N_y \times 1$ vector of zeros, $\mathbf{\dagger}_{(N_y)}$ is a $N_y \times 1$ vector of ``garbage" (ACHAR UM JEITO MELHOR DE DESCREVER, CHAN E ZHANG DESCREVEM APENAS COMO ``A PARTE A SER JOGADA FORA")  and $\mathbf{d}_{\ell}$ and $\mathbf{p}_{\ell}$ are $N_y \times 1$ vectors partitioned from the original $\mathbf{d}$ and $\mathbf{p}$ vectors (equation \ref{eq:predicted-data-vector}).

As demonstrated by \cite{grenander1984szeg}, circulant matrices can be diagonalized by taking the discrete Fourier transform (DFT), i.e., its eigenvalues can be easily calculated by a fast algorithm of DFT. For BCCB matrices \cite{chan2007introduction} demonstrate that they satisfy:

\begin{equation}
\mathbf{C} = \left(\mathbf{F}_{(2N_y)} \otimes \mathbf{F}_{(2N_x)} \right)^{\ast} \mathbf{\Lambda}\left( \mathbf{F}_{(2N_y)} \otimes \mathbf{F}_{(2N_x)} \right) \: ,
\label{eq:fft_BCCB}
\end{equation}

\noindent where $\mathbf{F}_{(2N_y)}$ and $\mathbf{F}_{(2N_x)}$ are the matrices of the discrete Fourier transform, ``$\otimes$'' represents the Kronecker product, ``$\ast$'' the conjugate matrix and $\mathbf{\Lambda}$ is the diagonal matrix $4N_xN_y \times 4N_xN_y$ containing the eigenvalues of $\mathbf{C}$. By proper manipulating equation \ref{eq:fft_BCCB} it is possible to calculate the eigenvalues of the BCCB matrix using only its first column:

\begin{eqnarray}
\left( \mathbf{F}_{(2N_y)} \otimes \mathbf{F}_{(2N_x)} \right) \mathbf{C} = \mathbf{\Lambda} \, \left( \mathbf{F}_{(2N_y)} \otimes \mathbf{F}_{(2N_x)} \right) \\
\left( \mathbf{F}_{(2N_y)} \otimes \mathbf{F}_{(2N_x)} \right) \, \mathbf{C} \, \mathbf{t}_{1} = \mathbf{\Lambda} \, \left( \mathbf{F}_{(2N_y)} \otimes \mathbf{F}_{(2N_x)} \right) \, \mathbf{t}_{1} \\
\left( \mathbf{F}_{(2N_y)} \otimes \mathbf{F}_{(2N_x)} \right) \, \mathbf{c}_{0} = \mathbf{\Lambda} \, \frac{1}{\sqrt{4N_xN_y}}\mathbf{1}_{(4N_xN_y)} \\
\left( \mathbf{F}_{(2N_y)} \otimes \mathbf{F}_{(2N_x)} \right) \, \mathbf{c}_{0} = \frac{1}{\sqrt{4N_xN_y}} \mathbf{\lambda} \quad ,
\label{eq:fft_BCCB_column}
\end{eqnarray}

\noindent where $\mathbf{t}_{1}$ is a $4N_xN_y \times 1$ vector with first element equal to 1 and all the remaining elements equal to 0, the vector $\mathbf{c}_{0}$ $4N_xN_y \times 1$ is the first column of $\mathbf{C}$, $\mathbf{1}_{(4N_xN_y)}$ is a $4N_xN_y \times 1$ with all elements equal to 1 and $\mathbf{\lambda}$ is a vector representing the diagonal of $\mathbf{\Lambda}$ (the eigenvalues of $\mathbf{C}$). Note that using one of the Kronecker product properties equation \ref{eq:fft_BCCB_column} can be calculated using the 2D-DFT \cite[]{jain1989fundamentals}:

\begin{eqnarray}
\mathbf{F}_{(2N_x)} \mathbf{G} \, \mathbf{F}_{(2N_y)} &= \frac{1}{\sqrt{4N_xN_y}} \mathbf{L}
\quad ,
\label{eq:fft_BCCB_column_2D}
\end{eqnarray}

\noindent where $\mathbf{G}$ is a $2N_x \times 2N_y$ row-oriented matrix containing the elements of the first column $\mathbf{c}_{0}$ of $\mathbf{C}$ and $\mathbf{L}$ is a $2N_x \times 2N_y$ row-oriented matrix containing the eigenvalues $\mathbf{\lambda}$.

By substituting equation \ref{eq:fft_BCCB} in equation \ref{eq:BCCB_vector_product}, using the fast calculation of the eigenvalues of BCCB matrices (equation \ref{eq:fft_BCCB_column_2D}) and using the same previous Kronecker product property, the auxiliary matrix-vetor product $\mathbf{C} \mathbf{v} = \mathbf{q}$ can be rewritten as follows:

\begin{equation}
\mathbf{F}_{(2 N_x)}^{\ast} \left[ \mathbf{L} \circ \left( \mathbf{F}_{(2N_x)} \mathbf{V} \, \mathbf{F}_{(2N_y)} \right) \right] \mathbf{F}_{(2N_y)}^{\ast} = \mathbf{Q} \: ,
\label{eq:fft_q}
\end{equation}

\noindent where ``$\circ$'' denotes the Hadamard product, $\mathbf{L}$ is a $2N_x \times 2N_y$ row-oriented matrix containing the eigenvalues of $\mathbf{C}$, and $\mathbf{V}$ and $\mathbf{Q}$ are $2N_x \times 2N_y$ row-oriented matrices obtained from the vectors $\mathbf{v}$ and $\mathbf{q}$.

Note that in general, the first column of blocks forming a BCCB matrix $\mathbf{C}_{\ell}$ (equation \ref{eq:circulant_elements}) is given by:

\begin{equation}
\left[\mathbf{C} \right]_{(0)} = 
\pmatrix{
\mathbf{C}_{0} \cr
\mathbf{C}_{1} \cr
\vdots \cr
\mathbf{C}_{N_x-2} \cr
\mathbf{C}_{N_x-1} \cr
\mathbf{0} \cr
\mathbf{C}_{N_x-1} \cr
\mathbf{C}_{N_x-2} \cr
\vdots \cr
\mathbf{C}_{1}
} \; ,
\end{equation}
where each block $\mathbf{C}_{\ell}$, $\ell = 0, \dots, N_x-1$, is a $2N_y \times 2N_y$ circulant matrix and $\mathbf{0}$ is a $2N_y \times 2N_y$ matrix with all elements equal to zero. Thus, the first column $\mathbf{c}_{0}$ of a circulant matrix $\mathbf{C}$ is given by:

\begin{equation}
\mathbf{c}_{0} =
\pmatrix{
	a_{00} \cr
	a_{10} \cr
	\vdots \cr
	a_{(N_y-2)0} \cr
	a_{N_y-1) 0} \cr
	0 \cr
	a_{N_y-1) 0} \cr
	a_{(N_y-2)0} \cr
	\vdots \cr
	a_{10}
} \:.
\end{equation}

To complete the process, after calculating the inverse to obtain $\mathbf{Q}$ it is necessary to rearrange its rows to obtain the vector $\mathbf{q}$ and also rearrange the elements of $\mathbf{q}$ to obtain the wanted vector $\mathbf{d(p)}$.

%The algorithm for computing the matrix-vector product $\mathbf{A}\mathbf{p}$ and obtaining the resulting vector $\mathbf{d(p)}$ can be summarized in the following steps:

%1. Compute the first column of the embedded $4N_x N_y \times 4N_x N_y$ BCCB matrix $\mathbf{C}$ by using the first column of the $N_x N_y \times N_x N_y$ BTTB matrix $\mathbf{A}$;

%2. Compute the eigenvalues of $\mathbf{C}$ rearranging them in rows of the $2N_x \times 2N_y$ matrix $\mathbf{L}$;

%3. Rearrange the vector $\mathbf{v}$ in the rows of the $2N_x \times 2N_y$ matrix $\mathbf{V}$;

%4. Compute the Hadammard product $\mathbf{L} \circ \tilde{\mathbf{P}}$, where $\tilde{\mathbf{P}} = \mathbf{F}_{(2N_x)} \mathbf{P} \, \mathbf{F}_{(2N_y)}$ is computed by using a fast algorithm for 2D DFT;

%5. Compute the resulting $2N_x \times 2N_y$ matrix $\mathbf{Q} = \mathbf{F}_{(2N_x)}^{\ast} \left[ \mathbf{L} \circ \tilde{\mathbf{P}}\right] \mathbf{F}_{(2N_y)}^{\ast}$ by using a fast algorithm for inverse 2D DFT;

%6. Rearrange the rows of $\mathbf{Q}$ to obtain the vector $\mathbf{q}$;

%7. Rearrange the elements of $\mathbf{q}$ to obtain the vector $\mathbf{d(p)}$.

\subsection{Computational performance}
In a normal procedure of the fast equivalent layer proposed by \cite{siqueira2017fast}, at each iteration a full matrix $\tensor{A}$ (equation \ref{eq:predicted-data-vector}) is multiplied by the estimated mass distribution parameter vector $\hat{\mathbf{p}}^k$ producing the predicted gravity data $\mathbf{d(p)}$  iteratively . As pointed in \cite{siqueira2017fast} the number of flops (floating-point operations) necessary to estimate the $N$-dimensional parameter vector inside the iteration loop is:

\begin{equation}
f_0 = N^{it} (3N + 2N^2) \; ,
\label{eq:float_classic}
\end{equation}

\noindent where $N^{it}$ is the number of iterations. From equation \ref{eq:float_classic} it is clear that the matrix-vector product ($2N^2$) accounts for most of the computational complexity in this method.

It is well known that FFT takes $N \log_2(N)$ flops \cite[]{brigham1988fast}. Computing the eigenvalues of the BCCB matrix ($4N \times 4N$) and applying 2D-FFT on the parameter vector (equation \ref{eq:fft_q}), takes $4N \log(4N)$ each. The point-multiplication takes $4N$. As it is necessary to compute the inverse FFT another two $4N \log(4N)$ must be taken in account. However, the sensibility matrix does not change during the process, thus, the eigenvalues of BCCB must be calculated only once, outside of the iteration. This lead us to a flops count in our method of:

\begin{equation}
f_1 = 4N\log(4N) + N^{it} (7N + 8N\log(4N)).
\label{eq:float_bccb}
\end{equation}

Another major improvement of this methodology is the exoneration of calculating the full sensibility matrix $\tensor{A}$ (equation \ref{eq:predicted-data-vector}). Each element needs $12$ flops (equation \ref{eq:aij}), totalizing $12N^2$ flops for the full matrix. Calculating only the first row of the BTTB matrix $12N$ flops is required.
Thus, the full flops count of Siqueira et al.'s (2017) method:

\begin{equation}
f_s = 12N^2 + N^{it} (3N + 2N^2),
\label{eq:float_siqueira}
\end{equation}

it is decreased in our method to:

\begin{equation}
f_f = 12N + 4N \log(4N) + N^{it} (7N + 8N\log(4N)).
\label{eq:float_new}
\end{equation}

Figure \ref{fig:float} shows the floating points to estimate the parameter vector using the fast equivalent layer with Siqueira et al.'s (2017) method (equation \ref{eq:float_classic}) and our approach (equation \ref{eq:float_bccb}) versus the number of observation points varyig from $N = 5000$ to $N = 1000000$ with $50$ iterations. The number of operations is drastically decreased.

Table 1 shows the system memory RAM usage needed to store the full sensibility matrix, the BTTB first row and the BCCB eigenvalues ($8$ times the BTTB first row). The quantities were computed for different numbers of data (N) with the same corresponding number of equivalent sources (N). Table 1 considers that each element of the matrix is a double-precision number, which requires 8 bytes of storage, except for the BCCB complex eigenvalues, which requires 16 bytes per element. Notice that $1,000.000$ observation points requires nearly $7.6$ Terabytes of memory RAM to store the whole sensibility matrix of the equivalent layer.

Using a PC with a Intel Core i7 4790@3.6GHz processor and 16 Gb of memory RAM, Figure \ref{fig:time_comparison}  compares the running time  of the Siqueira et al.'s (2017) method with the one of our work, considering a constant number of iterations equal to 50. Clearly, the major advantage of our approach is its computational efficiency that allows a rapid calculation of the gravity forward modeling  with number of observations greater than $10,000$. Because of the RAM available in this system, we could not perform this comparison with more observations. Therefore, the number of observation is limited to $22,500$. Disregarding the limitation of 16 Gb of RAM, Figure \ref{fig:time_bccb} shows the running time of our method with 50 iterations and with the number of observations up to  25 millions. Our method requires $26:8$ seconds to run one million of observations, whereas  Siqueira et al.'s (2017) method took $48:3$ seconds to run $22,500$ observations

\section{Synthetic tests}
In this section, we  investigate the effectiveness of using the properties of BTTB and BCCB matrices (equation \ref{eq:BCCB_vector_product})  to solve, at each iteration, the forward modeling (the matrix-vector product $\mathbf{A} \hat{\mathbf{p}}^k$)  required in the fast equivalent layer proposed by Siqueira et al. (2017).  We simulated three sources whose horizontal projections are shown in Figure \ref{fig:synthetic_data} as black lines.  These sources are two vertical prisms with density contrasts of $0.35 g/cm^3$ (upper-left prism) and $0.4 g/cm^3$ (upper-right prism) and a sphere with radius of 1,000 m with density contrast of $-0.5 g/cm^3$. Figure \ref{fig:synthetic_data}  shows the vertical component of gravity field generated by these sources contaminated with additive pseudorandom Gaussian noise with zero mean and standard deviation of $0.01486$ mGal.

The advantage of using the structures of BTTB and BCCB matrices to compute forward modeling in  the fast equivalent layer proposed by Siqueira et al. (2017) is   grounded  on the use of regular grids of data and equivalent sources. Hence, we created 10,000 observation points regularly spaced in a grid of $100 \times 100$ at $100$ m height. We also set a grid of equivalent point masses, each one directly beneath each observation points, located at $300$ m deep.  Figures \ref{fig:classic_fast_val}a and \ref{fig:bccb_fast_val}a  show the fitted gravity data obtained, respectively, by the fast equivalent layer using Siqueira et al.'s (2017) method and by our modified form of this method that computes the forward modeling using equation \ref{eq:BCCB_vector_product}. The corresponding residuals (Figures \ref{fig:classic_fast_val}b and \ref{fig:bccb_fast_val}b), defined as the difference between the observed (Figure  \ref{fig:synthetic_data}) and fitted gravity data (Figures \ref{fig:classic_fast_val}a and \ref{fig:bccb_fast_val}a), show means close to zero  and standard deviations of 0.0144 mGal.  Therefore, Figures \ref{fig:classic_fast_val} and \ref{fig:bccb_fast_val} show that Siqueira et al.'s (2017) method and our modified version of this method produced virtually the same results. This excellent agreement is confirmed in Figures \ref{fig:delta_comparison} and \ref{fig:delta_rho} which shows that there are virtually no differences, respectively,  in the fitted data presented in Figures \ref{fig:classic_fast_val}b and \ref{fig:bccb_fast_val}b and in the estimated mass distributions within the equivalent layers (not shown) yielded by both Siqueira et al.'s (2017) method and our modification of this method.  These results (\ref{fig:delta_comparison} and \ref{fig:delta_rho})  show that computing the matrix-vector product ($\mathbf{A} \hat{\mathbf{p}}^k$), required in the forward modeling, by means of embedding the BTTB matrix into a BCCB matrix (equation \ref{eq:BCCB_vector_product}) yields practically the same result as the one produced by computing this matrix-vector product  with a full matrix $\mathbf{A}$.

We perform two forms of processing the gravity data (Figure \ref{fig:synthetic_data}) through the equivalent layer technique: the upward (Figure \ref{fig:upward_comparison}) and the downward (Figure \ref{fig:downward_comparison}) continuations. The upward height is 300m and the downward  is at 50m.   Either in the upward continuation (Figure \ref{fig:upward_comparison}) or in the downward continuation (Figure \ref{fig:downward_comparison}), the continued gravity data using the fast equivalent layer proposed by \cite{siqueira2017fast} (Figures \ref{fig:upward_comparison}a and \ref{fig:downward_comparison}a) are in close agreement with those produced by our modification of Siqueira et al.'s (2017) method (Figures \ref{fig:upward_comparison}b and \ref{fig:downward_comparison}b). The residuals (Figures \ref{fig:upward_comparison}c and \ref{fig:downward_comparison}c) quantify this agreement since their means and standard deviations are close to zero in both continued gravity data using both methods.  All the continued gravity data shown here (Figures \ref{fig:upward_comparison} and \ref{fig:downward_comparison}) agree with the true ones (not shown). The most striking feature of these upward or the downward continuations concerns the total computation time. The computation time spent by our method is approximately $1500$ times faster than Siqueira et al.'s (2017) method.

\section{Real data test}
Test with real data are conducted with the gravity data from Caraj\'as, north of Brazil, were provided by the  Geological Survey of Brazil (CPRM). The real aerogravimetric data were collected in 113 flight lines along north–south direction with flight line spacing  of 3 km and tie lines along east–west direction at 12 km.

This airborne gravity survey was divided in two different areas, collected in different times, having samples spacing of 7.65 m and 15.21 m, totalizing  4,353,428 observation points. The height of the flight was fixed at 900 m. The gravity data (Figure \ref{fig:carajas_real_data}) were gridded into a regularly spaced dataset of $250,000$ observation points ($500 \times 500$) with a grid spacing of $716.9311$ km north-south and $781.7387$ km east-west.

To apply our modification of the fast equivalent layer method (\cite{siqueira2017fast})  that computes the forward modeling using the properties of BTTB and BCCB matrices (equation \ref{eq:BCCB_vector_product}), we set an equivalent layer at 300 m deep. Figure \ref{fig:carajas_gz_predito_val}a shows the fitted gravity data after with 50 iterations by applying our method approach. The residuals (Figure \ref{fig:carajas_gz_predito_val}b), defined as the difference between the observed (Figure \ref{fig:carajas_real_data}) and the predicted (Figure \ref{fig:carajas_gz_predito_val}a) data,  show an acceptable data fitting because they have a mean close to zero (0.000292 mGal) and a small standard deviation of 0.105 mGal which corresponds to approximately 0.1 \% of the amplitude of the gravity data.

These small residuals indicate that our method yielded an estimated mass distribution (not shown) that can be used in the data processing. We perform upward-continuation of the real gravity data (Figure \ref{fig:carajas_real_data}) at a constant height of 5000 m over the real data. The upward-continued gravity data (Figure \ref{fig:up2000_carajas_500x500}) seem a reasonable processing because of the attenuatinon of the short-wave lenghts. By using our approach, the processing of the 250,000 observations was extremely fast and took 0.216 seconds.

\section{Conclusions}
By exploring the properties related to Block-Toeplitz Toeplitz-block (BTTB) and Block-Circulant Circulant-Block (BCCB) matrices in the gravity data processing, we have proposed a new efficient approach for calculating the gravity-data forward modeling required in the iterative fast equivalent-layer technique grounded on excess mass constraint that does not demand  the solution of linear systems.  Its efficiency is grounded on the use of regular grids of observations and equivalent sources (point masses). Our algorithm greatly reduces the number of  flops necessary to estimate a 2D mass distribution within the equivalent layer that fits the observed gravity data. For example, when processing one million observations the number of flops is reduced in 104 times. Traditionally, such amount of data impractically requires 7.6 Terabytes of RAM memory to handle the full sensibility matrix. Rather, in our method, this matrix takes 61,035 Megabytes of RAM memory only.

Our method takes advantage of the symmetric BTTB system that arises when processing a harmonic function as the vertical component of gravity, that depends on the inverse
of distance between the observation and the point mass over the equivalent layer. Symmetric BTTB matrices can be stored by its only first row and can be embedded into a symmetric BCCB matrix, which in turn also only needs its first row.

Using the fast Fourier transform it is possible to calculate the eigenvalues of BCCB matrices which can
be used to compute a matrix-vector product (gravity-data forward modeling) in a very low computational cost. The time
needed to process medium-sized grids of observation, for example 22,500 points, is reduced in 102 times. We have successfully applied the proposed method to upward (or downward) synthetic gravity data. Testing on field data from the Caraj\'as Province, north of Brazil, confirms the potential of our approach in upward-continuing  gravity data with 250,000 observations in  about 0.2 seconds.  Our method allows, in future research, applying the equivalent layer technique for processing and interpreting massive data set such as collected in continental and global scales studies.

\newpage

\subsection*{Figures}
\renewcommand{\figdir}{Fig} % figure directory

Figure~\ref{fig:float}

\plot{float}{width=\textwidth}
{floating points to estimate the parameter vector using the fast equivalent layer with Siqueira et al.'s (2017) method (equation \ref{eq:float_classic}) and our approach (equation \ref{eq:float_bccb}) versus the numbers of observation points varyig from $N = 5000$ to $N = 1000000$ with $50$ iterations. The number of operations is drastically decreased.}
\newpage

Figure~\ref{fig:time_comparison}

\plot{time_comparison}{width=10cm}
{time necessary to run 50 iterations of the Siqueira et al.'s (2017) method and the one presented in this work. With the limitation of $16$ Gb of memory RAM in our system, we could test only up to $22500$ obervation points.}

Figure~\ref{fig:time_bccb}

\plot{time_bccb}{width=10cm}
{time necessary to run the equivalent layer technique with 50 iterations using only this new approach, where the RAM is not a limitation factor. We could run up to $25$ million observation points. In comparison, $1$ million observation points took $26.8$ seconds to run, where the maximun $22500$ observation points in figure \ref{fig:time_comparison}, with Siqueira et al.'s (2017) method, took $48.3$ seconds.}
\newpage

Figure~\ref{fig:synthetic_data}

\plot{synthetic_data}{width=13.5cm}
{model with two polygonal prisms, with density contrast of $0.35$ (upper-left body) and $0.4 g/cm^3$ (upper-right body), and a sphere with radius of $1000 m$ with density contrast of $-0.5 g/cm^3$. The vertical component of gravity generated by this bodies were calculated and are shown together with their horizontal projections. A gaussian noise was added to the data with mean of zero and maximum value of $0.5\%$ of the maximum of the original data. As previous said only in regular grids the BTTB matrix structures appears. We created $10000$ observation points regularly spaced in a grid of $100 \times 100$, with a uniform $100 m$ of height for all the observations.}
\newpage

Figure~\ref{fig:classic_fast_val}

\plot{classic_fast_val}{width=8cm}
{(a) Fitted gravity data produced by the fast equivalent layer proposed by \cite{siqueira2017fast}. (b) Gravity residuals, defined as the difference between the observed data in Figure \ref{fig:synthetic_data} and the predicted data in (a), with their mean of $8.264e-7$ and standard deviation of $0.0144$ mGal.}
\newpage

Figure~\ref{fig:bccb_fast_val}

\plot{bccb_fast_val}{width=8cm}
{(a) Fitted gravity data produced by our modification of the fast equivalent layer proposed by \cite{siqueira2017fast}. (b) Gravity residuals, defined as the difference between the observed data in Figure 4 and the predicted data in (a), with their mean of $8.264e-7$ and standard deviation of $0.0144$ mGal.}
\newpage

Figure~\ref{fig:delta_comparison}

\plot{delta_comparison}{width=13.5cm}
{Difference between the fitted gravity data produced by Siqueira et al.'s (2017) method (Figure \ref{fig:classic_fast_val}a) and by our modified form of this method that computes the forward modeling using the properties of BTTB and BCCB matrices (equation \ref{eq:BCCB_vector_product}).}
\newpage

Figure~\ref{fig:delta_rho}

\plot{delta_rho}{width=13.5cm}
{Difference between the estimated mass distribution within the equivalent layer produced by Siqueira et al.'s (2017) method (Figure \ref{fig:classic_fast_val}a) and by our modified form of this method that computes the forward modeling using the properties of BTTB and BCCB matrices (equation \ref{eq:BCCB_vector_product}).}
\newpage

Figure~\ref{fig:upward_comparison}

\plot{upward_comparison}{width=5cm}
{The upward-continued gravity data using: (a) the fast equivalent layer proposed by \cite{siqueira2017fast} and (b) our modified form of Siqueira et al.'s (2017) method by using the properties of BTTB and BCCB matrices (equation \ref{eq:BCCB_vector_product}) to calculate the forward modeling.  (c) Residuals, defined as the difference between a and b with their mean of $-5.938e-18$ and standard deviation of $8.701e-18$.  The total computation times in the Siqueira et al.'s (2017) method and in our approach are $7.62026$ and $0.00834$ seconds, respectively.}
\newpage

Figure~\ref{fig:downward_comparison}

\plot{downward_comparison}{width=5cm}
{The downward-continued gravity data using: (a) the fast equivalent layer proposed by \cite{siqueira2017fast} and (b) our modified form of Siqueira et al.'s (2017) method by using the properties of BTTB and BCCB matrices (equation \ref{eq:BCCB_vector_product}) to calculate the forward modeling.  (c) Residuals, defined as the difference between a and b with their mean of $5.914e-18$ and standard deviation of $9.014e-18$.  The total computation times in the Siqueira et al.'s (2017) method and in our approach are $7.59654$ and $0.00547$ seconds, respectively.}
\newpage

Figure~\ref{fig:carajas_real_data}

\plot{carajas_real_data}{width=13.5cm}
{Caraj\'as Province, Brazil. Gravity data on a regular grid of $500 \times 500$ points, totaling $250,000$ observations. The inset shows the study area (blue rectangle) which covers the southeast part of the state of Par\'a, north of Brazil.}
\newpage

Figure~\ref{fig:carajas_gz_predito_val}

\plot{carajas_gz_predito_val}{width=8cm}
{Caraj\'as Province, Brazil. (a) Predicted gravity data produced by our modification of the fast equivalent layer method (\cite{siqueira2017fast})  that computes the forward modeling using the properties of BTTB and BCCB matrices (equation \ref{eq:BCCB_vector_product}). (b) Gravity residuals, defined as the difference between the observed data in Figure \ref{fig:carajas_real_data} and the predicted data in a, with their mean of 0.000292 mGal and standard deviation of $0.105$ mGal.}
\newpage

Figure~\ref{fig:up2000_carajas_500x500}

\plot{up2000_carajas_500x500}{width=13.5cm}
{Caraj\'as Province, Brazil. The upward-continued gravity data using our modification of the fast equivalent layer method (\cite{siqueira2017fast}) that computes the forward modeling using the properties of BTTB and BCCB matrices (equation \ref{eq:BCCB_vector_product}). The total computation time is 0.216 seconds for processing of the $250,000$ observations.}
\newpage

%\subsubsection{Multiplot} 
%Sometimes it is convenient to put two or more figures from different
%files in an array (see Figure~\ref{fig:exph,exgr}). Individual plots
%are Figures~\ref{fig:exph} and~\ref{fig:exgr}.
%
%\multiplot{2}{exph,exgr}{width=0.4\textwidth}
%{This figure is specified in the document by \newline \texttt{
%    $\backslash$multiplot\{2\}\{exph,exgr\}\{width=0.4$\backslash$textwidth\}\{This caption.\}}.
%}
%
%The first argument of the \texttt{multiplot} command specifies the
%number of plots per row.

\subsection{Tables}

\begin{table}[h]
	\begin{center}
		\begin{tabular}{|l|c|c|c|}
			\hline
			\textbf{$N \times N$} & \textbf{Full RAM (Mb)} & \textbf{BTTB RAM (Mb)}  & \textbf{BCCB RAM (Mb)}\\
			\hline 
			$100 \times 100$ & 0.0763 & 0.0000763 & 0.0006104\\
			\hline
			$400 \times 400$ & 1.22 & 0.0031 & 0.0248\\
			\hline
			$2500 \times 2500$ & 48 & 0.0191 & 0.1528\\
			\hline
			$10000 \times 10000$ & 763 & 0.00763 & 0.6104\\
			\hline
			$40000 \times 40000$ & 12207 & 0.305 & 2.4416 \\
			\hline
			$250000 \times 250000$ & 476837 & 1.907 & 15.3 \\
			\hline
			$500000 \times 500000$ & 1907349 & 3.815 & 30.518 \\
			\hline
			$1000000 \times 1000000$ & 7629395 & 7.629 & 61.035 \\
			\hline
		\end{tabular}
		\caption{Comparison between the system memory RAM usage needed to store the full matrix, the BTTB first row and the BCCB eigenvalues (eight times the BTTB). The quantities were computed for different numbers of data (N) with the same corresponding number of equivalent sources (N). This table considers that each element of the matrix is a double-precision number, which requires 8 bytes of storage, except for the BCCB complex eigenvalues, which requires 16 bytes per element.}
	\end{center}
\end{table}

%The discussion is summarized in Table~\ref{tbl:example}.
%
%\tabl{example}{This table is specified in the document by \texttt{
%    $\backslash$tabl\{example\}\{This caption.\}\{\ldots\}}.
%}{
%  \begin{center}
%    \begin{tabular}{|c|c|c|}
%      \hline
%      \multicolumn{3}{|c|}{Table Example} \\
%      \hline
%      migration\rule[-2ex]{0ex}{5ex} & 
%      $\omega \rightarrow k_z$ & 
%      $k_y^2+k-z^2\cos^2 \psi=4\omega^2/v^2$ \\
%      \hline
%      \parbox{1in}{zero-offset\\diffraction}\rule[-4ex]{0ex}{8ex} &
%      $k_z\rightarrow\omega_0$ &
%      $k_y^2+k_z^2=4\omega_0^2/v^2$ \\
%      \hline
%      DMO+NMO\rule[-2ex]{0in}{5ex} & $\omega\rightarrow\omega_0$ & 
%      $\frac{1}{4}
%      v^2k_y^2\sin^2\psi+\omega_0^2\cos^2\psi=\omega^2$ \\
%      \hline
%      radial DMO\rule[-2ex]{0in}{5ex} & $\omega\rightarrow\omega_s$ &
%      $\frac{1}{4}v^2k_y^2\sin^2\psi+\omega_s^2=\omega^2$\\
%      \hline
%      radial NMO\rule[-2ex]{0in}{5ex} & $\omega_s\rightarrow\omega_0$ &
%      $\omega_0\cos\psi=\omega_s$\\
%      \hline
%    \end{tabular}
%  \end{center}
%}

\section{ACKNOWLEDGMENTS}

This study was financed by the brazilian agencies CAPES (in the form of a scholarship), FAPERJ (grant n.$^{\circ}$ E-26 202.729/2018) and CNPq (grant n.$^{\circ}$ 308945/2017-4).

%\append{Appendix example}

%\append[equations]{Another appendix}

%\sideplot{errgrp}{width=0.8\textwidth}
%{This figure is specified in the document by \texttt{
%    $\backslash$sideplot\{errgrp\}\{width=0.8$\backslash$text\-width\}\{This caption.\}}.

%\append{The source of this document}

%\verbatiminput{geophysics_example.ltx}

%\append{The source of the bibliography}

%\verbatiminput{example.bib}

\newpage

\bibliographystyle{seg}  % style file is seg.bst
\bibliography{references}

\end{document}

### 2019-11-29 manuscript/conclusions.tex
split the tex files

\section{Conclusions}
By exploring the BTTB structure of the sensitivity matrix in the gravity data processing, we have proposed a new 
efficient approach for calculating the gravity-data forward modeling required in the iterative fast equivalent-layer 
technique grounded on excess mass constraint that does not demand  the solution of linear systems.  Its efficiency 
requires the use of regular grids of observations and equivalent sources (point masses). Our algorithm greatly reduces 
the number of  flops necessary to estimate a 2D mass distribution within the equivalent layer that fits the observed 
gravity data. For example, when processing one million observations the number of flops is reduced in 104 times. 
Traditionally, such amount of data impractically requires 7.6 Terabytes of RAM memory to handle the full sensitivity 
matrix. Rather, in our method, this matrix takes 61.035 Megabytes of RAM memory only.

Our method takes advantage of the symmetric BTTB system that arises when processing a harmonic function and considering 
that either the observations or the sources of the interpretative model (point of masses over the equivalent layer) 
are distributed on regular grids. Symmetric BTTB matrices can be stored by using only a single column and can be 
embedded into a symmetric BCCB matrix, which in turn also only needs a single column. 
This means that only a single column  of the sensitivity matrix needs to be calculated.  Once the jth column of the 
sensitivity matrix represents the influence of the jth source (jth point mass) has on the predicted data, our method 
only requires a single point mass that set up the equivalent layer to compute the gravity-data forward modeling.

Using the 2D FFT, it is possible to calculate the eigenvalues of BCCB matrices which can
be used to compute a matrix-vector product (gravity-data forward modeling) in a very low computational cost. 
We have successfully applied the proposed method to upward (or downward) synthetic gravity data. Testing on field 
data from the Caraj{\'a}s Province, north of Brazil, confirms the potential of our approach in upward-continuing 
gravity data with $250\,000$ observations in  about 0.2 seconds.  Our method allows, in future research, applying 
the equivalent layer-technique for processing and interpreting massive data set such as collected in continental 
and global scales studies.

### 2019-11-29 manuscript/field_data.tex
split the tex files

\section{Field data test}
Test with real data are conducted with the gravity data from Caraj\'as, north of Brazil, were provided by the  Geological 
Survey of Brazil (CPRM). The real aerogravimetric data were collected in 113 flight lines along north–south direction 
with flight line spacing  of 3 km and tie lines along east–west direction at 12 km.

This airborne gravity survey was divided in two different areas, collected in different times, having samples spacing of 
$7.65$ m and $15.21$ m, totalizing  $4\,353\,428$ observation points. The height of the flight was fixed at 900 m. 
The gravity data (Figure \ref{fig:carajas_real_data}) were gridded into a regularly spaced dataset of $250\,000$ 
observation points ($500 \times 500$) with a grid spacing of $716.9311$ km north-south and $781.7387$ km east-west.

To apply our modification of the fast equivalent-layer method \cite[]{siqueira-etal2017}  that computes the forward 
modeling using the properties of BTTB and BCCB matrices (equation \ref{eq:w_Cv}), we set an equivalent layer at 300 m deep. 
Figure \ref{fig:carajas_gz_predito_val}a shows the fitted gravity data after 50 iterations by applying our method. 
The residuals (Figure \ref{fig:carajas_gz_predito_val}b), defined as the difference between the observed 
(Figure \ref{fig:carajas_real_data}) and the predicted (Figure \ref{fig:carajas_gz_predito_val}a) data, show an 
acceptable data fitting because they have a mean close to zero ($0.0003$ mGal) and a small standard deviation of 
$0.105$ mGal which corresponds to approximately $0.1$ \% of the amplitude of the gravity data.

These small residuals indicate that our method yielded an estimated mass distribution (not shown) that can be used in 
the data processing. We perform upward-continuation of the real gravity data (Figure \ref{fig:carajas_real_data}) at 
a constant height of $5\,000$ m over the real data. The upward-continued gravity data 
(Figure \ref{fig:up2000_carajas_500x500}) seem a reasonable processing because of the attenuation of the short 
wavelenghts. By using our approach, the processing of the $250\,000$ observations was extremely fast and took 
$0.216$ seconds.

### 2019-11-29 manuscript/figures.tex
split the tex files

\renewcommand{\figdir}{Fig} % figure directory

\plot{float}{width=\textwidth}{
	{Comparison between the number of flops (equations \ref{eq:float_fast_eqlayer} and \ref{eq:float_bccb}) 
	associated with the fast equivalent-layer technique \citep{siqueira-etal2017} and our method, 
	for $N$ varying from $5\,000$ to $1\,000\,000$. All values are computed with $N^{it} = 50$ iterations.}
	\label{fig:float}
}

\plot{time_comparison}{width=10cm}{
	{Comparison between the running time of the fast equivalent-layer technique \citep{siqueira-etal2017} and our method.
	The values were obtained for $N^{it} = 50$ iterations.}
	\label{fig:time_fast_eqlayer_bccb}
}

\plot{time_bccb}{width=10cm}{
	{Running time of our method for a number of observations $N$ up to $25$ millions. 
	The values were obtained for $N^{it} = 50$ iterations.}
	\label{fig:time_bccb}
}

\plot{synthetic_data}{width=13.5cm}{
	{Noise-corrupted gravity data (in color map) produced by three simulated sources  whose horizontal projections 
	are shown in black lines. The simulated sources are: two polygonal prisms, with density contrast of 
	$0.35\,	\mathrm{g/cm^3}$ (upper-left body) and $0.4\, \mathrm{g/cm^3}$ (upper-right body), and a sphere with 
	radius of $1\,000$ m with density contrast	of $-0.5\, \mathrm{g/cm^3}.$}
	\label{fig:synthetic_data}
}

\plot{classic_fast_val}{width=8cm}{
	{(a) Fitted gravity data produced by the fast equivalent-layer technique proposed by \citet{siqueira-etal2017}. 
	(b) Gravity residuals, defined as the difference between the observed data in Figure \ref{fig:synthetic_data} and 
	the predicted data in panel a, with their mean of $8.264e^{-7}$ and standard deviation of $0.0144$ mGal.}
	\label{fig:classic_fast_val}
}

\plot{bccb_fast_val}{width=8cm}{
	{(a) Fitted gravity data produced by our modification of the fast equivalent layer \citep{siqueira-etal2017}. 
	(b) Gravity residuals, defined as the difference between the observed data in Figure 4 and the predicted data in 
	panel a, with their mean of $8.264e^{-7}$ and standard deviation of $0.0144$ mGal.}
	\label{fig:bccb_fast_val}
}

\plot{delta_comparison}{width=13.5cm}{
	{Difference between the fitted gravity data produced by \citet{siqueira-etal2017} method 
	(Figure \ref{fig:classic_fast_val}a) and by our modified form of this method (Figure \ref{fig:bccb_fast_val}a) 
	that computes the forward modeling using the properties of BTTB and BCCB matrices (equation \ref{eq:w_Cv}).}
	\label{fig:delta_comparison}
}

\plot{delta_rho}{width=13.5cm}{
	{Difference between the estimated mass distribution within the equivalent layer produced by \citet{siqueira-etal2017}
	method (Figure \ref{fig:classic_fast_val}) and by our modified form of this method (Figure \ref{fig:bccb_fast_val}) 
	that computes the forward modeling using the properties of BTTB and BCCB matrices (equation \ref{eq:w_Cv}).}
	\label{fig:delta_rho}
}

\plot{upward_comparison}{width=5cm}{
	{The upward-continued gravity data using: (a) the fast equivalent layer proposed by \citet{siqueira-etal2017} and 
	(b) our modified form of \citet{siqueira-etal2017} method by using the properties of BTTB and BCCB matrices 
	(equation \ref{eq:w_Cv}) to calculate the forward modeling.  (c) Residuals, defined as the difference between 
	panels a and b with their mean of $-5.938e^{-18}$ and standard deviation of $8.701e^{-18}$.  The total computation 
	times in the \citet{siqueira-etal2017} method and in our approach are $7.62026$ and $0.00834$ seconds, respectively.}
	\label{fig:upward_comparison}
}

\plot{downward_comparison}{width=5cm}{
	{The downward-continued gravity data using: (a) the fast equivalent layer proposed by \citet{siqueira-etal2017} and 
	(b) our modified form of \citet{siqueira-etal2017} method by using the properties of BTTB and BCCB matrices 
	(equation \ref{eq:w_Cv}) to calculate the forward modeling.  (c) Residuals, defined as the difference between panels 
	a and b with their mean of $5.914e^{-18}$ and standard deviation of $9.014e^{-18}$.  The total computation times in 
	the \citet{siqueira-etal2017} method and in our approach are $7.59654$ and $0.00547$ seconds, respectively.}
	\label{fig:downward_comparison}
}

\plot{carajas_real_data}{width=13.5cm}{
	{Caraj\'as Province, Brazil. Gravity data on a regular grid of $500 \times 500$ points, totaling $250,000$ observations.
	The inset shows the study area (blue rectangle) which covers the southeast part of the state of Par\'a, north of Brazil.}
	\label{fig:carajas_real_data}
}

\plot{carajas_gz_predito_val}{width=8cm}{
	{Caraj\'as Province, Brazil. (a) Predicted gravity data produced by our modification of the fast equivalent-layer 
	method \citep{siqueira-etal2017}  that computes the forward modeling using the properties of BTTB and BCCB matrices
	(equation \ref{eq:w_Cv}). (b) Gravity residuals, defined as the difference between the observed data in 
	Figure \ref{fig:carajas_real_data} and the predicted data in panel a, with their mean of 0.000292 mGal and standard
	deviation of $0.105$ mGal.}
	\label{fig:carajas_gz_predito_val}
}

\plot{up2000_carajas_500x500}{width=13.5cm}{
	{Caraj\'as Province, Brazil. The upward-continued gravity data using our modification of the fast equivalent layer 
	method \citep{siqueira-etal2017} that computes the forward modeling using the properties of BTTB and BCCB matrices
	(equation \ref{eq:w_Cv}). The total computation time is 0.216 seconds for processing of the $250,000$ observations.}
	\label{fig:up2000_carajas_500x500}
}

### 2019-11-29 manuscript/introduction.tex
split the tex files

\section{Introduction}

The equivalent layer is a well-known technique for processing potential-field data in applied geophysics since 1960. 
It comes from potential theory as a mathematical solution of the Laplace's equation, in the region above the sources, 
by using the Dirichlet boundary condition \citep{kellogg1929}.
This theory states that any potential-field data produced by an arbitrary 3D physical-property distribution can be 
exactly reproduced by a fictitious layer located at any depth and having a continuous 2D physical-property  
distribution. In practical situations, the layer is approximated by a finite set of sources (e.g., point masses or 
dipoles) and their physical properties are estimated by solving a linear system of equations that yield an 
acceptable potential-field data fit. These fictitious sources are called equivalent sources.

Many previous works have used the equivalent layer as a processing technique of potential-field data. 
\citet{dampney1969} used the equivalent-layer technique for gridding and for computing the upward 
continuation of the potential-field data. \citet{cordell1992} and \citet{mendonca-silva1994} used it for 
interpolating and gridding potential-field data. \citet{emilia1973}, \citet{hansen-miyazaki1984} and 
\citet{li-oldenburg2010} used it for upward continuation of the potential-field data. \citet{silva1986}, 
\citet{leao-silva1989}, \citet{guspi-novara2009}, and \citet{oliveirajr-etal2013} used it for reducing the 
magnetic data to the pole. \citet{boggs-dransfield2004} used it for combining multiple data sets and 
\citet{barnes-lumley2011} for gradient-data processing.

The classic equivalent-layer formulation consists in estimating the physical-property distribution within a layer, 
composed by a set of equivalent sources, by solving a linear system of equations formed by harmonic functions 
(e.g.,the inverse of the distance between the observation point and the equivalent source). When these observation 
points and equivalent sources are regularly spaced, a Toeplitz system arises. Toeplitz systems are well-known in 
many branches of science as in (1) mathematics, for solving partial and ordinary diferential equations 
\citep[e.g.,][]{lin-etal2003}; (2) image processing \citep[e.g.,][]{chan-etal1999} and; (3) computational 
neuroscience \citep[e.g.,][]{wray-green1994}. \citet{jin2003} and \citet{chan-jin2007} give many examples of 
applications for Toeplitz systems.

In potential-field methods, the properties of Toeplitz system were used for downward continuation 
\citep{zhang-etal2016} and for 3D gravity-data inversion using a 2D multilayer model \citep{zhang-wong2015}. 
In the particular case of gravity data, the kernel generates a linear system with a matrix known as symmetric 
Block-Toeplitz Toeplitz-Block (BTTB).

INCLUIR \citet{qiang_etal2019}

A wide variety of applications in mathematics and engineers that fall into Toeplitz systems propelled the development 
of a large variety of  methods for solving them. 
%Direct methods were conceived by \citet{levinson1946} and by \citet{trench1964}. Currently the conjugate gradient 
%is used in most cases.
\citet{grenander-szego1984} noticed that a circulant matrix can be diagonalized by taking the Fast Fourier Trasform 
(FFT) of its first column, making it possible to calculate the matrix-vector product and solve the system with 
low computational cost \citep{strang-aarikka1986, olkin1986}. \citet{chan-jin2007} show some preconditioners to embedd 
the Toeplitz and BTTB matrices into, respectively, circulant matrices and Block-Circulant Circulant-Block (BCCB) by 
solving the system applying the conjugate gradient method.

Although the use of the equivalent-layer technique increased over the last decades, one of the biggest problem is still 
its high computational cost for processing large-data sets. \citet{siqueira-etal2017} developed a computationally 
efficient scheme for processing gravity data. This scheme does not solve a linear system, instead uses an iterative 
process that corrects the physical-property distribution over the equivalent layer by adding mass corrections that 
are proportional to the gravity residual. Although efficient, the method presented by \citet{siqueira-etal2017} requires, 
at each iteration,  the full computation of the forward problem to guarantee the convergence of the algorithm. The 
time spent on forward modeling accounts for most of the total computation time of their method.

We propose the use of BTTB and BCCB matrices properties to efficiently  solve the forward modeling in method of 
\citet{siqueira-etal2017}, resulting in a very faster parameter estimation and the possibility to use very large 
datasets. Here, we show how the system memory (RAM) usage can be drastically decreased by calculating only the 
first column of the BTTB matrix and embedding into a BCCB matrix. Using the Szeg\"{o} theorema  combined with 
\cite{strang-aarikka1986}, the matrix-vector product can be accomplished with very low cost, reducing in some 
orders of magnitude the number of operations required to complete the process. We present synthetic tests to validate 
our proposal and real field data from Caraj\'as, Brazil to demonstrate its applicability.

### 2019-11-29 manuscript/manuscript.tex
split the tex files

%\documentclass[paper,twocolumn,twoside]{geophysics}
\documentclass[manuscript,blind]{geophysics}
%\documentclass[manuscript]{geophysics}

% An example of defining macros
\newcommand{\rs}[1]{\mathstrut\mbox{\scriptsize\rm #1}}
\newcommand{\rr}[1]{\mbox{\rm #1}}

% Extra packages
\usepackage{amsmath}
\usepackage{bm}
\usepackage[pdftex,colorlinks=true]{hyperref}
\hypersetup{
	allcolors=black,
}
\usepackage{lipsum}
\usepackage[table]{xcolor}

\begin{document}

\title{Fast equivalent-layer technique for gravity data processing with BTTB matrix systems}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\ms{GEO-XXXX} % manuscript number

%\address{
%\footnotemark[1]Observat\'{o}rio Nacional, \\
%Department of Geophysics, \\
%Rio de Janeiro, Brazil}
%\author{Diego Takahashi\footnotemark[1], Vanderlei C. Oliveira{ }Jr.\footnotemark[1] and 
%Val{\'e}ria C. F. Barbosa\footnotemark[1]}
%
%\footer{vanderlei@on.br}
%\lefthead{Takahashi, Oliveira{ }Jr. \& Barbosa}
%\righthead{Fast equivalent-layer for BTTB systems}

\maketitle

% Main body
\include{abstract}
\include{introduction}
\include{methodology}
\include{synthetic_data}
\include{field_data}
\include{conclusions}
\include{aknowledgements}

% Tables and figures
\include{tables}
\include{figures}

% Appendices
\include{DFT_computations}
\include{eigenvalues_C}

\bibliographystyle{seg}  % style file is seg.bst
\bibliography{references}

\end{document}

### 2019-11-29 manuscript/methodology.tex
split the tex files

\section{Methodology}

\subsection{Equivalent-layer technique for gravity data processing}

Let $d^{o}_{i}$ be the observed gravity data at
the point $(x_{i}, y_{i}, z_{i})$, $i = 1, ..., N$, of a local Cartesian
system with $x$ axis pointing to north, the $y$ axis pointing to east and 
the $z$ axis pointing downward.
Let us consider an equivalent layer compose by a set of $N$ point masses 
(equivalent sources) over a layer located at depth $z_0$ ($z_0 >z_i$) and whose 
$x$- and $y$- coordinates of each point mass coincide with the corresponding coordinates 
of the observation directly above.
There is a linear relationship that maps the unknown mass distribution onto the gravity 
data given by
\begin{equation}
\mathbf{d}(\mathbf{p}) = \mathbf{A} \mathbf{p} \: ,
\label{eq:predicted-data-vector}
\end{equation}
where $\mathbf{d}$ is an $N \times 1$ vector whose $i$th element is the predicted gravity 
data at the $i$th point ($x_i$,$y_i$,$z_i$), $\mathbf{p}$ is the unknown $N \times 1$ 
parameter vector whose $j$th element $p_j$  is the mass of the $j$th equivalent source 
(point mass) at the $j$th Cartesian coordinates ($x_j$,$y_j$,$z_0$) and $\mathbf{A}$ 
is an $N \times N$  sensitivity matrix whose $ij$th element is given by 
\begin{equation}
a_{ij}= \frac{c_{g} \, G \, (z_{0} - z_{i})}{\left[(x_{i} - x_{j})^{2} +
(y_{i} - y_{j})^{2} +	(z_{i} - z_{0})^{2} \right]^{\frac{3}{2}}} \; ,
\label{eq:aij}
\end{equation}
where $G$ is the Newton's gravitational constant and $c_{g} = 10^{5}$ 
transforms from $\mathrm{m/s^2}$ to mGal.
Notice that the sensitivity matrix depends on the $i$th coordinates of the observations 
and the $j$th coordinates of the equivalent sources. For convenience, we designate 
these coordinates as \textit{matrix coordinates} and the indices $i$ and $j$ as 
\textit{matrix indices}.
In the classical equivalent-layer technique, we estimate the regularized parameter vector 
from the observed gravity data $\mathbf{d}^{o}$ by
\begin{equation}
\hat{\mathbf{p}} = \left( \mathbf{A}^{\top}\mathbf{A} + 
\mu \, \mathbf{I} \right)^{-1}
\mathbf{A}^{\top} \mathbf{d}^{o} \: .
\label{eq:estimated-p-parameter-space}
\end{equation}

\subsection{Fast equivalent-layer technique}
\citet{siqueira-etal2017} developed an iterative least-squares method to estimate the mass 
distribution over the equivalent layer based on the excess of mass and the positive correlation 
between the observed gravity data and the masses on the equivalent layer. Those authors showed 
that the fast equivalent-layer technique has a better computational efficiency than the 
classical equivalent layer approach (equation \ref{eq:estimated-p-parameter-space}) if the 
dataset totalize at least 200 observation points, even using a large number of iterations.

Considering one equivalent source (point mass) directly beneath each observation point, 
the iteration of the \citeauthor{siqueira-etal2017}'s~(\citeyear{siqueira-etal2017}) method 
starts by an initial approximation of mass distribution given by
\begin{equation}
\hat{\mathbf{p}}^0 = \tilde{\mathbf{A}}^{-1} \mathbf{d}^{o} \: ,
\label{eq:p0_fast_eqlayer}
\end{equation}
where $\tilde{\mathbf{A}}^{-1}$ is an $N \times N$ diagonal matrix with elements
\begin{equation}
\tilde{a}_{ii}^{-1} = \frac{\Delta s_i}{(2 \pi \, G \, c_g)} \: ,
\label{eq:aii_tilde_inv_fast_eqlayer}
\end{equation}
where $\Delta s_i$ is the $i$th element of surface area located at the $i$th horizontal 
coordinates $x_i$ and $y_i$ of the $i$th observation.
At the $k$th iteration, the masses of the equivalent sources are updated by
\begin{equation}
\hat{\mathbf{p}}^{k+1} = \hat{\mathbf{p}}^{k} + \mathbf{\Delta} \hat{\mathbf{p}}^{k} \: ,
\label{eq:p_k+1_fast_eqlayer}
\end{equation}
where the mass correction is given by
\begin{equation}
\mathbf{\Delta} \hat{\mathbf{p}}^{k+1} = \tilde{\mathbf{A}}^{-1} (\mathbf{d}^{o} - \mathbf{A} \hat{\mathbf{p}}^{k}) \: .
\label{eq:delta_p_k_fast_eqlayer}
\end{equation}

At the $k$th iteration of \citeauthor{siqueira-etal2017}'s~(\citeyear{siqueira-etal2017}) method, 
the matrix-vector product 
$\tensor{A} \hat{\mathbf{p}}^{k} = \mathbf{d}(\hat{\mathbf{p}}^{k})$ must be calculated to get a 
new residual $\mathbf{d^0} - \tensor{A} \hat{\mathbf{p}}^{k}$, which represents a bottleneck. 
Considering the limitation of 16 Gb of RAM 
memory in our system, we could run their method only up to 22 500 observation points. 
Hence, for very large data sets it is costful and can be overwhelming in terms of RAM 
memory to maintain such operation.

\subsection{Structure of matrix $\mathbf{A}$ for regular grids}

Consider that the observed data are located on an $N_{x} \times N_{y}$ regular grid of
points regularly spaced from $\Delta x$ and $\Delta y$ along the $x$ and $y$ directions,
respectively, on a horizontal plane defined by the constant vertical coordinate $z_{1} < z_{0}$. 
As a consequence, a given pair of matrix coordinates $(x_{i}, y_{i})$, defined by the matrix index 
$i$, $i = 1, \dots, N = N_{x} N_{y}$, is equivalent to a pair of coordinates $(x_{k}, y_{l})$
given by:
\begin{equation}
x_{i} \equiv x_{k} = x_{1} + \left[ k(i) - 1 \right] \, \Delta x \: , 
\label{eq:xi}
\end{equation}
and
\begin{equation}
y_{i} \equiv y_{l} = y_{1} + \left[ l(i) - 1 \right] \, \Delta y \: ,
\label{eq:yi}
\end{equation}
where $k(i)$ and $l(i)$ are integer functions of the matrix index $i$.
These equations can also be used to define the matrix coordinates 
$x_{j}$ and $y_{j}$ associated with the $j$-th equivalent source,
$j = 1, \dots, N = N_{x}N_{y}$. In this case, the integer functions
are evaluated by using the index $j$ instead of $i$.
For convenience, we designate $x_{k}$ and $y_{l}$ as \textit{grid coordinates}
and the indices $k$ and $l$ as \textit{grid indices}, which are computed with
the integer functions.

The integer functions assume different forms depending on the 
orientation of the regular grid of data.
Consider the case in which the grid is oriented along the
$x$-axis (Figure 1a). For convenience, we designate these grids as 
$x$-\textit{oriented grids}. For them, we have the following integer functions:
\begin{equation}
i(k, l) = (l - 1) \, N_{x} + k \quad ,
\label{eq:i-x-oriented}
\end{equation}
\begin{equation}
l(i) = \Bigg\lceil \frac{i}{N_{x}} \Bigg\rceil
\label{eq:l-x-oriented}
\end{equation}
and
\begin{equation}
k(i)  = i - \Bigg\lceil \frac{i}{N_{x}} \Bigg\rceil N_{x} + N_{x} \quad ,
\label{eq:k-x-oriented}
\end{equation}
where $\lceil \cdot \rceil$ denotes the ceiling function \citep[][ p. 67]{graham-etal1994}.
These integer functions are defined in terms of the matrix index $i$, but they can 
be defined in the same way by using the index $j$.
Figure 1a illustrates an $x$-oriented grid defined by $N_{x} = 4$ and $N_{y} = 3$.
In this example, the matrix coordinates $x_{7}$ and $y_{7}$, defined by the matrix index $i = 7$ (or $j = 7$), 
are equivalent to the grid coordinates $x_{3}$ and $y_{2}$, which are defined by the grid indices
$k = 3$ and $l = 2$, respectively. These indices are computed with equations \ref{eq:l-x-oriented}
and \ref{eq:k-x-oriented}, by using the matrix index $i = 7$ (or $j = 7$).

Now, consider the case in which the regular grid of data is oriented along 
the $y$-axis (Figure 1b). For convenience, we call them $y$-\textit{oriented grids}.
Similarly to $x$-oriented grids, we have the following integer functions associated with
$y$-oriented grids:
\begin{equation}
i(k, l) = (k - 1) \, N_{y} + l \quad ,
\label{eq:i-y-oriented}
\end{equation}
\begin{equation}
k(i) = \Bigg\lceil \frac{i}{N_{y}} \Bigg\rceil
\label{eq:k-y-oriented}
\end{equation}
and
\begin{equation}
l(i) = i - \Bigg\lceil \frac{i}{N_{y}} \Bigg\rceil N_{y} + N_{y} \quad .
\label{eq:l-y-oriented}
\end{equation}
Figure 1b illustrates an $y$-oriented grid defined by $N_{x} = 4$ and $N_{y} = 3$.
In this example, the matrix coordinates $x_{7}$ and $y_{7}$, defined by the matrix index 
$i = 7$ (or $j = 7$), are equivalent to the grid coordinates $x_{3}$ and $y_{1}$, which are 
defined by the grid indices $k = 3$ and $l = 1$, respectively. Differently from the example
shown in Figure 1a, the grid indices of the present example are 
computed with equations \ref{eq:k-y-oriented} and \ref{eq:l-y-oriented}, by using the 
matrix index $i = 7$ (or $j = 7$).

The element $a_{ij}$ (equation \ref{eq:aij}) can be rewritten 
by using equations \ref{eq:xi} and \ref{eq:yi}, giving rise to:
\begin{equation}
a_{ij} = \frac{c_{g} \, G \, \Delta z}{ \left[ 
	\left( \Delta k_{ij} \, \Delta x \right)^{2} + 
	\left( \Delta l_{ij} \, \Delta y \right)^{2} + 
	\left( \Delta z \right)^{2} \right]^{\frac{3}{2}}} \: ,
\label{eq:aij-regular-grids}
\end{equation}
where $\Delta z = z_{0} - z_{1}$, 
$\Delta k_{ij} = k(i) - k(j)$ (equations \ref{eq:k-x-oriented} and \ref{eq:k-y-oriented}) and
$\Delta l_{ij} = l(i) - l(j)$ (equations \ref{eq:l-x-oriented} and \ref{eq:l-y-oriented}).
Notice that the structure of matrix $\mathbf{A}$ (equation \ref{eq:predicted-data-vector}) for 
the case in which its elements are given by $a_{ij}$ (equation \ref{eq:aij-regular-grids}) is 
defined by the coefficients $\Delta k_{ij}$ and $\Delta l_{ij}$.

For $x$-oriented grids, the coefficients $\Delta k_{ij}$ and $\Delta l_{ij}$ are 
computed by using equations \ref{eq:k-x-oriented} and \ref{eq:l-x-oriented}, respectively.
In this case, $\mathbf{A}$ (equation \ref{eq:predicted-data-vector}) is 
composed of $N_{y} \times N_{y}$ blocks, where each block is formed by $N_{x} \times N_{x}$ elements.
For $y$-oriented grids, the coefficients $\Delta k_{ij}$ and $\Delta l_{ij}$ are 
computed by using equations \ref{eq:k-y-oriented} and \ref{eq:l-y-oriented}, respectively.
In this case, $\mathbf{A}$ (equation \ref{eq:predicted-data-vector}) is a 
composed of $N_{x} \times N_{x}$ blocks, where each block is formed by $N_{y} \times N_{y}$ elements.
In both cases, $\mathbf{A}$ is Toeplitz blockwise, i.e., the blocks lying at the same block 
diagonal are equal to each other.
Besides, the blocks located above the main diagonal are equal to those 
located below and each block is itself a Toeplitz matrix.
These symmetries come from the fact that the coefficients
$\Delta k_{ij}$ and $\Delta l_{ij}$ are squared at the denominator of 
$a_{ij}$ (equation \ref{eq:aij-regular-grids}).
Matrices with this well-defined pattern are called 
Doubly Block Toeplitz \citep[][ p. 28]{jain1989} or symmetric Block-Toeplitz Toeplitz-Block (BTTB),
for example. We opted for using the second term.

This well-defined pattern is better represented by using the \textit{block indices} $q$ and $p$. 
We represent $\mathbf{A}$ (equation \ref{eq:predicted-data-vector}) as a grid of $Q \times Q$ blocks 
$\mathbf{A}_{q}$, $q = 0, \dots, Q - 1$, given by
\begin{equation}
	\mathbf{A} = \begin{bmatrix}
	\mathbf{A}_{0}   & \mathbf{A}_{1} & \cdots         & \mathbf{A}_{Q-1} \\
	\mathbf{A}_{1}   & \mathbf{A}_{0} & \ddots         & \vdots           \\ 
	\vdots           & \ddots         & \ddots         & \mathbf{A}_{1}   \\
	\mathbf{A}_{Q-1} & \cdots         & \mathbf{A}_{1} & \mathbf{A}_{0}                 
	\end{bmatrix}_{N \times N} \: ,
	\label{eq:BTTB_A}
\end{equation}
where each block has $P \times P$ elements conveniently represented by $a^{q}_{p}$, 
$p = 0, \dots, P - 1$, as follows:
\begin{equation}
	\mathbf{A}_{q} = \begin{bmatrix}
	a^{q}_{0}   & a^{q}_{1} & \cdots    & a^{q}_{P-1} \\
	a^{q}_{1}   & a^{q}_{0} & \ddots    & \vdots           \\ 
	\vdots      & \ddots    & \ddots    & a^{q}_{1}   \\
	a^{q}_{P-1} & \cdots    & a^{q}_{1} & a^{q}_{0}                 
	\end{bmatrix}_{P \times P} \: ,
	\label{eq:Aq_block}
\end{equation}
with $N = QP$. The index $q$ defines the block diagonal where $\mathbf{A}_{q}$ (equation \ref{eq:Aq_block}) 
lies within the BTTB matrix $\mathbf{A}$ (equation \ref{eq:BTTB_A}). 
This index varies from $0$, at the main diagonal, to $Q - 1$, at
the corners of $\mathbf{A}$. Similarly, the index $p$ defines the diagonal where $a^{q}_{p}$ 
lies within $\mathbf{A}_{q}$ (equation \ref{eq:Aq_block}). This index varies from $0$, 
at the main diagonal, to $P - 1$, at the corners of $\mathbf{A}_{q}$.
For $x$-oriented grids, $Q = N_{y}$, $P = N_{x}$ and the block indices
$q$ and $p$ are defined, respectively, by the following integer functions 
of the matrix indices $i$ and $j$:
\begin{equation}
q(i, j) = \; \mid l(i) - l(j) \mid
\label{eq:q-x-oriented}
\end{equation}
and
\begin{equation}
p(i, j) = \; \mid k(i) - k(j) \mid \quad ,
\label{eq:p-x-oriented}
\end{equation}
where $l(i)$ and $l(j)$ are defined by equation \ref{eq:l-x-oriented} 
and $k(i)$ and $k(j)$ are defined by equation \ref{eq:k-x-oriented}.
For $y$-oriented grids, $Q = N_{x}$, $P = N_{y}$ and the block indices
$q$ and $p$ are defined, respectively, by the following integer functions 
of the matrix indices $i$ and $j$:
\begin{equation}
q(i, j) = \; \mid k(i) - k(j) \mid 
\label{eq:q-y-oriented}
\end{equation}
and
\begin{equation}
p(i, j) = \; \mid l(i) - l(j) \mid \quad ,
\label{eq:p-y-oriented}
\end{equation}
where $k(i)$ and $k(j)$ are defined by equation \ref{eq:k-y-oriented}
and $l(i)$ and $l(j)$ are defined by equation \ref{eq:l-y-oriented}.
Notice that, for each element $a_{ij}$ (equation \ref{eq:aij-regular-grids}),
defined by matrix indices $i$ and $j$, there is a corresponding block element
$a^{q}_{p}$, defined by block indices $q$ (equations \ref{eq:q-x-oriented} and 
\ref{eq:q-y-oriented}) and $p$ (equations \ref{eq:p-x-oriented} and \ref{eq:p-y-oriented}),
so that
\begin{equation}
	a^{q}_{p} \equiv a_{ij} \quad .
	\label{eq:aqp_equiv_aij}
\end{equation}
We also stress that matrix $\mathbf{A}$ (equation \ref{eq:predicted-data-vector}) defined 
by elements $a_{ij}$ (equation \ref{eq:aij-regular-grids}) in terms of matrix indices 
$i$ and $j$ is strictly the same BTTB matrix $\mathbf{A}$ (equation \ref{eq:BTTB_A}) defined 
by the blocks $\mathbf{A}_{q}$ (equation \ref{eq:Aq_block}) and block elements 
$a^{q}_{p}$ (equation \ref{eq:aqp_equiv_aij}) in terms of the block indices 
$q$ (equations \ref{eq:q-x-oriented} and \ref{eq:q-y-oriented}) and $p$ 
(equations \ref{eq:p-x-oriented} and \ref{eq:p-y-oriented}).

It is important noting that different matrix indices $i$ or $j$ produce the same 
absolute values for the grid indices $k$ (equations \ref{eq:k-x-oriented} and
\ref{eq:k-y-oriented}) and $l$ (equations \ref{eq:l-x-oriented} and
\ref{eq:l-y-oriented}). As a consequence, different pairs of matrix indices $i$
and $j$ generate the same absolute values for the coefficients $\Delta k_{ij}$ and
$\Delta l_{ij}$ that compose the denominator of $a_{ij}$ 
(equation \ref{eq:aij-regular-grids}) and also the same values for the block indices
$q$ (equations \ref{eq:q-x-oriented} and \ref{eq:q-y-oriented}) and 
$p$ (equations \ref{eq:p-x-oriented} and \ref{eq:p-y-oriented}), as well as the same
block element $a^{q}_{p}$ (equation \ref{eq:aqp_equiv_aij}). 
It means that elements $a_{ij}$ defined
by different matrix indices $i$ and $j$ have the same value. The key point for
understanding the structure of BTTB matrix $\mathbf{A}$ (equation \ref{eq:BTTB_A})
is then, given a single element $a_{ij}$
defined by matrix indices $i$ and $j$, compute the grid indices 
$k$ (equations \ref{eq:k-x-oriented} and \ref{eq:k-y-oriented}) and
$l$ (equations \ref{eq:l-x-oriented} and \ref{eq:l-y-oriented}).
These grid indices are used to 
(1) compute the coefficients $\Delta k_{ij}$ and 
$\Delta l_{ij}$ and determine the value of $a_{ij}$ with equation 
\ref{eq:aij-regular-grids} and 
(2) compute the block indices $q$ (equations \ref{eq:q-x-oriented}, 
and \ref{eq:q-y-oriented}) and $p$ (equations \ref{eq:p-x-oriented} and
and \ref{eq:p-y-oriented}) and determine the corresponding block element $a^{q}_{p}$ 
(equation \ref{eq:aqp_equiv_aij}) forming $\mathbf{A}_{q}$ (equation \ref{eq:Aq_block}).

Consider the $x$-oriented grid of $N_{x} \times N_{y}$ points shown in Figure 1a, 
with $N_{x} = 4$, $N_{y} = 3$ and $N = N_{x} \, N_{y} = 12$.
To illustrate the relationship between the matrix indices ($i$ and $j$) and 
the block indices ($q$ and $p$), consider the element $a_{ij}$ defined by 
$i = 2$ and $j = 10$, which is
located at the upper right corner of $\mathbf{A}$ (equation \ref{eq:predicted-data-vector}), 
in the 2nd line and 10th column.
By using equations \ref{eq:l-x-oriented} and \ref{eq:k-x-oriented}, we obtain the 
grid indices $l(i) = 1$, $l(j) = 3$, $k(i) = 2$ and $k(j) = 2$.
These grid indices result in the coefficients $\Delta k_{ij} = 0$ and $\Delta l_{ij} = -2$,
which are used to compute the element $a_{ij}$ (equation \ref{eq:aij-regular-grids}),
as well as in the block indices $q = 2$ (equation \ref{eq:q-x-oriented}) and 
$p = 0$ (equation \ref{eq:p-x-oriented}).
These block indices indicate that this element $a_{ij}$ appears in the main diagonal
of the blocks $\mathbf{A}_{2}$ (equation \ref{eq:Aq_block}), which are located at the corners 
of $\mathbf{A}$ (equation \ref{eq:BTTB_A}).
To verify this, let us take the matrix indices associated with these elements.
They are $(i, j)$ = $(1, 9)$, $(2, 10)$, $(3, 11)$, $(4, 12)$, $(9, 1)$, $(10, 2)$, 
$(11, 3)$ and $(12, 4)$. By using these matrix indices, it is easy to verify that all
of them produce the same grid indices $l(i)$, $l(j)$, $k(i)$ and $k(j)$ 
(equations \ref{eq:l-x-oriented} and \ref{eq:k-x-oriented}) as those associated with
the element defined by $i = 2$ and $j = 10$. Consequently, all of them produce
elements $a_{ij}$ (equation \ref{eq:aij-regular-grids}) having the same value.
Besides, it is also easy to verify that all these matrix indices produce the same block
indices $q = 2$ (equation \ref{eq:q-x-oriented}) and $p = 0$ (equation \ref{eq:p-x-oriented})
and the same block element $a^{q}_{p}$ (equation \ref{eq:aqp_equiv_aij}).
By repeating this procedure for all elements $a_{ij}$, $i = 1, \dots, 12$, $j = 1, \dots, 12$, 
forming the matrix $\mathbf{A}$ (equation \ref{eq:predicted-data-vector}) obtained from our 
$x$-oriented grid (Figure 1a), we can verify that
\begin{equation}
\mathbf{A} = \begin{bmatrix}
\mathbf{A}_{0} & \mathbf{A}_{1} & \mathbf{A}_{2} \\
\mathbf{A}_{1} & \mathbf{A}_{0} & \mathbf{A}_{1} \\
\mathbf{A}_{2} & \mathbf{A}_{1} & \mathbf{A}_{0}
\end{bmatrix} \quad ,
\label{eq:A-x-oriented-example}
\end{equation}
where $\mathbf{A}_{q}$ (equation \ref{eq:Aq_block}), $q = 0, \dots, Q -1$, $Q = N_{y}$, 
are symmetric Toeplitz matrices given by:
\begin{equation}
\mathbf{A}_{q} = \begin{bmatrix}
a^{q}_{0} & a^{q}_{1} & a^{q}_{2} & a^{q}_{3} \\
a^{q}_{1} & a^{q}_{0} & a^{q}_{1} & a^{q}_{2} \\
a^{q}_{2} & a^{q}_{1} & a^{q}_{0} & a^{q}_{1} \\
a^{q}_{3} & a^{q}_{2} & a^{q}_{1} & a^{q}_{0}
\end{bmatrix} \quad ,
\label{eq:Aq-x-oriented}
\end{equation}
with elements $a^{q}_{p}$ (equation \ref{eq:aqp_equiv_aij}) defined by 
$p = 0, \dots, P - 1$, $P = N_{x}$.

This procedure can also be used to verify that the matrix $\mathbf{A}$ 
(equation \ref{eq:predicted-data-vector}) obtained
from the $y$-oriented grid illustrated in Figure 1b is given by
\begin{equation}
\mathbf{A} = \begin{bmatrix}
\mathbf{A}_{0} & \mathbf{A}_{1} & \mathbf{A}_{2} & \mathbf{A}_{3} \\
\mathbf{A}_{1} & \mathbf{A}_{0} & \mathbf{A}_{1} & \mathbf{A}_{2} \\
\mathbf{A}_{2} & \mathbf{A}_{1} & \mathbf{A}_{0} & \mathbf{A}_{1} \\
\mathbf{A}_{3} & \mathbf{A}_{2} & \mathbf{A}_{1} & \mathbf{A}_{0}
\end{bmatrix} \quad ,
\label{eq:A-y-oriented-example}
\end{equation}
where $\mathbf{A}_{q}$ (equation \ref{eq:Aq_block}), $q = 0, \dots, Q - 1$, $Q = N_{x}$, 
are symmetric Toeplitz matrices given by:
\begin{equation}
\mathbf{A}_{q} = \begin{bmatrix}
a^{q}_{0} & a^{q}_{1} & a^{q}_{2} \\
a^{q}_{1} & a^{q}_{0} & a^{q}_{1} \\
a^{q}_{2} & a^{q}_{1} & a^{q}_{0}
\end{bmatrix} \quad ,
\label{eq:Aq-y-oriented}
\end{equation}
with elements $a^{q}_{p}$ (equation \ref{eq:aqp_equiv_aij}) defined by 
$p = 0, \dots, P - 1$, $P = N_{y}$.

These examples (equations \ref{eq:A-x-oriented-example}--\ref{eq:Aq-y-oriented}) show 
that the entire $N \times N$ BTTB matrix $\mathbf{A}$ 
(equations \ref{eq:predicted-data-vector} and \ref{eq:BTTB_A}) 
can be defined by using only the elements 
forming its first column (or row). Notice that this column contains the gravitational effect 
produced by a single equivalent source at all $N$ observation points.

\subsection{BTTB matrix-vector product}

The matrix-vector product $\tensor{A} \hat{\mathbf{p}}^{k}$ (equation 
\ref{eq:delta_p_k_fast_eqlayer}) required by the fast equivalent-layer 
technique \citep{siqueira-etal2017} accounts for most of its total computation time 
and can cause RAM memory shortage when large data sets are used.
This computational load can be drastically lessen by exploring the well-defined structure of 
matrix $\mathbf{A}$ (equation \ref{eq:predicted-data-vector}) for the particular case in which 
its elements $a_{ij}$ are defined by equation \ref{eq:aij-regular-grids}. 
In this case, $\mathbf{A}$ is a symmetric BTTB matrix (equations \ref{eq:BTTB_A} and 
\ref{eq:A-x-oriented-example}--\ref{eq:Aq-y-oriented}) and the predicted data vector 
$\mathbf{d}(\mathbf{p})$ (equation \ref{eq:predicted-data-vector}) can be efficiently
computed by using the two-dimensional Discrete Fourier Transform (DFT).
To do this, let us first rewrite $\mathbf{d}(\mathbf{p})$ and
$\mathbf{p}$ (equation \ref{eq:predicted-data-vector}) as the following partitioned vectors:
\begin{equation}
\mathbf{d}(\mathbf{p}) = \begin{bmatrix}
\mathbf{d}_{0}(\mathbf{p}) \\
\vdots \\
\mathbf{d}_{Q - 1}(\mathbf{p})
\end{bmatrix}_{N \times 1}
\label{eq:predicted-data-vector-partitioned}
\end{equation}
and
\begin{equation}
\mathbf{p} = \begin{bmatrix}
\mathbf{p}_{0} \\
\vdots \\
\mathbf{p}_{Q - 1}
\end{bmatrix}_{N \times 1} \quad ,
\label{eq:parameter-vector-partitioned}
\end{equation}
where $\mathbf{d}_{q}(\mathbf{p})$ and $\mathbf{p}_{q}$, $q = 0, \dots, Q - 1$,
are $P \times 1$ vectors. Notice that $q$ is the block index defined by equations 
\ref{eq:q-x-oriented} and \ref{eq:q-y-oriented}, $Q$ defines the number of blocks
$\mathbf{A}_{q}$ (equation \ref{eq:Aq_block}) forming $\mathbf{A}$ (equation \ref{eq:BTTB_A}) 
and $P$ defines the number of elements forming each block $\mathbf{A}_{q}$.
Then, by using the partitioned vectors 
(equations \ref{eq:parameter-vector-partitioned} and \ref{eq:predicted-data-vector-partitioned}) 
and remembering that $N = QP$, we define the auxiliary linear system
\begin{equation}
\mathbf{w} = \mathbf{C} \mathbf{v} \: ,
\label{eq:w_Cv}
\end{equation}
where
\begin{equation}
\mathbf{w} = \begin{bmatrix}
\mathbf{w}_{0} \\
\vdots \\
\mathbf{w}_{Q - 1} \\
\mathbf{0}_{2N \times 1}
\end{bmatrix}_{4N \times 1} \quad ,
\label{eq:w-vector}
\end{equation}
\begin{equation}
\mathbf{w}_{q} = \begin{bmatrix}
\mathbf{d}_{q}(\mathbf{p}) \\
\mathbf{0}_{P \times 1}
\end{bmatrix}_{2P \times 1}
\label{eq:wq-vector} \quad ,
\end{equation}
\begin{equation}
\mathbf{v} = \begin{bmatrix}
\mathbf{v}_{0} \\
\vdots \\
\mathbf{v}_{Q - 1} \\
\mathbf{0}_{2N \times 1}
\end{bmatrix}_{4N \times 1} \quad ,
\label{eq:v-vector}
\end{equation}
and
\begin{equation}
\mathbf{v}_{q} = \begin{bmatrix}
\mathbf{p}_{q} \\
\mathbf{0}_{P \times 1}
\end{bmatrix}_{2P \times 1}
\label{eq:vq-vector} \quad ,
\end{equation}
with $\mathbf{d}_{q}(\mathbf{p})$ and $\mathbf{p}_{q}$ defined by
equations \ref{eq:predicted-data-vector-partitioned} and 
\ref{eq:parameter-vector-partitioned}, respectively.
Finally $\mathbf{C}$ (equation \ref{eq:w_Cv}) is a 
$4N \times 4N$ symmetric Block Circulant matrix with Circulant Blocks (BCCB) 
\citep[][ p. 184]{davis1979}.
What follows shows a step-by-step description of how we use the auxiliary 
system (equation \ref{eq:w_Cv}) to compute the matrix-vector product 
$\tensor{A} \hat{\mathbf{p}}^{k}$ (equation \ref{eq:delta_p_k_fast_eqlayer}) in 
a computationally efficient way by exploring the structure of matrix $\mathbf{C}$.
We show that this approach, that have been used in potential-field methods
\citep[e.g.,][]{zhang-wong2015, zhang-etal2016, qiang_etal2019}, is actually a fast 
convolution method \citep[e.g.,][ p. 213]{vanloan1992}.

Matrix $\mathbf{C}$ (equation \ref{eq:w_Cv})
is circulant blockwise, formed by $Q \times Q$ blocks, where
each block $\mathbf{C}_{q}$, $q = 0, \dots, Q-1$, is a $2P \times 2P$ circulant matrix. 
Similarly to the BTTB matrix $\mathbf{A}$ (equations \ref{eq:BTTB_A} and 
\ref{eq:A-x-oriented-example}--\ref{eq:Aq-y-oriented}), 
the index $q$ defines the block diagonal where $\mathbf{C}_{q}$ lies 
within $\mathbf{C}$. This index varies from $0$, at the main diagonal, to $Q - 1$, at
the corners of $\mathbf{C}$. Additionally, the blocks lying 
above the main diagonal are equal to those located below.
It is well-known that a circulant matrix can be defined by properly downshifting 
its first column \citep[][ p. 206]{vanloan1992}. Hence, the BCCB matrix $\mathbf{C}$ 
(equation \ref{eq:w_Cv}) can be obtained from its 
first column of blocks, which is given by
\begin{equation}
\left[\mathbf{C} \right]_{(0)} = 
	\begin{bmatrix}
	\mathbf{C}_{0} \\
	\vdots \\
	\mathbf{C}_{Q-1} \\
	\mathbf{0} \\
	\mathbf{C}_{Q-1} \\
	\vdots \\
	\mathbf{C}_{1}
	\end{bmatrix}_{4N \times 2P} \: ,
	\label{eq:C-first-column-blocks}
\end{equation}
where $\mathbf{0}$ is a $2P \times 2P$ matrix of zeros. Similarly, each block 
$\mathbf{C}_{q}$, $q = 0, \dots, Q-1$, can be obtained by downshifting its first 
column
\begin{equation}
\mathbf{c}^{q}_{0} = 
	\begin{bmatrix}
		a^{q}_{0} \\
		\vdots \\
		a^{q}_{P-1} \\
		0 \\
		a^{q}_{P-1} \\
		\vdots \\
		a^{q}_{1}
	\end{bmatrix}_{2P \times 1} \: ,
	\label{eq:Cq-first-column}
\end{equation}
where $a^{q}_{p}$ (equation \ref{eq:aqp_equiv_aij}), $p = 0, \dots, P-1$, are the elements 
forming the block $\mathbf{A}_{q}$ (equations \ref{eq:Aq_block} and 
\ref{eq:A-x-oriented-example}--\ref{eq:Aq-y-oriented}).
The downshift can be thought off as permutation that pushes the components of a column vector 
down one notch with wraparound \citep[][ p. 20]{golub-vanloan2013}.
To illustrate this operation, consider our $y$-oriented grid illustrated in Figure 1b. 
In this case, the resulting 
BCCB matrix $\mathbf{C}$ (equation \ref{eq:w_Cv}) is given by 
\begin{equation}
\mathbf{C} =
\begin{bmatrix}
	\mathbf{C_{0}} & \mathbf{C_{1}} & \mathbf{C_{2}} & \mathbf{C_{3}} & \mathbf{0}     & \mathbf{C_{3}} & \mathbf{C_{2}} & \mathbf{C_{1}} \\
	\mathbf{C_{1}} & \mathbf{C_{0}} & \mathbf{C_{1}} & \mathbf{C_{2}} & \mathbf{C_{3}} & \mathbf{0}     & \mathbf{C_{3}} & \mathbf{C_{2}} \\
	\mathbf{C_{2}} & \mathbf{C_{1}} & \mathbf{C_{0}} & \mathbf{C_{1}} & \mathbf{C_{2}} & \mathbf{C_{3}} & \mathbf{0}     & \mathbf{C_{3}} \\
	\mathbf{C_{3}} & \mathbf{C_{2}} & \mathbf{C_{1}} & \mathbf{C_{0}} & \mathbf{C_{1}} & \mathbf{C_{2}} & \mathbf{C_{3}} & \mathbf{0}     \\
	\mathbf{0}     & \mathbf{C_{3}} & \mathbf{C_{2}} & \mathbf{C_{1}} & \mathbf{C_{0}} & \mathbf{C_{1}} & \mathbf{C_{2}} & \mathbf{C_{3}} \\
	\mathbf{C_{3}} & \mathbf{0}     & \mathbf{C_{3}} & \mathbf{C_{2}} & \mathbf{C_{1}} & \mathbf{C_{0}} & \mathbf{C_{1}} & \mathbf{C_{2}} \\
	\mathbf{C_{2}} & \mathbf{C_{3}} & \mathbf{0}     & \mathbf{C_{3}} & \mathbf{C_{2}} & \mathbf{C_{1}} & \mathbf{C_{0}} & \mathbf{C_{1}} \\
	\mathbf{C_{1}} & \mathbf{C_{2}} & \mathbf{C_{3}} & \mathbf{0}     & \mathbf{C_{3}} & \mathbf{C_{2}} & \mathbf{C_{1}} & \mathbf{C_{0}}
\end{bmatrix},
\label{eq:C-y-oriented}
\end{equation}
where each block $\mathbf{C}_{q}$, $q = 0, 1, 2$, is represented as follows 
\begin{equation}
\tensor{C}_{q} =
\begin{bmatrix}
	a^{q}_{0} & a^{q}_{1} & a^{q}_{2} & 0         & a^{q}_{2} & a^{q}_{1} \\
	a^{q}_{1} & a^{q}_{0} & a^{q}_{1} & a^{q}_{2} & 0         & a^{q}_{2} \\
	a^{q}_{2} & a^{q}_{1} & a^{q}_{0} & a^{q}_{1} & a^{q}_{2} & 0         \\
	0         & a^{q}_{2} & a^{q}_{1} & a^{q}_{0} & a^{q}_{1} & a^{q}_{2} \\
	a^{q}_{2} & 0         & a^{q}_{2} & a^{q}_{0} & a^{q}_{0} & a^{q}_{1} \\
	a^{q}_{1} & a^{q}_{2} & 0         & a^{q}_{2} & a^{q}_{1} & a^{q}_{0}
\end{bmatrix}
\label{eq:Cq-y-oriented}
\end{equation}
in terms of the block elements $a^{q}_{p}$ (equation \ref{eq:aqp_equiv_aij}).
Similar matrices are obtained for our $x$-oriented grid illustrated in Figure 1a.

BCCB matrices are diagonalized by the two-dimensional unitary DFT 
\citep[][ p. 185]{davis1979}. It means that $\mathbf{C}$ (equation \ref{eq:w_Cv}) 
satisfies 
\begin{equation}
\mathbf{C} = 
\left(\mathbf{F}_{2Q} \otimes \mathbf{F}_{2P} \right)^{\ast} 
\boldsymbol{\Lambda}
\left(\mathbf{F}_{2Q} \otimes \mathbf{F}_{2P} \right) \: ,
\label{eq:C-diagonalized}
\end{equation}
where the symbol ``$\otimes$" denotes the Kronecker product \citep{neudecker1969},
$\mathbf{F}_{2Q}$ and $\mathbf{F}_{2P}$ are the $2Q \times 2Q$ and $2P \times 2P$ 
unitary DFT matrices \citep[][ p. 31]{davis1979}, respectively, the superscritpt 
``$\ast$" denotes the complex conjugate and $\boldsymbol{\Lambda}$ is a 
$4QP \times 4QP$ diagonal matrix containing the eigenvalues of $\mathbf{C}$.
By substituting equation \ref{eq:C-diagonalized} in the auxiliary system 
(equation \ref{eq:w_Cv}) and premultiplying both
sides of the result by $\left(\mathbf{F}_{2Q} \otimes \mathbf{F}_{2P} \right)$, 
we obtain
\begin{equation}
\boldsymbol{\Lambda} \left(\mathbf{F}_{2Q} \otimes \mathbf{F}_{2P} \right) 
\mathbf{v} = \left(\mathbf{F}_{2Q} \otimes \mathbf{F}_{2P} \right) 
\mathbf{w} \: .
\label{eq:vec-DFT-system}
\end{equation}
Now, by applying the $vec$-operator to both sides of equation \ref{eq:vec-DFT-system} 
(see the details in Appendix A), we obtain:
\begin{equation}
\mathbf{F}_{2Q}^{\ast} \left[ 
\mathbf{L} \circ \left(\mathbf{F}_{2Q} \, \mathbf{V} \, \mathbf{F}_{2P} \right) 
\right] \mathbf{F}_{2P}^{\ast} = \mathbf{W} \: ,
\label{eq:DFT-system}
\end{equation}
where ``$\circ$'' denotes the Hadamard product \citep[][ p. 298]{horn_johnson1991} and 
$\mathbf{L}$, $\mathbf{V}$ and $\mathbf{W}$ are $2Q \times 2P$ matrices obtained 
by rearranging, along their rows, the elements forming the diagonal of matrix 
$\boldsymbol{\Lambda}$, vector $\mathbf{v}$ and vector $\mathbf{w}$, respectively.
The left side of equation \ref{eq:DFT-system} contains the two-dimensional 
Inverse Discrete Fourier Transform (IDFT) of the term in brackets, which in turn
represents the Hadamard product of matrix $\mathbf{L}$ (equation \ref{eq:left_side_DFT_system_3})
and the two-dimensional DFT of matrix $\mathbf{V}$ (equation \ref{eq:left_side_DFT_system_1}).
Matrix $\mathbf{L}$ contains the eigenvalues 
of $\boldsymbol{\Lambda}$ (equation \ref{eq:C-diagonalized}) and can be 
efficiently computed by using only the first column of the BCCB matrix 
$\mathbf{C}$ (equation \ref{eq:w_Cv}) (see the details in Appendix B).
Here, we evaluate equation \ref{eq:DFT-system} and compute matrix $\mathbf{L}$
by using the two-dimensional Fast Fourier Transform (2D FFT).
This approach is a two-dimensional discrete convolution \citep[e.g.,][ p. 213]{vanloan1992}.

At each iteration $k$th of the fast equivalent-layer technique, 
(equation \ref{eq:delta_p_k_fast_eqlayer}), we efficiently compute 
$\mathbf{A} \hat{\mathbf{p}}^{k} = \mathbf{d}(\hat{\mathbf{p}}^{k})$ by following 
the steps below:

\begin{itemize}
\item[\textbf{(1)}] Use equation \ref{eq:aij-regular-grids} to compute the first column 
of each block $\mathbf{A}_{q}$ (equation \ref{eq:Aq_block}), $q = 0, \dots, Q-1$, forming 
the BTTB matrix $\mathbf{A}$ (equation \ref{eq:BTTB_A});

\item[\textbf{(2)}] Rearrange the first column of $\mathbf{A}$ according to equations 
\ref{eq:C-first-column-blocks} and \ref{eq:Cq-first-column} to obtain the
first column $\mathbf{c}_{0}$ of the BCCB matrix $\mathbf{C}$ (equation \ref{eq:w_Cv});

\item[\textbf{(3)}] Rearrange $\mathbf{c}_{0}$ along the rows of matrix $\mathbf{G}$ and
use the 2D FFT to compute matrix $\mathbf{L}$ (equation \ref{eq:DTF_G});

\item[\textbf{(4)}] Rearrange the parameter vector $\hat{\mathbf{p}}^{k}$ 
(equation \ref{eq:predicted-data-vector}) in its partitioned form 
(equation \ref{eq:parameter-vector-partitioned}) to define the auxiliary vector 
$\mathbf{v}$ (equation \ref{eq:v-vector});

\item[\textbf{(5)}] Rearrange $\mathbf{v}$ to obtain matrix $\mathbf{V}$, use the 2D FFT 
to compute its DFT and evaluate the left side of equation \ref{eq:DFT-system-preliminary};

\item[\textbf{(6)}] Use the 2D FFT to compute the IDFT of the result obtained at item (5) and 
obtain the matrix $\mathbf{W}$ (equation \ref{eq:DFT-system});

\item[\textbf{(7)}] Use the $vec$-operator (equation \ref{eq:vec-operator}) and equations 
\ref{eq:w-vector} and \ref{eq:wq-vector} to rearrange $\mathbf{W}$ and obtain the predicted 
data vector $\mathbf{d}(\hat{\mathbf{p}}^{k})$.

\end{itemize}

\subsection{Computational performance}

The number of flops (floating-point operations) necessary to estimate the 
$N \times 1$ parameter vector $\mathbf{p}$ in the fast equivalent-layer technique 
\citep{siqueira-etal2017} is
\begin{equation}
f_{0} = N^{it} (3N + 2N^{2}) \; ,
\label{eq:float_fast_eqlayer}
\end{equation}
where $N^{it}$ is the number of iterations. In this equation, the term $2N^2$ is associated 
with the matrix-vector product $\mathbf{A} \hat{\mathbf{p}}^{k}$ (equation 
\ref{eq:delta_p_k_fast_eqlayer}) and accounts for most of the computational complexity 
of this method.
Our method replace this matrix-vector product by three operations: 
one DFT, one Hadammard product and one IDFT involving $2Q \times 2P$ matrices 
(left side of equation \ref{eq:DFT-system}). 
The Hadamard product requires six times $4N$ flops, $N = QP$, because the entries are 
complex numbers.
We consider that a DFT/IDFT requires $\kappa \, 4N \log_{2}(4N)$ flops to be computed via 2D FFT, 
where $\kappa$ is a constant depending on the algorithm. 
Then, the resultant flops count of our method is given by:
\begin{equation}
f_{1} = N^{it} \left[ 27N + \kappa \, 8N \log_{2}(4N) \right] \: .
\label{eq:float_bccb}
\end{equation}
Figure \ref{fig:float} shows the flops counts $f_{0}$ and $f_{1}$ (equations \ref{eq:float_fast_eqlayer}
and \ref{eq:float_bccb}) associated with the fast equivalent-layer technique \citep{siqueira-etal2017} and 
our method, respectively, as a function of the number $N$ of observation points. 
We considered a fixed number of $N^{it} = 50$ iterations and $\kappa = 5$ (equation \ref{eq:float_bccb}),
which is compatible with a radix-2 FFT \citep[][ p. 16]{vanloan1992}.
As we can see, the number of flops is drastically decreased in our method.

Another advantage of our method is concerned with the real $N \times N$ matrix $\mathbf{A}$ 
(equations \ref{eq:predicted-data-vector} and \ref{eq:BTTB_A}).
In the fast equivalent-layer technique, the full matrix 
is computed once and stored during the entire iterative process.
On the other hand, our method computes only one column of $\mathbf{A}$ 
and uses it to compute the complex $2Q \times 2P$ matrix $\mathbf{L}$ 
(equation \ref{eq:DTF_G}) via 2D FFT, which is stored during all iterations.
Table \ref{tab:RAM-usage} shows the RAM memory usage needed to store the 
full matrix $\mathbf{A}$, a single column of $\mathbf{A}$ and the full matrix 
$\mathbf{L}$. The quantities were computed for different numbers of observations $N$. 
Notice that $N = 1\,000\,000$ observations require nearly $7.6$ Terabytes of RAM memory 
to store the whole matrix $\mathbf{A}$.

Figure \ref{fig:time_fast_eqlayer_bccb} compares the running time of the fast equivalent-layer technique 
\citep{siqueira-etal2017} and of our method, considering a constant number of iterations $N^{it} = 50$. 
We used a PC with an Intel Core i7 4790@3.6GHz processor and 16 GB of RAM memory.
The computational efficiency of our approach is significantly superior to that of the 
fast equivalent-layer technique for a number of observations $N$ greater than $10\,000$. 
We could not perform this comparison with a number of observations greater than $22\,500$
due to limitations of our PC in storing the full matrix $\mathbf{A}$.
Figure \ref{fig:time_bccb} shows the running time of our method with a number of observations 
up to  $25$ millions. 
These results shows that, while the running time of our method is $26.8$ seconds for 
$N = 1\,000\,000$, the fast equivalent-layer technique spends $48.3$ seconds for $N = 22\,500$.

### 2019-11-29 manuscript/synthetic_data.tex
split the tex files

\section{Synthetic tests}

In this section, we  investigate the effectiveness of using the properties of BTTB and BCCB matrices 
(equation \ref{eq:w_Cv})  to solve, at each iteration, the forward modeling (the matrix-vector product 
$\mathbf{A} \hat{\mathbf{p}}^k$)  required in the fast equivalent-layer method proposed by \citet{siqueira-etal2017}. 
We simulated three sources whose horizontal projections are shown in Figure \ref{fig:synthetic_data} as black lines. 
These sources are two vertical prisms with density contrasts of $0.35\, \mathrm{g/cm^3}$ (upper-left prism) and 
$0.4\, \mathrm{g/cm^3}$ (upper-right prism) and a sphere with radius of $1\,000$ m with density contrast of 
$-0.5\, \mathrm{g/cm^3}$. Figure \ref{fig:synthetic_data}  shows the vertical component of gravity field generated 
by these sources contaminated with additive pseudorandom Gaussian noise with zero mean and standard deviation of 
$0.015$ mGal.

The advantage of using the structures of BTTB and BCCB matrices to compute forward modeling in  the fast-equivalent 
layer method \cite[]{siqueira-etal2017} is grounded on the use of regular grids of data and equivalent sources. 
Hence, we created $10\,000$ observation points regularly spaced in a grid of $100 \times 100$ at $100$ m height. 
We also set a grid of equivalent point masses, each one directly beneath each observation points, located at 
$300$ m deep.  Figures \ref{fig:classic_fast_val}a and \ref{fig:bccb_fast_val}a  show the fitted gravity data 
obtained, respectively, by the fast equivalent layer method and by our modified form of this method that computes 
the forward modeling using equation \ref{eq:w_Cv}. The corresponding residuals (Figures \ref{fig:classic_fast_val}b 
and \ref{fig:bccb_fast_val}b), defined as the difference between the observed (Figure  \ref{fig:synthetic_data}) 
and fitted gravity data (Figures \ref{fig:classic_fast_val}a and \ref{fig:bccb_fast_val}a), show means close to 
zero  and standard deviations of 0.0144 mGal.  Therefore, Figures \ref{fig:classic_fast_val} and 
\ref{fig:bccb_fast_val} show that \citet{siqueira-etal2017} method and our modified version of this method produced 
virtually the same results. This excellent agreement is confirmed in Figures \ref{fig:delta_comparison} and 
\ref{fig:delta_rho} which shows that there are virtually no differences, respectively,  in the fitted data presented 
in Figures \ref{fig:classic_fast_val}b and \ref{fig:bccb_fast_val}b and in the estimated mass distributions within 
the equivalent layers (not shown) yielded by both \citet{siqueira-etal2017} method and our modification of this method. 
These results (Figures \ref{fig:delta_comparison} and \ref{fig:delta_rho})  show that computing the matrix-vector 
product ($\mathbf{A} \hat{\mathbf{p}}^k$), required in the forward modeling, by means of embedding the BTTB matrix into 
a BCCB matrix (equation \ref{eq:w_Cv}) yields practically the same result as the one produced by computing this 
matrix-vector product  with a full matrix $\mathbf{A}$ as used in \citet{siqueira-etal2017}.

We perform two forms of processing the gravity data (Figure \ref{fig:synthetic_data}) through the equivalent layer 
technique: the upward (Figure \ref{fig:upward_comparison}) and the downward (Figure \ref{fig:downward_comparison}) 
continuations. The upward height is 300 m and the downward  is at 50 m.   Either in the upward continuation 
(Figure \ref{fig:upward_comparison}) or in the downward continuation (Figure \ref{fig:downward_comparison}), the 
continued gravity data using the fast equivalent layer proposed by \citet{siqueira-etal2017} 
(Figures \ref{fig:upward_comparison}a and \ref{fig:downward_comparison}a) are in close agreement with those produced 
by our modification of \citet{siqueira-etal2017} method (Figures \ref{fig:upward_comparison}b and 
\ref{fig:downward_comparison}b). The residuals (Figures \ref{fig:upward_comparison}c and \ref{fig:downward_comparison}c)
quantify this agreement since their means and standard deviations are close to zero in both continued gravity data using 
both methods.  All the continued gravity data shown here (Figures \ref{fig:upward_comparison} and 
\ref{fig:downward_comparison}) agree with the true ones (not shown). The most striking feature of these upward or 
the downward continuations concerns the total computation time. The computation time spent by our method is 
approximately $1\,500$ times faster than \citet{siqueira-etal2017} method.

### 2019-11-29 manuscript/tables.tex
split the tex files

\begin{table}[h]
	\label{tab:RAM-usage}
	\begin{center}
		\begin{tabular}{|l|c|c|c|}
			\hline
			\textbf{$N$} & \textbf{Full matrix $\mathbf{A}$} & \textbf{Single column of $\mathbf{A}$} 
			& \textbf{Full matrix $\mathbf{L}$}\\
			\hline 
			$100$ & 0.0763 & 0.0000763 & 0.0006104\\
			\hline
			$400$ & 1.22 & 0.0031 & 0.0248\\
			\hline
			$2\,500$ & 48 & 0.0191 & 0.1528\\
			\hline
			$10\,000$ & 763 & 0.00763 & 0.6104\\
			\hline
			$40\,000$ & 12\,207 & 0.305 & 2.4416 \\
			\hline
			$250\,000$ & 476\,837 & 1.907 & 15.3 \\
			\hline
			$500\,000$ & 1\,907\,349 & 3.815 & 30.518 \\
			\hline
			$1\,000\,000$ & 7\,629\,395 & 7.629 & 61.035 \\
			\hline
		\end{tabular}
		\caption{Comparison between the system RAM memory usage needed to store the full matrix $\mathbf{A}$ 
		(equations \ref{eq:predicted-data-vector} and \ref{eq:BTTB_A}), a single column of $\mathbf{A}$ and the 
		full matrix $\mathbf{L}$ (equation \ref{eq:DTF_G}). The quantities are represented in megabyte (MB), for 
		different total number of observations $N$. We consider that the elements of $\mathbf{A}$ and $\mathbf{L}$ 
		occupy, respectively, 8 and 16 bytes each in computer memory.}
	\end{center}
\end{table}

### 2019-11-29 manuscript/DFT_computations.tex

\append{Computations with the 2D DFT}

In the present Appendix, we deduce equation \ref{eq:DFT-system}
by using the row-ordered $vec$-operator (here designated simply as $vec$-operator).
This equation can be efficiently computed by using the two-dimensional 
fast Fourier Transform. 
This operator was implicitly used by \citet[][ p. 31]{jain1989} to 
show the relationship between Kronecker products and separable 
transformations. The $vec$-operator defined here 
transforms a matrix into a column vector by stacking its rows. It is important 
to stress that this operator is different from that defined by \citet{neudecker1969}, which 
is most commonly found in the literature and 
transforms a matrix into a column vector by stacking its columns.

Let $\mathbf{M}$ be an arbitrary $N \times M$ matrix given by:
\begin{equation}
\mathbf{M} = \begin{bmatrix}
\mathbf{m}^{\top}_{1} \\ 
\vdots \\
\mathbf{m}^{\top}_{N}
\end{bmatrix} \: ,
\label{eq:matrix-M}
\end{equation}
where $\mathbf{m}_{i}$, $i = 1, \dots, N$, are $M \times 1$ vectors containing 
the rows of $\mathbf{M}$.
The elements of this matrix can be rearranged into a column vector by using the
$vec$-operator \citep[][ p. 31]{jain1989} as follows:
\begin{equation}
vec \left( \mathbf{M} \right) = \begin{bmatrix}
\mathbf{m}_{1} \\
\vdots \\
\mathbf{m}_{N}
\end{bmatrix}_{NM \times 1} \: .
\label{eq:vec-operator}
\end{equation}
This rearrangement is known as lexicographic ordering \citep[][ p. 150]{jain1989}.

Two important properties of the $vec$-operator (equation \ref{eq:vec-operator}) 
are necessary to us. 
To define the first one, consider an 
$N \times M$ matrix $\mathbf{H}$ given by
\begin{equation}
\mathbf{H} = \mathbf{P} \circ \mathbf{Q} \: ,
\label{eq:matrix-H}
\end{equation}
where $\mathbf{P}$ and $\mathbf{Q}$ are arbitrary $N \times M$ matrices and 
``$\circ$" represents the Hadamard product \citep[][ p. 298]{horn_johnson1991}.
By applying the $vec$-operator to $\mathbf{H}$ (equation \ref{eq:matrix-H}), 
it can be shown that
\begin{equation}
vec \left( \mathbf{H} \right) = 
vec \left( \mathbf{P} \right) \circ vec \left( \mathbf{Q} \right) \: .
\label{eq:vec-matrix-H}
\end{equation}
To define the second important property of $vec$-operator, 
consider an $N \times M$ matrix $\mathbf{S}$ defined by 
the separable transformation \citet[][ p. 31]{jain1989}:
\begin{equation}
\mathbf{S} = \mathbf{P \, M \, Q} \: ,
\label{eq:matrix-S}
\end{equation}
where $\mathbf{P}$ and $\mathbf{Q}$ are arbitrary $N \times N$ and $M \times M$ 
matrices, respectively.
By implicitly applying the $vec$-operator to 
the $\mathbf{S}$ (equation \ref{eq:matrix-S}), 
\citet[][ p. 31]{jain1989} showed that:
\begin{equation}
vec \left( \mathbf{S} \right) = 
\left( \mathbf{P} \otimes \mathbf{Q}^{\top} \right) 
vec \left( \mathbf{M} \right) \: ,
\label{eq:vec-matrix-S}
\end{equation}
where ``$\otimes$" denotes the Kronecker product \citep{neudecker1969}.
It is important to stress the difference between equation \ref{eq:vec-matrix-S}
and that presented by \citet{neudecker1969}, which is more commonly found in 
the literature.
While that equation uses a $vec$-operator that transforms a matrix into a column 
vector by stacking its columns, equation \ref{eq:vec-matrix-S} 
uses the $vec$-operator defined by equation \ref{eq:vec-operator}, which 
transforms a matrix into a column vector by stacking its rows.

Now, let us deduce equation \ref{eq:DFT-system} by 
using the above-defined properties (equation \ref{eq:vec-matrix-H}
and \ref{eq:vec-matrix-S}).
We start calling attention to the right side of equation \ref{eq:vec-DFT-system}.
Consider that vector $\mathbf{w}$ (equation \ref{eq:vec-DFT-system}) 
is obtained by applying the $vec$-operator (equation \ref{eq:vec-operator}) to a matrix 
$\mathbf{W}$, whose two-dimensional DFT $\tilde{\mathbf{W}}$ is represented by the 
following separable transformation \citep[][ p. 146]{jain1989}:
\begin{equation}
\tilde{\mathbf{W}} = \mathbf{F}_{2Q} \, \mathbf{W} \, \mathbf{F}_{2P} \: ,
\label{eq:2D-DFT-W}
\end{equation}
where $\mathbf{F}_{2Q}$ and $\mathbf{F}_{2P}$ are the $2Q \times 2Q$ and $2P \times 2P$ 
unitary DFT matrices. 
Using equation \ref{eq:vec-matrix-S} and the symmetry of unitary DFT 
matrices, we rewrite the right side of equation \ref{eq:vec-DFT-system} 
as follows:
\begin{equation}
vec \left( \tilde{\mathbf{W}} \right) = 
\left( \mathbf{F}_{2Q} \otimes \mathbf{F}_{2P} \right) 
vec \left( \mathbf{W} \right) \: .
\label{eq:right_side_DFT_system_1}
\end{equation}
Similarly, consider that $\mathbf{v}$ (equation \ref{eq:vec-DFT-system}) 
is obtained by applying the $vec$-operator (equation \ref{eq:vec-operator}) to a matrix 
$\mathbf{V}$, whose two-dimensional DFT (equation \ref{eq:2D-DFT-W}) is 
represented by $\tilde{\mathbf{V}}$, and use
equation \ref{eq:vec-matrix-S} to rewrite the 
left side of equation \ref{eq:vec-DFT-system} as follows:
\begin{equation}
\boldsymbol{\Lambda} \, vec \left( \tilde{\mathbf{V}} \right) = 
\boldsymbol{\Lambda}
\left( \mathbf{F}_{2Q} \otimes \mathbf{F}_{2P} \right) 
vec \left( \mathbf{V} \right) \: .
\label{eq:left_side_DFT_system_1}
\end{equation}
Note that both sides of equation \ref{eq:left_side_DFT_system_1}
are defined as the product of the diagonal matrix $\boldsymbol{\Lambda}$ (equation \ref{eq:C-diagonalized}) 
and a vector. In this case, the matrix-vector product can be conveniently replaced by
\begin{equation}
\boldsymbol{\lambda} \circ vec \left( \tilde{\mathbf{V}} \right) = 
\boldsymbol{\lambda} \circ
\left( \mathbf{F}_{2Q} \otimes \mathbf{F}_{2P} \right) 
vec \left( \mathbf{V} \right) \: ,
\label{eq:left_side_DFT_system_2}
\end{equation}
where $\boldsymbol{\lambda}$ is a $4QP \times 1$ vector containing the diagonal of 
$\boldsymbol{\Lambda}$ (equation \ref{eq:C-diagonalized}).
Then, consider that $\boldsymbol{\lambda}$ is obtained by applying the $vec$-operator 
(equation \ref{eq:vec-operator}) to a $2Q \times 2P$ matrix $\mathbf{L}$ and use 
equations \ref{eq:vec-matrix-H} and \ref{eq:vec-matrix-S} to rewrite equation 
\ref{eq:left_side_DFT_system_2} as follows:
\begin{equation}
vec \left( \mathbf{L} \circ \tilde{\mathbf{V}} \right) = 
vec \left[ \mathbf{L} \circ 
\left( \mathbf{F}_{2Q} \, \mathbf{V} \, \mathbf{F}_{2P} \right) 
\right] \: .
\label{eq:left_side_DFT_system_3}
\end{equation}
Equations \ref{eq:2D-DFT-W}, \ref{eq:right_side_DFT_system_1} and 
\ref{eq:left_side_DFT_system_3} show that equation \ref{eq:vec-DFT-system}
is obtained by applying the $vec$-operator to 
\begin{equation}
\mathbf{L} \circ \left( \mathbf{F}_{2Q} \, \mathbf{V} \, \mathbf{F}_{2P} \right) = 
\mathbf{F}_{2Q} \, \mathbf{W} \, \mathbf{F}_{2P} \: .
\label{eq:DFT-system-preliminary}
\end{equation}
Finally, we premultiply both sides of equation \ref{eq:DFT-system-preliminary} by 
$\mathbf{F}_{2Q}^{\ast}$ and then postmultiply both sides of the result by 
$\mathbf{F}_{2P}^{\ast}$ to deduce equation \ref{eq:DFT-system}.

### 2019-11-29 manuscript/abstract.tex

\begin{abstract}

We have developed an efficient and very fast equivalent-layer technique for gravity data processing by modifying an 
iterative method grounded on excess mass constraint that does not require the solution of linear systems. 
Taking advantage of the symmetric Block-Toeplitz Toeplitz-block (BTTB) structure of the sensitivity matrix, that raises 
when regular grids of observation points and equivalent sources (point masses) are used to set up a fictitious 
equivalent layer, we have developed an algorithm which greatly reduces the number of flops and RAM memory necessary 
to estimate a 2D mass distribution over the equivalent layer. The structure of symmetric BTTB matrix consists of 
the elements of the first column of the sensitivity matrix, which in turn can be embedded into a symmetric 
Block-Circulant Circulant-Block (BCCB) matrix. Likewise, only the first column of the BCCB matrix is needed 
to reconstruct the full sensitivity matrix completely. From the first column of BCCB matrix, its eigenvalues 
can be calculated using the two-dimensional Fast Fourier Transform (2D FFT), which can be used to readily 
compute the matrix-vector product of the forward modeling in the fast equivalent-layer technique. As a result, 
our method is efficient to process very large datasets using either fine- or mid-grid meshes. The larger the 
dataset, the faster and more efficient our method becomes compared to the available equivalent-layer techniques. 
Synthetic tests demonstrate the ability of our method to satisfactorily upward- and downward-continuing the 
gravity data. For example, while our method requires 26:8 seconds to run one million of observations, the fast 
equivalent-layer technique required 48:3 seconds to run 22,500 observations. Test with real data from Caraj\'as, 
Brazil, shows its applicability to process very large dataset at low computational cost.

\end{abstract}

### 2019-11-29 manuscript/eigenvalues_C.tex

\append{The eigenvalues of $\mathbf{C}$}

In the present Appendix, we show how to efficiently compute matrix $\mathbf{L}$
(equations \ref{eq:left_side_DFT_system_3}, \ref{eq:DFT-system-preliminary} 
and \ref{eq:DFT-system}) by using only the first column of the BCCB matrix
$\mathbf{C}$ (equation \ref{eq:w_Cv}).

We need first premultiply both sides of equation \ref{eq:C-diagonalized}
by $\left(\mathbf{F}_{2Q} \otimes \mathbf{F}_{2P} \right)$ to obtain
\begin{equation}
\left(\mathbf{F}_{2Q} \otimes \mathbf{F}_{2P} \right) \mathbf{C} = 
\boldsymbol{\Lambda}
\left(\mathbf{F}_{2Q} \otimes \mathbf{F}_{2P} \right) \: .
\label{eq:C-diagonalized2}
\end{equation}
From equation \ref{eq:C-diagonalized2}, we can easily show that 
\citep[][ p. 77]{chan-jin2007}:
\begin{equation}
	\left(\mathbf{F}_{2Q} \otimes \mathbf{F}_{2P} \right) \, 
	\mathbf{c}_{0} = \frac{1}{\sqrt{4QP}} \, \boldsymbol{\lambda} \: ,
	\label{eq:DFT_C_column}
\end{equation}
where $\mathbf{c}_{0}$ is a $4QP \times 1$ vector representing the first column of 
$\mathbf{C}$ (equation \ref{eq:w_Cv}) and 
$\boldsymbol{\lambda}$ (equation \ref{eq:left_side_DFT_system_2}) is the $4QP \times 1$ 
vector that contains the diagonal of matrix $\boldsymbol{\Lambda}$ (equation \ref{eq:C-diagonalized}) 
and is obtained by applying the $vec$-operator (equation \ref{eq:vec-operator}) to matrix $\mathbf{L}$.
Now, let us conveniently consider that $\mathbf{c}_{0}$ is obtained by applying the $vec$-operator 
to a $2Q \times 2P$ matrix $\mathbf{G}$.
Using this matrix, the property of the $vec$-operator for separable transformations 
(equation \ref{eq:matrix-S}) and the symmetry of unitary DFT matrices, equation \ref{eq:DFT_C_column} 
can be rewritten as follows
\begin{eqnarray}
	\mathbf{F}_{2Q} \, \mathbf{G} \, \mathbf{F}_{2P} = 
	\frac{1}{\sqrt{4QP}} \, \mathbf{L} \: .
	\label{eq:DTF_G}
\end{eqnarray}
This equation shows that the eigenvalues of the BCCB matrix $\mathbf{C}$ 
(equation \ref{eq:w_Cv}), forming the rows of $\mathbf{L}$,
are obtained by computing the two-dimensional DFT of matrix $\mathbf{G}$,
which contains the elements forming the first column of the BCCB matrix 
$\mathbf{C}$ (equation \ref{eq:w_Cv}).

### 2019-11-29 manuscript/methodology.tex

@@
 by using the two-dimensional Fast Fourier Transform (2D FFT).
-This approach is a two-dimensional discrete convolution \citep[e.g.,][ p. 213]{vanloan1992}.
+This approach is a fast two-dimensional discrete convolution \citep[e.g.,][ p. 213]{vanloan1992}.
 
@@
 up to  $25$ millions. 
-These results shows that, while the running time of our method is $26.8$ seconds for 
-$N = 1\,000\,000$, the fast equivalent-layer technique spends $48.3$ seconds for $N = 22\,500$.
+These results shows that, while the running time of our method is $\approx 30.9$ seconds for 
+$N = 1\,000\,000$, the fast equivalent-layer technique spends $\approx 46.8$ seconds for $N = 22\,500$.

### 2019-11-29 manuscript/tables.tex

@@
-\begin{table}[h]
+\tabl{RAM-usage}{Comparison between the system RAM memory usage needed to store the full 
+matrix $\mathbf{A}$ (equations \ref{eq:predicted-data-vector} and \ref{eq:BTTB_A}), a single 
+column of $\mathbf{A}$ and the full matrix $\mathbf{L}$ (equation \ref{eq:DTF_G}). The quantities 
+are represented in megabyte (MB), for different total number of observations $N$. We consider 
+that the elements of $\mathbf{A}$ and $\mathbf{L}$ occupy, respectively, 8 and 16 bytes each 
+in computer memory.
 \label{tab:RAM-usage}
+}{
 	\begin{center}
-		\begin{tabular}{|l|c|c|c|}
+		\begin{tabular}[]{|l|c|c|c|}
 			\hline
@@
 		\end{tabular}
-		\caption{Comparison between the system RAM memory usage needed to store the full matrix $\mathbf{A}$ 
-		(equations \ref{eq:predicted-data-vector} and \ref{eq:BTTB_A}), a single column of $\mathbf{A}$ and the 
-		full matrix $\mathbf{L}$ (equation \ref{eq:DTF_G}). The quantities are represented in megabyte (MB), for 
-		different total number of observations $N$. We consider that the elements of $\mathbf{A}$ and $\mathbf{L}$ 
-		occupy, respectively, 8 and 16 bytes each in computer memory.}
 	\end{center} 
-\end{table}
+}

### 2019-11-29 manuscript/figures.tex
review the application to synthetic data

@@
 \renewcommand{\figdir}{Fig} % figure directory
+
+% Computational performance
 
@@
 	associated with the fast equivalent-layer technique \citep{siqueira-etal2017} and our method, 
-	for $N$ varying from $5\,000$ to $1\,000\,000$. All values are computed with $N^{it} = 50$ iterations.}
+	for $N$ varying from $5\,000$ to $1\,000\,000$. All values are computed with $N^{it} = 50$ 
+	iterations and $\kappa = 5$.}
 	\label{fig:float}
@@
 
+% Application to synthetic data
+
 \plot{synthetic_data}{width=13.5cm}{
-	{Noise-corrupted gravity data (in color map) produced by three simulated sources  whose horizontal projections 
-	are shown in black lines. The simulated sources are: two polygonal prisms, with density contrast of 
-	$0.35\,	\mathrm{g/cm^3}$ (upper-left body) and $0.4\, \mathrm{g/cm^3}$ (upper-right body), and a sphere with 
-	radius of $1\,000$ m with density contrast	of $-0.5\, \mathrm{g/cm^3}.$}
+	{Noise-corrupted gravity data (in color map) produced by three synthetic sources: 
+	a sphere with density contrast $-0.5\, \mathrm{g/cm^3}$ and 
+	two rectangular prisms with density contrasts $0.35\, \mathrm{g/cm^3}$ (upper-left body) and 
+	$0.4\, \mathrm{g/cm^3}$ (upper-right body). The black lines represent the horizontal projection 
+	of the sources.}
 	\label{fig:synthetic_data}
@@
 \plot{classic_fast_val}{width=8cm}{
-	{(a) Fitted gravity data produced by the fast equivalent-layer technique proposed by \citet{siqueira-etal2017}. 
-	(b) Gravity residuals, defined as the difference between the observed data in Figure \ref{fig:synthetic_data} and 
-	the predicted data in panel a, with their mean of $8.264e^{-7}$ and standard deviation of $0.0144$ mGal.}
+	{(a) Predicted data produced by the fast equivalent-layer technique \citep{siqueira-etal2017}. 
+	(b) Residuals between the simulated (Figure \ref{fig:synthetic_data}) and 
+	predicted (shown in a) data. The mean and standard deviation are $8.264 \times 10^{-7}$ and $0.0144$ mGal,
+	respectively.}
 	\label{fig:classic_fast_val}
@@
 \plot{bccb_fast_val}{width=8cm}{
-	{(a) Fitted gravity data produced by our modification of the fast equivalent layer \citep{siqueira-etal2017}. 
-	(b) Gravity residuals, defined as the difference between the observed data in Figure 4 and the predicted data in 
-	panel a, with their mean of $8.264e^{-7}$ and standard deviation of $0.0144$ mGal.}
+	{(a) Predicted data produced by our method. 
+	(b) Residuals between the simulated (Figure \ref{fig:synthetic_data}) and 
+	predicted (shown in a) data. The mean and standard deviation are $8.264 \times 10^{-7}$ and $0.0144$ mGal,
+	respectively.}
 	\label{fig:bccb_fast_val}
@@
 \plot{delta_comparison}{width=13.5cm}{
-	{Difference between the fitted gravity data produced by \citet{siqueira-etal2017} method 
-	(Figure \ref{fig:classic_fast_val}a) and by our modified form of this method (Figure \ref{fig:bccb_fast_val}a) 
-	that computes the forward modeling using the properties of BTTB and BCCB matrices (equation \ref{eq:w_Cv}).}
+	{Difference between the predicted data produced by the fast equivalent-layer technique 
+	(Figure \ref{fig:classic_fast_val}a) and by our method (Figure \ref{fig:bccb_fast_val}a).}
 	\label{fig:delta_comparison}
@@
 \plot{delta_rho}{width=13.5cm}{
-	{Difference between the estimated mass distribution within the equivalent layer produced by \citet{siqueira-etal2017}
-	method (Figure \ref{fig:classic_fast_val}) and by our modified form of this method (Figure \ref{fig:bccb_fast_val}) 
-	that computes the forward modeling using the properties of BTTB and BCCB matrices (equation \ref{eq:w_Cv}).}
+	{Difference between the estimated mass distribution within the equivalent layers produced by 
+	the fast equivalent-layer technique (not shown) and by our method (not shown).}
 	\label{fig:delta_rho}
@@
 \plot{upward_comparison}{width=5cm}{
-	{The upward-continued gravity data using: (a) the fast equivalent layer proposed by \citet{siqueira-etal2017} and 
-	(b) our modified form of \citet{siqueira-etal2017} method by using the properties of BTTB and BCCB matrices 
-	(equation \ref{eq:w_Cv}) to calculate the forward modeling.  (c) Residuals, defined as the difference between 
-	panels a and b with their mean of $-5.938e^{-18}$ and standard deviation of $8.701e^{-18}$.  The total computation 
-	times in the \citet{siqueira-etal2017} method and in our approach are $7.62026$ and $0.00834$ seconds, respectively.}
+	{The upward-continued gravity data obtained with (a) the fast equivalent-layer technique 
+	\citep{siqueira-etal2017} and (b) our method. (c) Residuals between 
+	panels a and b with mean $-5.938 \times 10^{-18}$ mGal and standard deviation $8.701 \times 10^{-18}$ mGal. 
+	The total computation times spent by the fast equivalent-layer technique and our method 
+	are $7.62026$ and $0.00834$ seconds, respectively.}
 	\label{fig:upward_comparison}
@@
 \plot{downward_comparison}{width=5cm}{
-	{The downward-continued gravity data using: (a) the fast equivalent layer proposed by \citet{siqueira-etal2017} and 
-	(b) our modified form of \citet{siqueira-etal2017} method by using the properties of BTTB and BCCB matrices 
-	(equation \ref{eq:w_Cv}) to calculate the forward modeling.  (c) Residuals, defined as the difference between panels 
-	a and b with their mean of $5.914e^{-18}$ and standard deviation of $9.014e^{-18}$.  The total computation times in 
-	the \citet{siqueira-etal2017} method and in our approach are $7.59654$ and $0.00547$ seconds, respectively.}
+	{The downward-continued gravity data obtained with (a) the fast equivalent-layer technique 
+	\citep{siqueira-etal2017} and (b) our method. (c) Residuals between 
+	panels a and b with mean $5.914 \times 10^{-18}$ mGal and standard deviation $9.014 \times 10^{-18}$ mGal. 
+	The total computation times spent by the fast equivalent-layer technique and our method 
+	are $7.59654$ and $0.00547$ seconds, respectively.}
 	\label{fig:downward_comparison}
 }
+
+% Application to field data
 

### 2019-11-29 manuscript/synthetic_data.tex
review the application to synthetic data

@@
 
-In this section, we  investigate the effectiveness of using the properties of BTTB and BCCB matrices 
-(equation \ref{eq:w_Cv})  to solve, at each iteration, the forward modeling (the matrix-vector product 
-$\mathbf{A} \hat{\mathbf{p}}^k$)  required in the fast equivalent-layer method proposed by \citet{siqueira-etal2017}. 
 We simulated three sources whose horizontal projections are shown in Figure \ref{fig:synthetic_data} as black lines. 
-These sources are two vertical prisms with density contrasts of $0.35\, \mathrm{g/cm^3}$ (upper-left prism) and 
-$0.4\, \mathrm{g/cm^3}$ (upper-right prism) and a sphere with radius of $1\,000$ m with density contrast of 
-$-0.5\, \mathrm{g/cm^3}$. Figure \ref{fig:synthetic_data}  shows the vertical component of gravity field generated 
-by these sources contaminated with additive pseudorandom Gaussian noise with zero mean and standard deviation of 
-$0.015$ mGal.
+These sources are a sphere with radius of $1\,000$ m with density contrast of 
+$-0.5\, \mathrm{g/cm^3}$ and two rectangular prisms with density contrasts of $0.35\, \mathrm{g/cm^3}$ (upper-left prism) 
+and $0.4\, \mathrm{g/cm^3}$ (upper-right prism). 
+Figure \ref{fig:synthetic_data} shows the gravity disturbance (vertical component of gravitational attraction) 
+produced by these sources. The synthetic data are contaminated with additive pseudo-random Gaussian 
+noise with zero mean and standard deviation of $0.015$ mGal.
+We created $10\,000$ observation points regularly spaced on a grid of $100 \times 100$, on a 
+horizontal plane at $-100$ m, and set a grid of equivalent point masses, each one directly beneath each 
+observation point, located at depth $300$ m.
 
-The advantage of using the structures of BTTB and BCCB matrices to compute forward modeling in  the fast-equivalent 
-layer method \cite[]{siqueira-etal2017} is grounded on the use of regular grids of data and equivalent sources. 
-Hence, we created $10\,000$ observation points regularly spaced in a grid of $100 \times 100$ at $100$ m height. 
-We also set a grid of equivalent point masses, each one directly beneath each observation points, located at 
-$300$ m deep.  Figures \ref{fig:classic_fast_val}a and \ref{fig:bccb_fast_val}a  show the fitted gravity data 
-obtained, respectively, by the fast equivalent layer method and by our modified form of this method that computes 
-the forward modeling using equation \ref{eq:w_Cv}. The corresponding residuals (Figures \ref{fig:classic_fast_val}b 
-and \ref{fig:bccb_fast_val}b), defined as the difference between the observed (Figure  \ref{fig:synthetic_data}) 
-and fitted gravity data (Figures \ref{fig:classic_fast_val}a and \ref{fig:bccb_fast_val}a), show means close to 
-zero  and standard deviations of 0.0144 mGal.  Therefore, Figures \ref{fig:classic_fast_val} and 
-\ref{fig:bccb_fast_val} show that \citet{siqueira-etal2017} method and our modified version of this method produced 
-virtually the same results. This excellent agreement is confirmed in Figures \ref{fig:delta_comparison} and 
-\ref{fig:delta_rho} which shows that there are virtually no differences, respectively,  in the fitted data presented 
-in Figures \ref{fig:classic_fast_val}b and \ref{fig:bccb_fast_val}b and in the estimated mass distributions within 
-the equivalent layers (not shown) yielded by both \citet{siqueira-etal2017} method and our modification of this method. 
-These results (Figures \ref{fig:delta_comparison} and \ref{fig:delta_rho})  show that computing the matrix-vector 
-product ($\mathbf{A} \hat{\mathbf{p}}^k$), required in the forward modeling, by means of embedding the BTTB matrix into 
-a BCCB matrix (equation \ref{eq:w_Cv}) yields practically the same result as the one produced by computing this 
-matrix-vector product  with a full matrix $\mathbf{A}$ as used in \citet{siqueira-etal2017}.
+Figures \ref{fig:classic_fast_val}a and \ref{fig:bccb_fast_val}a show the data fits 
+obtained, respectively, by the fast equivalent-layer technique \citep{siqueira-etal2017} 
+and by our method. The corresponding residuals (Figures \ref{fig:classic_fast_val}b 
+and \ref{fig:bccb_fast_val}b), defined as the difference between the simulated (Figure \ref{fig:synthetic_data}) 
+and predicted data (Figures \ref{fig:classic_fast_val}a and \ref{fig:bccb_fast_val}a), show means close to 
+zero and standard deviations of $0.0144$ mGal, indicating that both methods produced virtually the same results.
+This excellent agreement is confirmed by Figures \ref{fig:delta_comparison} and 
+\ref{fig:delta_rho}. They show that there are virtually no differences between, respectively, 
+the predicted data (Figures \ref{fig:classic_fast_val}b and \ref{fig:bccb_fast_val}b) and the 
+estimated mass distributions within the equivalent layers (not shown) yielded by 
+both methods.
 
-We perform two forms of processing the gravity data (Figure \ref{fig:synthetic_data}) through the equivalent layer 
-technique: the upward (Figure \ref{fig:upward_comparison}) and the downward (Figure \ref{fig:downward_comparison}) 
-continuations. The upward height is 300 m and the downward  is at 50 m.   Either in the upward continuation 
-(Figure \ref{fig:upward_comparison}) or in the downward continuation (Figure \ref{fig:downward_comparison}), the 
-continued gravity data using the fast equivalent layer proposed by \citet{siqueira-etal2017} 
-(Figures \ref{fig:upward_comparison}a and \ref{fig:downward_comparison}a) are in close agreement with those produced 
-by our modification of \citet{siqueira-etal2017} method (Figures \ref{fig:upward_comparison}b and 
+We performed the upward and downward continuations of the simulated gravity data 
+(Figure \ref{fig:synthetic_data}) by using the estimated equivalent layers.
+The results show that the continued gravity data obtained by using the 
+fast equivalent-layer (Figures \ref{fig:upward_comparison}a and \ref{fig:downward_comparison}a) 
+are in close agreement with those produced by our method (Figures \ref{fig:upward_comparison}b and 
 \ref{fig:downward_comparison}b). The residuals (Figures \ref{fig:upward_comparison}c and \ref{fig:downward_comparison}c)
-quantify this agreement since their means and standard deviations are close to zero in both continued gravity data using 
-both methods.  All the continued gravity data shown here (Figures \ref{fig:upward_comparison} and 
-\ref{fig:downward_comparison}) agree with the true ones (not shown). The most striking feature of these upward or 
-the downward continuations concerns the total computation time. The computation time spent by our method is 
-approximately $1\,500$ times faster than \citet{siqueira-etal2017} method.
+quantify this agreement since their means and standard deviations are close to zero.
+
+The continued gravity data obtained with both methods agree with the true ones (not shown).
+The difference is that the computation time spent by our method is 
+approximately $1\,500$ times faster than that of the fast-equivalent-layer 
+technique \citep{siqueira-etal2017}.

### 2019-11-29 manuscript/manuscript.tex
review the application to synthetic data
change the title

@@
 
-\title{Fast equivalent-layer technique for gravity data processing with BTTB matrix systems}
+\title{Convolutional equivalent layer for gravity data processing}
 
@@
 
-%\address{
-%\footnotemark[1]Observat\'{o}rio Nacional, \\
-%Department of Geophysics, \\
-%Rio de Janeiro, Brazil}
-%\author{Diego Takahashi\footnotemark[1], Vanderlei C. Oliveira{ }Jr.\footnotemark[1] and 
-%Val{\'e}ria C. F. Barbosa\footnotemark[1]}
-%
-%\footer{vanderlei@on.br}
+\address{
+\footnotemark[2]Observat\'{o}rio Nacional, Department of Geophysics, Rio de Janeiro, Brazil\\
+\footnotemark[1] Corresponding author: diego.takahashi@gmail.com
+}
+\author{Diego Takahashi\footnotemark[2]\footnotemark[1], Vanderlei C. Oliveira{ }Jr.\footnotemark[2] and 
+Val{\'e}ria C. F. Barbosa\footnotemark[2]}
+
+%\footer{Example}
 %\lefthead{Takahashi, Oliveira{ }Jr. \& Barbosa}
-%\righthead{Fast equivalent-layer for BTTB systems}
+\righthead{Convolutional equivalent layer}
 
@@
 \include{conclusions}
-\include{aknowledgements}
+%\include{aknowledgements}
 

### 2019-11-29 manuscript/synthetic_data.tex
review application to field data

@@
-\section{Synthetic tests}
+\section{Application to synthetic data}
 
@@
 noise with zero mean and standard deviation of $0.015$ mGal.
-We created $10\,000$ observation points regularly spaced on a grid of $100 \times 100$, on a 
-horizontal plane at $-100$ m, and set a grid of equivalent point masses, each one directly beneath each 
-observation point, located at depth $300$ m.
+We created $N = 10\,000$ observation points regularly spaced on a grid of $100 \times 100$, on a 
+horizontal plane at $z_{1} = -100$ m, and set a grid of equivalent point masses, each one directly beneath each 
+observation point, located at depth $z_{0} = 300$ m.
 
@@
 and \ref{fig:bccb_fast_val}b), defined as the difference between the simulated (Figure \ref{fig:synthetic_data}) 
-and predicted data (Figures \ref{fig:classic_fast_val}a and \ref{fig:bccb_fast_val}a), show means close to 
-zero and standard deviations of $0.0144$ mGal, indicating that both methods produced virtually the same results.
+and predicted data (Figures \ref{fig:classic_fast_val}a and \ref{fig:bccb_fast_val}a) after $N^{it} = XXXX$ iterations, 
+show means close to zero and standard deviations of $0.0144$ mGal, indicating that both methods produced virtually 
+the same results.
 This excellent agreement is confirmed by Figures \ref{fig:delta_comparison} and 
@@
 The difference is that the computation time spent by our method is 
-approximately $1\,500$ times faster than that of the fast-equivalent-layer 
+approximately $1\,500$ times smaller than that of the fast-equivalent-layer 
 technique \citep{siqueira-etal2017}.

### 2019-11-29 manuscript/conclusions.tex
review the conclusions

@@
 \section{Conclusions}
-By exploring the BTTB structure of the sensitivity matrix in the gravity data processing, we have proposed a new 
-efficient approach for calculating the gravity-data forward modeling required in the iterative fast equivalent-layer 
-technique grounded on excess mass constraint that does not demand  the solution of linear systems.  Its efficiency 
-requires the use of regular grids of observations and equivalent sources (point masses). Our algorithm greatly reduces 
-the number of  flops necessary to estimate a 2D mass distribution within the equivalent layer that fits the observed 
-gravity data. For example, when processing one million observations the number of flops is reduced in 104 times. 
-Traditionally, such amount of data impractically requires 7.6 Terabytes of RAM memory to handle the full sensitivity 
-matrix. Rather, in our method, this matrix takes 61.035 Megabytes of RAM memory only.
 
-Our method takes advantage of the symmetric BTTB system that arises when processing a harmonic function and considering 
-that either the observations or the sources of the interpretative model (point of masses over the equivalent layer) 
-are distributed on regular grids. Symmetric BTTB matrices can be stored by using only a single column and can be 
-embedded into a symmetric BCCB matrix, which in turn also only needs a single column. 
-This means that only a single column  of the sensitivity matrix needs to be calculated.  Once the jth column of the 
-sensitivity matrix represents the influence of the jth source (jth point mass) has on the predicted data, our method 
-only requires a single point mass that set up the equivalent layer to compute the gravity-data forward modeling.
-
-Using the 2D FFT, it is possible to calculate the eigenvalues of BCCB matrices which can
-be used to compute a matrix-vector product (gravity-data forward modeling) in a very low computational cost. 
-We have successfully applied the proposed method to upward (or downward) synthetic gravity data. Testing on field 
-data from the Caraj{\'a}s Province, north of Brazil, confirms the potential of our approach in upward-continuing 
-gravity data with $250\,000$ observations in  about 0.2 seconds.  Our method allows, in future research, applying 
-the equivalent layer-technique for processing and interpreting massive data set such as collected in continental 
-and global scales studies.
+We show that the sensitivity matrix associated with the equivalent layer technique 
+for processing gravity data has a BTTB structure when the following conditions are
+satisfied: (i) the observed data are disposed at a regular grid on a horizontal plane
+and (ii) the equivalent sources (point masses) are placed at a constant depth, one directly 
+beneath each observation point.
+By exploring the BTTB structure of the sensitivity matrix, we formulate the forward modeling as a 
+fast FFT convolution that requires only one column of the sensitivity matrix 
+and propose an efficient approach for optimizing the computational time of an 
+iterative method called fast equivalent-layer technique.
+This iterative method is grounded on the excess mass constraint and does not demand the solution of 
+linear systems.
+Our approach greatly reduces the number of flops and the RAM memory necessary to estimate a 
+2D mass distribution within a planar equivalent layer that fits the observed gravity data. 
+For example, the number of flops of the fast equivalent-layer technique is reduced by two orders 
+of magnitude when processing one million of observations by using our approach. 
+Traditionally, such amount of data impractically requires $7.6$ Terabytes of RAM memory to handle 
+the full sensitivity matrix whereas our method takes only $61.035$ Megabytes.
+This drastic reduction comes from the fact that, by using only one column of the sensitivity matrix,
+our method is able to compute the gravity data produced by the whole equivalent layer with a single 
+point mass via FFT.
+We have successfully applied the proposed method to compute the upward/downward continuation of 
+synthetic gravity data. Application to field data over the Caraj{\'a}s Province, north of Brazil, 
+confirms the potential of our approach in processing a gravity data set with $250\,000$ observations 
+in approximately $0.2$ seconds. Further studies need to be carried out in order to apply our 
+method to process other potential-field data.

### 2019-12-03 manuscript/introduction.tex

@@
 
-The equivalent layer is a well-known technique for processing potential-field data in applied geophysics since 1960. 
-It comes from potential theory as a mathematical solution of the Laplace's equation, in the region above the sources, 
-by using the Dirichlet boundary condition \citep{kellogg1929}.
-This theory states that any potential-field data produced by an arbitrary 3D physical-property distribution can be 
-exactly reproduced by a fictitious layer located at any depth and having a continuous 2D physical-property  
-distribution. In practical situations, the layer is approximated by a finite set of sources (e.g., point masses or 
-dipoles) and their physical properties are estimated by solving a linear system of equations that yield an 
+The equivalent layer is a well-known technique for processing potential-field data in applied geophysics 
+since the 60's. 
+It comes from potential theory as a mathematical solution of the Laplace's equation, in the region above the
+sources, by using the Dirichlet boundary condition \citep{kellogg1929}.
+This theory states that any potential-field data produced by an arbitrary 3D physical-property distribution can 
+be exactly reproduced by a fictitious layer located at any depth and having a continuous 2D physical-property  
+distribution. In practical situations, the layer is approximated by a finite set of sources (e.g., point masses 
+or dipoles) and their physical properties are estimated by solving a linear system of equations that yield an 
 acceptable potential-field data fit. These fictitious sources are called equivalent sources.
 
-Many previous works have used the equivalent layer as a processing technique of potential-field data. 
-\citet{dampney1969} used the equivalent-layer technique for gridding and for computing the upward 
-continuation of the potential-field data. \citet{cordell1992} and \citet{mendonca-silva1994} used it for 
-interpolating and gridding potential-field data. \citet{emilia1973}, \citet{hansen-miyazaki1984} and 
-\citet{li-oldenburg2010} used it for upward continuation of the potential-field data. \citet{silva1986}, 
-\citet{leao-silva1989}, \citet{guspi-novara2009}, and \citet{oliveirajr-etal2013} used it for reducing the 
-magnetic data to the pole. \citet{boggs-dransfield2004} used it for combining multiple data sets and 
-\citet{barnes-lumley2011} for gradient-data processing.
+Many previous works have used the equivalent layer to perform different potential-field data 
+transformations such as gridding \citep[e.g.,][]{dampney1969, cordell1992, mendonca-silva1994},
+upward/downward continuation \citep[e.g.,][]{emilia1973, hansen-miyazaki1984, 
+li-oldenburg2010}, reduction to the pole \citep[e.g.,]{silva1986, leao-silva1989, guspi-novara2009, 
+oliveirajr-etal2013}, combining multiple data sets \citep[e.g.,][]{boggs-dransfield2004} and 
+gradient data processing \citep[e.g.,][]{barnes-lumley2011}.
 
-The classic equivalent-layer formulation consists in estimating the physical-property distribution within a layer, 
-composed by a set of equivalent sources, by solving a linear system of equations formed by harmonic functions 
-(e.g.,the inverse of the distance between the observation point and the equivalent source). When these observation 
-points and equivalent sources are regularly spaced, a Toeplitz system arises. Toeplitz systems are well-known in 
-many branches of science as in (1) mathematics, for solving partial and ordinary diferential equations 
-\citep[e.g.,][]{lin-etal2003}; (2) image processing \citep[e.g.,][]{chan-etal1999} and; (3) computational 
-neuroscience \citep[e.g.,][]{wray-green1994}. \citet{jin2003} and \citet{chan-jin2007} give many examples of 
-applications for Toeplitz systems.
+Although the use of the equivalent-layer technique increased over the last decades, one of the biggest 
+problems is still its high computational cost for processing large-data sets. 
+This problem propelled several studies to improve the computational efficiency of the 
+equivalent layer technique. 
+\citet{leao-silva1989} developed a fast method for processing a regular grid of potential-field data.
+The method consists in estimating an equivalent layer which exactly reproduces the potential-field data 
+within a small data window. The data window is shifted over the whole gridded data set in a procedure similar 
+to a discrete convolution. 
+The equivalent layer extends beyond the moving-data window and is located at a depth between two and six 
+times the grid spacing of the observations. For each data window, the equivalent layer is estimated by 
+solving an underdetermined linear system.
+After estimating an equivalent layer, the transformed-potential field is computed only at the center of the 
+moving-data window. The use of a small moving-data window greatly reduces the total number of floating-point 
+operations (\textit{flops}) and RAM memory storage. The computational efficiency of this method relies on the 
+strategy of constructing the equivalent layer by successively solving small linear systems instead of solving 
+just one large linear system for the entire equivalent layer. 
+\citet{mendonca-silva1994} also followed the strategy of solving successive small linear systems for 
+constructing an equivalent layer. Their method is based on the  equivalent-data concept, which 
+consists in determining a subset of all potential-field data (named equivalent-data set), such that the 
+interpolating surface that fits the chosen subset also automatically fits all remaining data. 
+The equivalent data set is obtained by iteratively introducing the potential-field observation with the 
+greatest residual in the preceding iteration. By applying to the interpolation problem, the method is 
+optimized by approximating dot products by the discrete form of an analytic integration that can be 
+evaluated with less computational effort. According to the authors, the equivalent-data set is usually 
+smaller than the total number of potential-field observations, leading to computational savings. 
+The authors also pointed out that the computational efficiency of the method depends on the number of 
+equivalent data. If the potential-field anomaly is nonsmooth, the number of equivalent data can be large and 
+the method will be less efficient than the classical approach.
 
-In potential-field methods, the properties of Toeplitz system were used for downward continuation 
-\citep{zhang-etal2016} and for 3D gravity-data inversion using a 2D multilayer model \citep{zhang-wong2015}. 
-In the particular case of gravity data, the kernel generates a linear system with a matrix known as symmetric 
-Block-Toeplitz Toeplitz-Block (BTTB).
+By following a different strategy, \citet{li-oldenburg2010} developed a rapid method that transforms the 
+dense sensitivity matrix associated with the linear system into a sparse one by using a wavelet technique. 
+After obtaining a sparse representation of the sensitivity matrix, those authors estimate the 
+physical-property distribution within the equivalent layer by using an overdetermined formulation. 
+Those authors pointed out that, given the sparse representation, their method reduces the computational time 
+required for solving the linear system by as many as two orders of magnitude if compared with the same 
+formulation using a dense matrix. 
+\citet{barnes-lumley2011} followed a similar strategy and transformed the dense sensitivity matrix 
+into a sparse one. However, differently from \citet{li-oldenburg2010}, their method operates in the space 
+domain by grouping equivalent sources far from an observation point into blocks with average physical 
+property. This procedure aims at reducing the memory storage and achieving computational efficiency 
+by solving the transformed linear system with a weighted-least-squares conjugate-gradient algorithm.
+Notice that, instead of constructing the equivalent layer by solving successive small linear systems, 
+these last two methods first transform the large linear system into a sparse one and then take advantage of 
+this sparseness.
 
-INCLUIR \citet{qiang_etal2019}
+\citet{oliveirajr-etal2013} developed a fast method based on the reparameterization of the physical-property 
+distribution within the equivalent layer. Those authors divided the equivalent layer into a regular grid of 
+equivalent-source windows inside which the physical-property distribution is described by bivariate 
+polynomial functions. By using this polynomial representation, the inverse problem for estimating the 
+equivalent layer is posed in the space of the total number of polynomial coefficients within all 
+equivalent-source windows instead of in the space of the total number of equivalent sources. According to 
+\citet{oliveirajr-etal2013}, the computational efficiency of their method relies on the fact that the total 
+number of polynomial coefficients needed to describe the physical property distribution within the 
+equivalent layer is generally much smaller than the number of equivalent sources, leading to a very smaller 
+linear system to be solved. Those authors could verify that the total number of \textit{flops} needed for 
+building and solving the linear inverse problem of estimating the total number of polynomial coefficients can 
+be reduced by as many as three and four orders of magnitude, respectively, if compared with the same inverse 
+problem of estimating the physical property of each equivalent source via Cholesky decomposition.
 
-A wide variety of applications in mathematics and engineers that fall into Toeplitz systems propelled the development 
-of a large variety of  methods for solving them. 
-%Direct methods were conceived by \citet{levinson1946} and by \citet{trench1964}. Currently the conjugate gradient 
-%is used in most cases.
-\citet{grenander-szego1984} noticed that a circulant matrix can be diagonalized by taking the Fast Fourier Trasform 
-(FFT) of its first column, making it possible to calculate the matrix-vector product and solve the system with 
-low computational cost \citep{strang-aarikka1986, olkin1986}. \citet{chan-jin2007} show some preconditioners to embedd 
-the Toeplitz and BTTB matrices into, respectively, circulant matrices and Block-Circulant Circulant-Block (BCCB) by 
-solving the system applying the conjugate gradient method.
+There is another class of methods that iteratively estimates the physical-property distribution within 
+the equivalent layer without solving linear systems.
+The method presented by \citet{cordell1992} and after generalized by \citet{guspi-novara2009} updates 
+the physical property of the sources, which are located below each potential-field data, using a 
+procedure that removes the maximum residual between the observed and predicted data.
+\citet{xia-sprowl1991} and \citet{xia-etal1993} developed 
+fast iterative schemes for updating the physical-property distribution
+within the equivalent layer in the wavenumber and space domains, respectively.
+Grounded on excess mass constraint, \cite{siqueira-etal2017} developed an iterative scheme 
+starting with a mass distribution within the equivalent layer that is proportional to observed gravity data.
+Then, their method iteratively adds mass corrections that are proportional to the gravity residuals.
+The total number of \textit{flops} required by these iterative methods for estimating the physical-property 
+distribution within the equivalent layer depends on the total number of iterations, however this number is 
+generally much smaller than the total number of \textit{flops} required to solve a large-scaled linear system. 
+Generally, the most computational expensive step in each iteration of these methods is the forward problem 
+of calculating the potential-field data produced by the equivalent layer.
 
-Although the use of the equivalent-layer technique increased over the last decades, one of the biggest problem is still 
-its high computational cost for processing large-data sets. \citet{siqueira-etal2017} developed a computationally 
-efficient scheme for processing gravity data. This scheme does not solve a linear system, instead uses an iterative 
-process that corrects the physical-property distribution over the equivalent layer by adding mass corrections that 
-are proportional to the gravity residual. Although efficient, the method presented by \citet{siqueira-etal2017} requires, 
-at each iteration,  the full computation of the forward problem to guarantee the convergence of the algorithm. The 
-time spent on forward modeling accounts for most of the total computation time of their method.
-
-We propose the use of BTTB and BCCB matrices properties to efficiently  solve the forward modeling in method of 
-\citet{siqueira-etal2017}, resulting in a very faster parameter estimation and the possibility to use very large 
-datasets. Here, we show how the system memory (RAM) usage can be drastically decreased by calculating only the 
-first column of the BTTB matrix and embedding into a BCCB matrix. Using the Szeg\"{o} theorema  combined with 
-\cite{strang-aarikka1986}, the matrix-vector product can be accomplished with very low cost, reducing in some 
-orders of magnitude the number of operations required to complete the process. We present synthetic tests to validate 
-our proposal and real field data from Caraj\'as, Brazil to demonstrate its applicability.
+In the present work, we show that the sensitivity matrix associated with a planar equivalent layer 
+of point masses has a very well-defined structure called Block-Toeplitz Toeplitz-Block (BTTB) for 
+the case in which (i) the observed gravity data is located on a regularly spaced grid at constant 
+height and (ii) there is one point mass directly beneath each observation point.
+By using this property, we propose an efficient algorithm based on fast FFT convolution 
+\citep[e.g.,][ p. 207]{vanloan1992} for computing the forward problem at each iteration of 
+the fast equivalent-layer technique proposed by \citet{siqueira-etal2017}.
+Our method uses the gravitational effect produced by a single point mass to compute the 
+effect produced by the whole equivalent layer, which results in a drastic reduction 
+not only in the number of flops, but also in the RAM memory usage of the fast equivalent-layer technique.
+This technique have been successfully used in potential-field methods for 
+3D gravity inversion \citep{zhang-wong2015}, downward continuation 
+\citep{zhang-etal2016} and 3D magnetic modeling \citep{qiang_etal2019}.
+Tests with synthetic and field data illustrate the good performance of our method in processing 
+large gravity data sets.

### 2019-12-03 manuscript/field_data.tex

@@
-\section{Field data test}
-Test with real data are conducted with the gravity data from Caraj\'as, north of Brazil, were provided by the  Geological 
-Survey of Brazil (CPRM). The real aerogravimetric data were collected in 113 flight lines along north–south direction 
-with flight line spacing  of 3 km and tie lines along east–west direction at 12 km.
+\section{Application to field data}
 
-This airborne gravity survey was divided in two different areas, collected in different times, having samples spacing of 
-$7.65$ m and $15.21$ m, totalizing  $4\,353\,428$ observation points. The height of the flight was fixed at 900 m. 
-The gravity data (Figure \ref{fig:carajas_real_data}) were gridded into a regularly spaced dataset of $250\,000$ 
-observation points ($500 \times 500$) with a grid spacing of $716.9311$ km north-south and $781.7387$ km east-west.
+We applied our method to airborne gravity data from Caraj{\'a}s, north of Brazil, which were provided by 
+the Geological Survey of Brazil (CPRM). The data were collected along $113$ north–south flight lines separated 
+by $3$ km and XXXXXX east–west tie lines separated by $12$ km.
+This data set was divided in two different areas, collected in different times, having samples spacing of 
+$7.65$ m and $15.21$ m along the lines, totalizing  $4\,353\,428$ observation points at a fixed height 
+of $900$ m ($z_{1} = -900$ m). 
+The gravity data were interpolated (Figure \ref{fig:carajas_real_data}) into a regularly spaced grid of 
+$500 \times 500$ observation points ($N = 250\,000$) with a grid spacing of $716.9311$ m north-south and 
+$781.7387$ m east-west.
 
-To apply our modification of the fast equivalent-layer method \cite[]{siqueira-etal2017}  that computes the forward 
-modeling using the properties of BTTB and BCCB matrices (equation \ref{eq:w_Cv}), we set an equivalent layer at 300 m deep. 
-Figure \ref{fig:carajas_gz_predito_val}a shows the fitted gravity data after 50 iterations by applying our method. 
+To apply our method, we set an equivalent layer at $z_{0} = -300$ m. 
+Figure \ref{fig:carajas_gz_predito_val}a shows the predicted data obtained with our method after 
+$N^{it} = 50$ iterations.
 The residuals (Figure \ref{fig:carajas_gz_predito_val}b), defined as the difference between the observed 
-(Figure \ref{fig:carajas_real_data}) and the predicted (Figure \ref{fig:carajas_gz_predito_val}a) data, show an 
-acceptable data fitting because they have a mean close to zero ($0.0003$ mGal) and a small standard deviation of 
-$0.105$ mGal which corresponds to approximately $0.1$ \% of the amplitude of the gravity data.
-
-These small residuals indicate that our method yielded an estimated mass distribution (not shown) that can be used in 
-the data processing. We perform upward-continuation of the real gravity data (Figure \ref{fig:carajas_real_data}) at 
-a constant height of $5\,000$ m over the real data. The upward-continued gravity data 
-(Figure \ref{fig:up2000_carajas_500x500}) seem a reasonable processing because of the attenuation of the short 
-wavelenghts. By using our approach, the processing of the $250\,000$ observations was extremely fast and took 
+(Figure \ref{fig:carajas_real_data}) and predicted (Figure \ref{fig:carajas_gz_predito_val}a) data, show a 
+very good data fit with mean close to zero ($0.0003$ mGal) and small standard deviation ($0.105$ mGal), 
+which corresponds to approximately $0.1$ \% of the maximum amplitude of the gravity data.
+By using the estimated mass distribution (not shown), we performed an upward-continuation of the 
+observed gravity data to a horizontal plane located $5\,000$ m above. Figure \ref{fig:up2000_carajas_500x500} 
+shows a very consistent upward-continued gravity data, with a clear attenuation of the short 
+wave-lengths. By using our approach, the processing of the $250\,000$ observations took only 
 $0.216$ seconds.

### 2019-12-03 manuscript/figures.tex

@@
 \plot{synthetic_data}{width=13.5cm}{
-	{Noise-corrupted gravity data (in color map) produced by three synthetic sources: 
+	{Application to synthetic data. 
+	Noise-corrupted gravity data (in color map) produced by three synthetic sources: 
 	a sphere with density contrast $-0.5\, \mathrm{g/cm^3}$ and 
@@
 \plot{classic_fast_val}{width=8cm}{
-	{(a) Predicted data produced by the fast equivalent-layer technique \citep{siqueira-etal2017}. 
+	{Application to synthetic data. 
+	(a) Predicted data produced by the fast equivalent-layer technique \citep{siqueira-etal2017}. 
 	(b) Residuals between the simulated (Figure \ref{fig:synthetic_data}) and 
@@
 \plot{bccb_fast_val}{width=8cm}{
-	{(a) Predicted data produced by our method. 
+	{Application to synthetic data. 
+	(a) Predicted data produced by our method. 
 	(b) Residuals between the simulated (Figure \ref{fig:synthetic_data}) and 
@@
 \plot{delta_comparison}{width=13.5cm}{
-	{Difference between the predicted data produced by the fast equivalent-layer technique 
+	{Application to synthetic data. 
+	Difference between the predicted data produced by the fast equivalent-layer technique 
 	(Figure \ref{fig:classic_fast_val}a) and by our method (Figure \ref{fig:bccb_fast_val}a).}
@@
 \plot{delta_rho}{width=13.5cm}{
-	{Difference between the estimated mass distribution within the equivalent layers produced by 
+	{Application to synthetic data. 
+	Difference between the estimated mass distribution within the equivalent layers produced by 
 	the fast equivalent-layer technique (not shown) and by our method (not shown).}
@@
 \plot{upward_comparison}{width=5cm}{
-	{The upward-continued gravity data obtained with (a) the fast equivalent-layer technique 
+	{Application to synthetic data. 
+	The upward-continued gravity data obtained with (a) the fast equivalent-layer technique 
 	\citep{siqueira-etal2017} and (b) our method. (c) Residuals between 
@@
 \plot{downward_comparison}{width=5cm}{
-	{The downward-continued gravity data obtained with (a) the fast equivalent-layer technique 
+	{Application to synthetic data. 
+	The downward-continued gravity data obtained with (a) the fast equivalent-layer technique 
 	\citep{siqueira-etal2017} and (b) our method. (c) Residuals between 
@@
 \plot{carajas_real_data}{width=13.5cm}{
-	{Caraj\'as Province, Brazil. Gravity data on a regular grid of $500 \times 500$ points, totaling $250,000$ observations.
-	The inset shows the study area (blue rectangle) which covers the southeast part of the state of Par\'a, north of Brazil.}
+	{Application to field data over the Caraj{\'a}s Province, Brazil. 
+	Observed gravity data on a regular grid of $500 \times 500$ points, totaling $N = 250,000$ 
+	observations. The inset shows the study area (blue rectangle) which covers the southeast part of 
+	the state of Par{\'a}, north of Brazil.}
 	\label{fig:carajas_real_data}
@@
 \plot{carajas_gz_predito_val}{width=8cm}{
-	{Caraj\'as Province, Brazil. (a) Predicted gravity data produced by our modification of the fast equivalent-layer 
-	method \citep{siqueira-etal2017}  that computes the forward modeling using the properties of BTTB and BCCB matrices
-	(equation \ref{eq:w_Cv}). (b) Gravity residuals, defined as the difference between the observed data in 
-	Figure \ref{fig:carajas_real_data} and the predicted data in panel a, with their mean of 0.000292 mGal and standard
-	deviation of $0.105$ mGal.}
+	{Application to field data over the Caraj{\'a}s Province, Brazil. 
+	(a) Predicted data produced by our method. (b) Residuals between the observed  
+	(Figure \ref{fig:carajas_real_data}) and the predicted data (panel a), with mean 
+	$0.000292$ mGal and standard deviation of $0.105$ mGal.}
 	\label{fig:carajas_gz_predito_val}
@@
 \plot{up2000_carajas_500x500}{width=13.5cm}{
-	{Caraj\'as Province, Brazil. The upward-continued gravity data using our modification of the fast equivalent layer 
-	method \citep{siqueira-etal2017} that computes the forward modeling using the properties of BTTB and BCCB matrices
-	(equation \ref{eq:w_Cv}). The total computation time is 0.216 seconds for processing of the $250,000$ observations.}
+	{Application to field data over the Caraj{\'a}s Province, Brazil. 
+	The upward-continued gravity data obtained with our method $5\,000$ m above the 
+	observed data (Figure \ref{fig:carajas_real_data}).
+	The total computation time for processing of the $250,000$ observations was $0.216$ seconds.}
 	\label{fig:up2000_carajas_500x500}

### 2019-12-20 manuscript/DFT_computations.tex
final version pressumission

@@
 by using the row-ordered $vec$-operator (here designated simply as $vec$-operator).
-This equation can be efficiently computed by using the two-dimensional 
+This equation can be efficiently computed by using the 2D 
 fast Fourier Transform. 
@@
 is obtained by applying the $vec$-operator (equation \ref{eq:vec-operator}) to a matrix 
-$\mathbf{W}$, whose two-dimensional DFT $\tilde{\mathbf{W}}$ is represented by the 
+$\mathbf{W}$, whose 2D DFT $\tilde{\mathbf{W}}$ is represented by the 
 following separable transformation \citep[][ p. 146]{jain1989}:
@@
 is obtained by applying the $vec$-operator (equation \ref{eq:vec-operator}) to a matrix 
-$\mathbf{V}$, whose two-dimensional DFT (equation \ref{eq:2D-DFT-W}) is 
-represented by $\tilde{\mathbf{V}}$, and use
-equation \ref{eq:vec-matrix-S} to rewrite the 
+$\mathbf{V}$, whose 2D DFT (equation \ref{eq:2D-DFT-W}) is 
+represented by $\tilde{\mathbf{V}}$. Using equation \ref{eq:vec-matrix-S} and the symmetry 
+of unitary DFT matrices, we can rewrite the 
 left side of equation \ref{eq:vec-DFT-system} as follows:
@@
 Then, consider that $\boldsymbol{\lambda}$ is obtained by applying the $vec$-operator 
-(equation \ref{eq:vec-operator}) to a $2Q \times 2P$ matrix $\mathbf{L}$ and use 
+(equation \ref{eq:vec-operator}) to a $2Q \times 2P$ matrix $\mathbf{L}$, we can use 
 equations \ref{eq:vec-matrix-H} and \ref{eq:vec-matrix-S} to rewrite equation 

### 2019-12-20 manuscript/eigenvalues_C.tex
final version pressumission

@@
 (equation \ref{eq:w_Cv}), forming the rows of $\mathbf{L}$,
-are obtained by computing the two-dimensional DFT of matrix $\mathbf{G}$,
+are obtained by computing the 2D DFT of matrix $\mathbf{G}$,
 which contains the elements forming the first column of the BCCB matrix 

### 2019-12-20 manuscript/figures.tex
final version pressumission

@@
 \renewcommand{\figdir}{Fig} % figure directory
+
+\plot{Figure1}{width=\textwidth}{
+	{Schematic representation of an $N_{x} \times N_{y}$ regular grid of points (black dots) defined by 
+	$N_{x} = 4$ and $N_{y} = 3$. The grids are oriented along the (a) $x$-axis and (b) $y$-axis. The grid 
+	coordinates are $x_{k}$ and $y_{k}$, where the $k = 1, \dots, N_{x}$ and $l = 1, \dots, N_{y}$ are 
+	called the grid indices. The insets show the grid indices $k$ and $l$.}
+	\label{fig:methodology}
+}
 
@@
 
-\plot{float}{width=\textwidth}{
+\plot{Figure2}{width=\textwidth}{
 	{Comparison between the number of flops (equations \ref{eq:float_fast_eqlayer} and \ref{eq:float_bccb}) 
@@
 
-\plot{time_comparison}{width=10cm}{
+\plot{Figure3}{width=10cm}{
 	{Comparison between the running time of the fast equivalent-layer technique \citep{siqueira-etal2017} and our method.
@@
 
-\plot{time_bccb}{width=10cm}{
+\plot{Figure4}{width=10cm}{
 	{Running time of our method for a number of observations $N$ up to $25$ millions. 
@@
 
-\plot{synthetic_data}{width=13.5cm}{
+\plot{Figure5}{width=13.5cm}{
 	{Application to synthetic data. 
@@
 
-\plot{classic_fast_val}{width=8cm}{
+\plot{Figure6}{width=8cm}{
 	{Application to synthetic data. 
@@
 
-\plot{bccb_fast_val}{width=8cm}{
+\plot{Figure7}{width=8cm}{
 	{Application to synthetic data. 
@@
 
-\plot{delta_comparison}{width=13.5cm}{
+\plot{Figure8}{width=13.5cm}{
 	{Application to synthetic data. 
@@
 
-\plot{delta_rho}{width=13.5cm}{
+\plot{Figure9}{width=13.5cm}{
 	{Application to synthetic data. 
@@
 
-\plot{upward_comparison}{width=5cm}{
+\plot{Figure10}{width=5cm}{
 	{Application to synthetic data. 
@@
 
-\plot{downward_comparison}{width=5cm}{
+\plot{Figure11}{width=5cm}{
 	{Application to synthetic data. 
@@
 
-\plot{carajas_real_data}{width=13.5cm}{
+\plot{Figure12}{width=13.5cm}{
 	{Application to field data over the Caraj{\'a}s Province, Brazil. 
@@
 
-\plot{carajas_gz_predito_val}{width=8cm}{
+\plot{Figure13}{width=8cm}{
 	{Application to field data over the Caraj{\'a}s Province, Brazil. 
@@
 
-\plot{up2000_carajas_500x500}{width=13.5cm}{
+\plot{Figure14}{width=13.5cm}{
 	{Application to field data over the Caraj{\'a}s Province, Brazil. 

### 2019-12-20 manuscript/introduction.tex
final version pressumission

@@
 distribution. In practical situations, the layer is approximated by a finite set of sources (e.g., point masses 
-or dipoles) and their physical properties are estimated by solving a linear system of equations that yield an 
+or dipoles) and their physical properties are estimated by solving a linear system of equations that yields an 
 acceptable potential-field data fit. These fictitious sources are called equivalent sources.
@@
 height and (ii) there is one point mass directly beneath each observation point.
+This technique have been successfully used in potential-field methods for 
+3D gravity inversion \citep{zhang-wong2015}, downward continuation 
+\citep{zhang-etal2016} and 3D magnetic modeling \citep{qiang_etal2019}.
 By using this property, we propose an efficient algorithm based on fast FFT convolution 
@@
 not only in the number of flops, but also in the RAM memory usage of the fast equivalent-layer technique.
-This technique have been successfully used in potential-field methods for 
-3D gravity inversion \citep{zhang-wong2015}, downward continuation 
-\citep{zhang-etal2016} and 3D magnetic modeling \citep{qiang_etal2019}.
 Tests with synthetic and field data illustrate the good performance of our method in processing 

### 2019-12-20 manuscript/methodology.tex
final version pressumission

@@
 the $z$ axis pointing downward.
-Let us consider an equivalent layer compose by a set of $N$ point masses 
+Let us consider an equivalent layer composed by a set of $N$ point masses 
 (equivalent sources) over a layer located at depth $z_0$ ($z_0 >z_i$) and whose 
-$x$- and $y$- coordinates of each point mass coincide with the corresponding coordinates 
+$x$- and $y$- coordinates of each point mass coincides with the corresponding coordinates 
 of the observation directly above.
@@
 transforms from $\mathrm{m/s^2}$ to mGal.
-Notice that the sensitivity matrix depends on the $i$th coordinates of the observations 
-and the $j$th coordinates of the equivalent sources. For convenience, we designate 
+Notice that the sensitivity matrix depends on the $i$th coordinate of the observation 
+and the $j$th coordinate of the equivalent source. For convenience, we designate 
 these coordinates as \textit{matrix coordinates} and the indices $i$ and $j$ as 
@@
 \subsection{Fast equivalent-layer technique}
+
 \citet{siqueira-etal2017} developed an iterative least-squares method to estimate the mass 
 distribution over the equivalent layer based on the excess of mass and the positive correlation 
-between the observed gravity data and the masses on the equivalent layer. Those authors showed 
+between the observed gravity data and the masses on the equivalent layer. They showed 
 that the fast equivalent-layer technique has a better computational efficiency than the 
 classical equivalent layer approach (equation \ref{eq:estimated-p-parameter-space}) if the 
-dataset totalize at least 200 observation points, even using a large number of iterations.
+dataset is greater than at least 200 observation points, even using a large number of iterations.
 
@@
 new residual $\mathbf{d^0} - \tensor{A} \hat{\mathbf{p}}^{k}$, which represents a bottleneck. 
-Considering the limitation of 16 Gb of RAM 
-memory in our system, we could run their method only up to 22 500 observation points. 
-Hence, for very large data sets it is costful and can be overwhelming in terms of RAM 
+Considering the limitation of 16 Gb of RAM memory in our system, we could run the 
+\citeauthor{siqueira-etal2017}'s~(\citeyear{siqueira-etal2017}) method only up to 22 500 observation points;
+Otherwise, it is costly and can be prohibitive in terms of RAM 
 memory to maintain such operation.
@@
 Consider that the observed data are located on an $N_{x} \times N_{y}$ regular grid of
-points regularly spaced from $\Delta x$ and $\Delta y$ along the $x$ and $y$ directions,
+points regularly spaced by $\Delta x$ and $\Delta y$ along the $x$ and $y$ directions,
 respectively, on a horizontal plane defined by the constant vertical coordinate $z_{1} < z_{0}$. 
@@
 These equations can also be used to define the matrix coordinates 
-$x_{j}$ and $y_{j}$ associated with the $j$-th equivalent source,
+$x_{j}$ and $y_{j}$ associated with the $j$th equivalent source,
 $j = 1, \dots, N = N_{x}N_{y}$. In this case, the integer functions
@@
 Consider the case in which the grid is oriented along the
-$x$-axis (Figure 1a). For convenience, we designate these grids as 
+$x$-axis (Figure \ref{fig:methodology}a). For convenience, we designate these grids as 
 $x$-\textit{oriented grids}. For them, we have the following integer functions:
@@
 be defined in the same way by using the index $j$.
-Figure 1a illustrates an $x$-oriented grid defined by $N_{x} = 4$ and $N_{y} = 3$.
+Figure \ref{fig:methodology}a illustrates an $x$-oriented grid defined by $N_{x} = 4$ and $N_{y} = 3$.
 In this example, the matrix coordinates $x_{7}$ and $y_{7}$, defined by the matrix index $i = 7$ (or $j = 7$), 
@@
 Now, consider the case in which the regular grid of data is oriented along 
-the $y$-axis (Figure 1b). For convenience, we call them $y$-\textit{oriented grids}.
+the $y$-axis (Figure \ref{fig:methodology}b). For convenience, we call them $y$-\textit{oriented grids}.
 Similarly to $x$-oriented grids, we have the following integer functions associated with
@@
 \end{equation}
-Figure 1b illustrates an $y$-oriented grid defined by $N_{x} = 4$ and $N_{y} = 3$.
+Figure \ref{fig:methodology}b illustrates an $y$-oriented grid defined by $N_{x} = 4$ and $N_{y} = 3$.
 In this example, the matrix coordinates $x_{7}$ and $y_{7}$, defined by the matrix index 
@@
 defined by the grid indices $k = 3$ and $l = 1$, respectively. Differently from the example
-shown in Figure 1a, the grid indices of the present example are 
+shown in Figure \ref{fig:methodology}a, the grid indices of the present example are 
 computed with equations \ref{eq:k-y-oriented} and \ref{eq:l-y-oriented}, by using the 
@@
 where $\Delta z = z_{0} - z_{1}$, 
-$\Delta k_{ij} = k(i) - k(j)$ (equations \ref{eq:k-x-oriented} and \ref{eq:k-y-oriented}) and
-$\Delta l_{ij} = l(i) - l(j)$ (equations \ref{eq:l-x-oriented} and \ref{eq:l-y-oriented}).
+$\Delta k_{ij} = k(i) - k(j)$ (equations \ref{eq:k-x-oriented} or \ref{eq:k-y-oriented}) and
+$\Delta l_{ij} = l(i) - l(j)$ (equations \ref{eq:l-x-oriented} or \ref{eq:l-y-oriented}).
 Notice that the structure of matrix $\mathbf{A}$ (equation \ref{eq:predicted-data-vector}) for 
@@
 defined by matrix indices $i$ and $j$, there is a corresponding block element
-$a^{q}_{p}$, defined by block indices $q$ (equations \ref{eq:q-x-oriented} and 
-\ref{eq:q-y-oriented}) and $p$ (equations \ref{eq:p-x-oriented} and \ref{eq:p-y-oriented}),
+$a^{q}_{p}$, defined by block indices $q$ (equations \ref{eq:q-x-oriented} or 
+\ref{eq:q-y-oriented}) and $p$ (equations \ref{eq:p-x-oriented} or \ref{eq:p-y-oriented}),
 so that
@@
 $a^{q}_{p}$ (equation \ref{eq:aqp_equiv_aij}) in terms of the block indices 
-$q$ (equations \ref{eq:q-x-oriented} and \ref{eq:q-y-oriented}) and $p$ 
-(equations \ref{eq:p-x-oriented} and \ref{eq:p-y-oriented}).
+$q$ (equations \ref{eq:q-x-oriented} or \ref{eq:q-y-oriented}) and $p$ 
+(equations \ref{eq:p-x-oriented} or \ref{eq:p-y-oriented}).
 
 It is important noting that different matrix indices $i$ or $j$ produce the same 
-absolute values for the grid indices $k$ (equations \ref{eq:k-x-oriented} and
-\ref{eq:k-y-oriented}) and $l$ (equations \ref{eq:l-x-oriented} and
+absolute values for the grid indices $k$ (equations \ref{eq:k-x-oriented} or 
+\ref{eq:k-y-oriented}) and $l$ (equations \ref{eq:l-x-oriented} or 
 \ref{eq:l-y-oriented}). As a consequence, different pairs of matrix indices $i$
@@
 (equation \ref{eq:aij-regular-grids}) and also the same values for the block indices
-$q$ (equations \ref{eq:q-x-oriented} and \ref{eq:q-y-oriented}) and 
-$p$ (equations \ref{eq:p-x-oriented} and \ref{eq:p-y-oriented}), as well as the same
+$q$ (equations \ref{eq:q-x-oriented} or \ref{eq:q-y-oriented}) and 
+$p$ (equations \ref{eq:p-x-oriented} or \ref{eq:p-y-oriented}), as well as the same
 block element $a^{q}_{p}$ (equation \ref{eq:aqp_equiv_aij}). 
@@
 defined by matrix indices $i$ and $j$, compute the grid indices 
-$k$ (equations \ref{eq:k-x-oriented} and \ref{eq:k-y-oriented}) and
-$l$ (equations \ref{eq:l-x-oriented} and \ref{eq:l-y-oriented}).
+$k$ (equations \ref{eq:k-x-oriented} or \ref{eq:k-y-oriented}) and
+$l$ (equations \ref{eq:l-x-oriented} or \ref{eq:l-y-oriented}).
 These grid indices are used to 
@@
 \ref{eq:aij-regular-grids} and 
-(2) compute the block indices $q$ (equations \ref{eq:q-x-oriented}, 
-and \ref{eq:q-y-oriented}) and $p$ (equations \ref{eq:p-x-oriented} and
-and \ref{eq:p-y-oriented}) and determine the corresponding block element $a^{q}_{p}$ 
+(2) compute the block indices $q$ (equations \ref{eq:q-x-oriented} 
+or \ref{eq:q-y-oriented}) and $p$ (equations \ref{eq:p-x-oriented} or 
+\ref{eq:p-y-oriented}) and determine the corresponding block element $a^{q}_{p}$ 
 (equation \ref{eq:aqp_equiv_aij}) forming $\mathbf{A}_{q}$ (equation \ref{eq:Aq_block}).
 
-Consider the $x$-oriented grid of $N_{x} \times N_{y}$ points shown in Figure 1a, 
+Consider the $x$-oriented grid of $N_{x} \times N_{y}$ points shown in Figure \ref{fig:methodology}a, 
 with $N_{x} = 4$, $N_{y} = 3$ and $N = N_{x} \, N_{y} = 12$.
@@
 $i = 2$ and $j = 10$, which is
-located at the upper right corner of $\mathbf{A}$ (equation \ref{eq:predicted-data-vector}), 
-in the 2nd line and 10th column.
+located in the 2nd line and 10th column of $\mathbf{A}$ (equation \ref{eq:predicted-data-vector}).
 By using equations \ref{eq:l-x-oriented} and \ref{eq:k-x-oriented}, we obtain the 
@@
 (equations \ref{eq:l-x-oriented} and \ref{eq:k-x-oriented}) as those associated with
-the element defined by $i = 2$ and $j = 10$. Consequently, all of them produce
+the element defined by $i = 2$ and $j = 10$ or vice versa ($i = 10$ and $j = 2$). 
+Consequently, all of them produce
 elements $a_{ij}$ (equation \ref{eq:aij-regular-grids}) having the same value.
@@
 forming the matrix $\mathbf{A}$ (equation \ref{eq:predicted-data-vector}) obtained from our 
-$x$-oriented grid (Figure 1a), we can verify that
+$x$-oriented grid (Figure \ref{fig:methodology}a), we can verify that
 \begin{equation}
@@
 \mathbf{A}_{2} & \mathbf{A}_{1} & \mathbf{A}_{0}
-\end{bmatrix} \quad ,
+\end{bmatrix}_{N \times N} \quad ,
 \label{eq:A-x-oriented-example}
@@
 a^{q}_{3} & a^{q}_{2} & a^{q}_{1} & a^{q}_{0}
-\end{bmatrix} \quad ,
+\end{bmatrix}_{N_{x} \times N_{x}} \quad ,
 \label{eq:Aq-x-oriented}
@@
 (equation \ref{eq:predicted-data-vector}) obtained
-from the $y$-oriented grid illustrated in Figure 1b is given by
+from the $y$-oriented grid illustrated in Figure \ref{fig:methodology}b is given by
 \begin{equation}
@@
 \mathbf{A}_{3} & \mathbf{A}_{2} & \mathbf{A}_{1} & \mathbf{A}_{0}
-\end{bmatrix} \quad ,
+\end{bmatrix}_{N \times N} \quad ,
 \label{eq:A-y-oriented-example}
@@
 a^{q}_{2} & a^{q}_{1} & a^{q}_{0}
-\end{bmatrix} \quad ,
+\end{bmatrix}_{N_{y} \times N_{y}} \quad ,
 \label{eq:Aq-y-oriented}
@@
 can be defined by using only the elements 
-forming its first column (or row). Notice that this column contains the gravitational effect 
+forming its first column (or row). Notice that a column contains the gravitational effect 
 produced by a single equivalent source at all $N$ observation points.
@@
 and can cause RAM memory shortage when large data sets are used.
-This computational load can be drastically lessen by exploring the well-defined structure of 
+This computational load can be drastically reduced by exploring the well-defined structure of 
 matrix $\mathbf{A}$ (equation \ref{eq:predicted-data-vector}) for the particular case in which 
@@
 $\mathbf{d}(\mathbf{p})$ (equation \ref{eq:predicted-data-vector}) can be efficiently
-computed by using the two-dimensional Discrete Fourier Transform (DFT).
+computed by using the 2D Discrete Fourier Transform (DFT).
 To do this, let us first rewrite $\mathbf{d}(\mathbf{p})$ and
@@
 a computationally efficient way by exploring the structure of matrix $\mathbf{C}$.
-We show that this approach, that have been used in potential-field methods
-\citep[e.g.,][]{zhang-wong2015, zhang-etal2016, qiang_etal2019}, is actually a fast 
-convolution method \citep[e.g.,][ p. 213]{vanloan1992}.
 
 Matrix $\mathbf{C}$ (equation \ref{eq:w_Cv})
-is circulant blockwise, formed by $Q \times Q$ blocks, where
+is circulant blockwise, formed by $2Q \times 2Q$ blocks, where
 each block $\mathbf{C}_{q}$, $q = 0, \dots, Q-1$, is a $2P \times 2P$ circulant matrix. 
 Similarly to the BTTB matrix $\mathbf{A}$ (equations \ref{eq:BTTB_A} and 
-\ref{eq:A-x-oriented-example}--\ref{eq:Aq-y-oriented}), 
-the index $q$ defines the block diagonal where $\mathbf{C}_{q}$ lies 
-within $\mathbf{C}$. This index varies from $0$, at the main diagonal, to $Q - 1$, at
-the corners of $\mathbf{C}$. Additionally, the blocks lying 
+\ref{eq:A-x-oriented-example}--\ref{eq:Aq-y-oriented}), the index $q$ 
+varies from $0$ to $Q - 1$.
+
+Additionally, the blocks lying 
 above the main diagonal are equal to those located below.
+
 It is well-known that a circulant matrix can be defined by properly downshifting 
@@
 down one notch with wraparound \citep[][ p. 20]{golub-vanloan2013}.
-To illustrate this operation, consider our $y$-oriented grid illustrated in Figure 1b. 
+To illustrate this operation, consider our $y$-oriented grid illustrated in Figure \ref{fig:methodology}b. 
 In this case, the resulting 
@@
 	\mathbf{C_{1}} & \mathbf{C_{2}} & \mathbf{C_{3}} & \mathbf{0}     & \mathbf{C_{3}} & \mathbf{C_{2}} & \mathbf{C_{1}} & \mathbf{C_{0}}
-\end{bmatrix},
+\end{bmatrix}_{4N \times 4N},
 \label{eq:C-y-oriented}
@@
 	a^{q}_{1} & a^{q}_{2} & 0         & a^{q}_{2} & a^{q}_{1} & a^{q}_{0}
-\end{bmatrix}
+\end{bmatrix}_{2P \times 2P}
 \label{eq:Cq-y-oriented}
@@
 in terms of the block elements $a^{q}_{p}$ (equation \ref{eq:aqp_equiv_aij}).
-Similar matrices are obtained for our $x$-oriented grid illustrated in Figure 1a.
+Similar matrices are obtained for our $x$-oriented grid illustrated in Figure \ref{fig:methodology}a.
 
-BCCB matrices are diagonalized by the two-dimensional unitary DFT 
+BCCB matrices are diagonalized by the 2D unitary DFT 
 \citep[][ p. 185]{davis1979}. It means that $\mathbf{C}$ (equation \ref{eq:w_Cv}) 
@@
 $\boldsymbol{\Lambda}$, vector $\mathbf{v}$ and vector $\mathbf{w}$, respectively.
-The left side of equation \ref{eq:DFT-system} contains the two-dimensional 
+The left side of equation \ref{eq:DFT-system} contains the 2D 
 Inverse Discrete Fourier Transform (IDFT) of the term in brackets, which in turn
 represents the Hadamard product of matrix $\mathbf{L}$ (equation \ref{eq:left_side_DFT_system_3})
-and the two-dimensional DFT of matrix $\mathbf{V}$ (equation \ref{eq:left_side_DFT_system_1}).
+and the 2D DFT of matrix $\mathbf{V}$ (equation \ref{eq:left_side_DFT_system_1}).
 Matrix $\mathbf{L}$ contains the eigenvalues 
@@
 Here, we evaluate equation \ref{eq:DFT-system} and compute matrix $\mathbf{L}$
-by using the two-dimensional Fast Fourier Transform (2D FFT).
-This approach is a fast two-dimensional discrete convolution \citep[e.g.,][ p. 213]{vanloan1992}.
+by using the 2D Fast Fourier Transform (2D FFT).
+This approach, that have been used in potential-field methods
+\citep[e.g.,][]{zhang-wong2015, zhang-etal2016, qiang_etal2019}, is actually 
+a fast 2D discrete convolution \citep[e.g.,][ p. 213]{vanloan1992}.
 
@@
 
-\item[\textbf{(6)}] Use the 2D FFT to compute the IDFT of the result obtained at item (5) and 
+\item[\textbf{(6)}] Use the 2D FFT to compute the IDFT of the result obtained in step (5) to 
 obtain the matrix $\mathbf{W}$ (equation \ref{eq:DFT-system});
@@
 \item[\textbf{(7)}] Use the $vec$-operator (equation \ref{eq:vec-operator}) and equations 
-\ref{eq:w-vector} and \ref{eq:wq-vector} to rearrange $\mathbf{W}$ and obtain the predicted 
+\ref{eq:w-vector} and \ref{eq:wq-vector} to rearrange $\mathbf{W}$ in order to obtain the predicted 
 data vector $\mathbf{d}(\hat{\mathbf{p}}^{k})$.
@@
 (left side of equation \ref{eq:DFT-system}). 
-The Hadamard product requires six times $4N$ flops, $N = QP$, because the entries are 
+The Hadamard product requires $24N$ flops, $N = QP$, because the entries are 
 complex numbers.
@@
 full matrix $\mathbf{A}$, a single column of $\mathbf{A}$ and the full matrix 
-$\mathbf{L}$. The quantities were computed for different numbers of observations $N$. 
+$\mathbf{L}$. These quantities were computed for different numbers of observations $N$. 
 Notice that $N = 1\,000\,000$ observations require nearly $7.6$ Terabytes of RAM memory 

### 2019-12-20 manuscript/synthetic_data.tex
final version pressumission

@@
 Figure \ref{fig:synthetic_data} shows the gravity disturbance (vertical component of gravitational attraction) 
-produced by these sources. The synthetic data are contaminated with additive pseudo-random Gaussian 
+produced by these sources. The synthetic data are contaminated with additive pseudorandom Gaussian 
 noise with zero mean and standard deviation of $0.015$ mGal.
@@
 and \ref{fig:bccb_fast_val}b), defined as the difference between the simulated (Figure \ref{fig:synthetic_data}) 
-and predicted data (Figures \ref{fig:classic_fast_val}a and \ref{fig:bccb_fast_val}a) after $N^{it} = XXXX$ iterations, 
+and predicted data (Figures \ref{fig:classic_fast_val}a and \ref{fig:bccb_fast_val}a) after $N^{it} = 23$ iterations, 
 show means close to zero and standard deviations of $0.0144$ mGal, indicating that both methods produced virtually 

### 2020-03-21 manuscript/abstract.tex

@@
 to reconstruct the full sensitivity matrix completely. From the first column of BCCB matrix, its eigenvalues 
-can be calculated using the two-dimensional Fast Fourier Transform (2D FFT), which can be used to readily 
+can be calculated using the 2D Fast Fourier Transform (2D FFT), which can be used to readily 
 compute the matrix-vector product of the forward modeling in the fast equivalent-layer technique. As a result, 
@@
 Synthetic tests demonstrate the ability of our method to satisfactorily upward- and downward-continuing the 
-gravity data. For example, while our method requires 26:8 seconds to run one million of observations, the fast 
-equivalent-layer technique required 48:3 seconds to run 22,500 observations. Test with real data from Caraj\'as, 
+gravity data. For example, while our method requires 26.8 seconds to run one million of observations, the fast 
+equivalent-layer technique required 48.3 seconds to run 22,500 observations. Test with real data from Caraj{\'a}s, 
 Brazil, shows its applicability to process very large dataset at low computational cost.

### 2020-03-21 manuscript/field_data.tex

@@
 We applied our method to airborne gravity data from Caraj{\'a}s, north of Brazil, which were provided by 
-the Geological Survey of Brazil (CPRM). The data were collected along $113$ north–south flight lines separated 
-by $3$ km and XXXXXX east–west tie lines separated by $12$ km.
+the Geological Survey of Brazil (CPRM). The data were collected along $131$ north-south flight lines separated 
+by $3$ km and $29$ east-west tie lines separated by $12$ km.
 This data set was divided in two different areas, collected in different times, having samples spacing of 
@@
 
-To apply our method, we set an equivalent layer at $z_{0} = -300$ m. 
+To apply our method, we set an equivalent layer at $z_{0} = 300$ m. 
 Figure \ref{fig:carajas_gz_predito_val}a shows the predicted data obtained with our method after 
@@
 shows a very consistent upward-continued gravity data, with a clear attenuation of the short 
-wave-lengths. By using our approach, the processing of the $250\,000$ observations took only 
+wavelengths. By using our approach, the processing of the $250\,000$ observations took only 
 $0.216$ seconds.

### 2020-03-21 manuscript/methodology.tex
create appendix A and fix small typo errors

@@
 $4N \times 4N$ symmetric Block Circulant matrix with Circulant Blocks (BCCB) 
-\citep[][ p. 184]{davis1979}.
+\citep[][ p. 184]{davis1979}. A detailed description of this matrix and
+some of its relevant properties are presented in Appendix A.
+
 What follows shows a step-by-step description of how we use the auxiliary 
@@
 a computationally efficient way by exploring the structure of matrix $\mathbf{C}$.
-
-Matrix $\mathbf{C}$ (equation \ref{eq:w_Cv})
-is circulant blockwise, formed by $2Q \times 2Q$ blocks, where
-each block $\mathbf{C}_{q}$, $q = 0, \dots, Q-1$, is a $2P \times 2P$ circulant matrix. 
-Similarly to the BTTB matrix $\mathbf{A}$ (equations \ref{eq:BTTB_A} and 
-\ref{eq:A-x-oriented-example}--\ref{eq:Aq-y-oriented}), the index $q$ 
-varies from $0$ to $Q - 1$.
-
-Additionally, the blocks lying 
-above the main diagonal are equal to those located below.
-
-It is well-known that a circulant matrix can be defined by properly downshifting 
-its first column \citep[][ p. 206]{vanloan1992}. Hence, the BCCB matrix $\mathbf{C}$ 
-(equation \ref{eq:w_Cv}) can be obtained from its 
-first column of blocks, which is given by
-\begin{equation}
-\left[\mathbf{C} \right]_{(0)} = 
-	\begin{bmatrix}
-	\mathbf{C}_{0} \\
-	\vdots \\
-	\mathbf{C}_{Q-1} \\
-	\mathbf{0} \\
-	\mathbf{C}_{Q-1} \\
-	\vdots \\
-	\mathbf{C}_{1}
-	\end{bmatrix}_{4N \times 2P} \: ,
-	\label{eq:C-first-column-blocks}
-\end{equation}
-where $\mathbf{0}$ is a $2P \times 2P$ matrix of zeros. Similarly, each block 
-$\mathbf{C}_{q}$, $q = 0, \dots, Q-1$, can be obtained by downshifting its first 
-column
-\begin{equation}
-\mathbf{c}^{q}_{0} = 
-	\begin{bmatrix}
-		a^{q}_{0} \\
-		\vdots \\
-		a^{q}_{P-1} \\
-		0 \\
-		a^{q}_{P-1} \\
-		\vdots \\
-		a^{q}_{1}
-	\end{bmatrix}_{2P \times 1} \: ,
-	\label{eq:Cq-first-column}
-\end{equation}
-where $a^{q}_{p}$ (equation \ref{eq:aqp_equiv_aij}), $p = 0, \dots, P-1$, are the elements 
-forming the block $\mathbf{A}_{q}$ (equations \ref{eq:Aq_block} and 
-\ref{eq:A-x-oriented-example}--\ref{eq:Aq-y-oriented}).
-The downshift can be thought off as permutation that pushes the components of a column vector 
-down one notch with wraparound \citep[][ p. 20]{golub-vanloan2013}.
-To illustrate this operation, consider our $y$-oriented grid illustrated in Figure \ref{fig:methodology}b. 
-In this case, the resulting 
-BCCB matrix $\mathbf{C}$ (equation \ref{eq:w_Cv}) is given by 
-\begin{equation}
-\mathbf{C} =
-\begin{bmatrix}
-	\mathbf{C_{0}} & \mathbf{C_{1}} & \mathbf{C_{2}} & \mathbf{C_{3}} & \mathbf{0}     & \mathbf{C_{3}} & \mathbf{C_{2}} & \mathbf{C_{1}} \\
-	\mathbf{C_{1}} & \mathbf{C_{0}} & \mathbf{C_{1}} & \mathbf{C_{2}} & \mathbf{C_{3}} & \mathbf{0}     & \mathbf{C_{3}} & \mathbf{C_{2}} \\
-	\mathbf{C_{2}} & \mathbf{C_{1}} & \mathbf{C_{0}} & \mathbf{C_{1}} & \mathbf{C_{2}} & \mathbf{C_{3}} & \mathbf{0}     & \mathbf{C_{3}} \\
-	\mathbf{C_{3}} & \mathbf{C_{2}} & \mathbf{C_{1}} & \mathbf{C_{0}} & \mathbf{C_{1}} & \mathbf{C_{2}} & \mathbf{C_{3}} & \mathbf{0}     \\
-	\mathbf{0}     & \mathbf{C_{3}} & \mathbf{C_{2}} & \mathbf{C_{1}} & \mathbf{C_{0}} & \mathbf{C_{1}} & \mathbf{C_{2}} & \mathbf{C_{3}} \\
-	\mathbf{C_{3}} & \mathbf{0}     & \mathbf{C_{3}} & \mathbf{C_{2}} & \mathbf{C_{1}} & \mathbf{C_{0}} & \mathbf{C_{1}} & \mathbf{C_{2}} \\
-	\mathbf{C_{2}} & \mathbf{C_{3}} & \mathbf{0}     & \mathbf{C_{3}} & \mathbf{C_{2}} & \mathbf{C_{1}} & \mathbf{C_{0}} & \mathbf{C_{1}} \\
-	\mathbf{C_{1}} & \mathbf{C_{2}} & \mathbf{C_{3}} & \mathbf{0}     & \mathbf{C_{3}} & \mathbf{C_{2}} & \mathbf{C_{1}} & \mathbf{C_{0}}
-\end{bmatrix}_{4N \times 4N},
-\label{eq:C-y-oriented}
-\end{equation}
-where each block $\mathbf{C}_{q}$, $q = 0, 1, 2$, is represented as follows 
-\begin{equation}
-\tensor{C}_{q} =
-\begin{bmatrix}
-	a^{q}_{0} & a^{q}_{1} & a^{q}_{2} & 0         & a^{q}_{2} & a^{q}_{1} \\
-	a^{q}_{1} & a^{q}_{0} & a^{q}_{1} & a^{q}_{2} & 0         & a^{q}_{2} \\
-	a^{q}_{2} & a^{q}_{1} & a^{q}_{0} & a^{q}_{1} & a^{q}_{2} & 0         \\
-	0         & a^{q}_{2} & a^{q}_{1} & a^{q}_{0} & a^{q}_{1} & a^{q}_{2} \\
-	a^{q}_{2} & 0         & a^{q}_{2} & a^{q}_{0} & a^{q}_{0} & a^{q}_{1} \\
-	a^{q}_{1} & a^{q}_{2} & 0         & a^{q}_{2} & a^{q}_{1} & a^{q}_{0}
-\end{bmatrix}_{2P \times 2P}
-\label{eq:Cq-y-oriented}
-\end{equation}
-in terms of the block elements $a^{q}_{p}$ (equation \ref{eq:aqp_equiv_aij}).
-Similar matrices are obtained for our $x$-oriented grid illustrated in Figure \ref{fig:methodology}a.
-
-BCCB matrices are diagonalized by the 2D unitary DFT 
-\citep[][ p. 185]{davis1979}. It means that $\mathbf{C}$ (equation \ref{eq:w_Cv}) 
-satisfies 
-\begin{equation}
-\mathbf{C} = 
-\left(\mathbf{F}_{2Q} \otimes \mathbf{F}_{2P} \right)^{\ast} 
-\boldsymbol{\Lambda}
-\left(\mathbf{F}_{2Q} \otimes \mathbf{F}_{2P} \right) \: ,
-\label{eq:C-diagonalized}
-\end{equation}
-where the symbol ``$\otimes$" denotes the Kronecker product \citep{neudecker1969},
-$\mathbf{F}_{2Q}$ and $\mathbf{F}_{2P}$ are the $2Q \times 2Q$ and $2P \times 2P$ 
-unitary DFT matrices \citep[][ p. 31]{davis1979}, respectively, the superscritpt 
-``$\ast$" denotes the complex conjugate and $\boldsymbol{\Lambda}$ is a 
-$4QP \times 4QP$ diagonal matrix containing the eigenvalues of $\mathbf{C}$.
 By substituting equation \ref{eq:C-diagonalized} in the auxiliary system 
 (equation \ref{eq:w_Cv}) and premultiplying both
-sides of the result by $\left(\mathbf{F}_{2Q} \otimes \mathbf{F}_{2P} \right)$, 
-we obtain
+sides of the result by $\left(\mathbf{F}_{2Q} \otimes \mathbf{F}_{2P} \right)$
+(see the details in Appendix A), we obtain
 \begin{equation}
@@
 Now, by applying the $vec$-operator to both sides of equation \ref{eq:vec-DFT-system} 
-(see the details in Appendix A), we obtain:
+(see the details in Appendix B), we obtain:
 \begin{equation}
@@
 efficiently computed by using only the first column of the BCCB matrix 
-$\mathbf{C}$ (equation \ref{eq:w_Cv}) (see the details in Appendix B).
+$\mathbf{C}$ (equation \ref{eq:w_Cv}) (see the details in Appendix C).
 Here, we evaluate equation \ref{eq:DFT-system} and compute matrix $\mathbf{L}$
@@
 Our method replace this matrix-vector product by three operations: 
-one DFT, one Hadammard product and one IDFT involving $2Q \times 2P$ matrices 
+one DFT, one Hadamard product and one IDFT involving $2Q \times 2P$ matrices 
 (left side of equation \ref{eq:DFT-system}). 

### 2020-03-25 manuscript/field_data.tex
correct some numbers in field data application

@@
 This data set was divided in two different areas, collected in different times, having samples spacing of 
-$7.65$ m and $15.21$ m along the lines, totalizing  $4\,353\,428$ observation points at a fixed height 
+$7.65$ m and $15.21$ m along the lines, totalizing  $5\,492\,551$ observation points at a fixed height 
 of $900$ m ($z_{1} = -900$ m). 
@@
 (Figure \ref{fig:carajas_real_data}) and predicted (Figure \ref{fig:carajas_gz_predito_val}a) data, show a 
-very good data fit with mean close to zero ($0.0003$ mGal) and small standard deviation ($0.105$ mGal), 
+very good data fit with mean close to zero ($0.0003$ mGal) and small standard deviation ($0.1160$ mGal), 
 which corresponds to approximately $0.1$ \% of the maximum amplitude of the gravity data.

### 2020-03-26 manuscript/figures.tex
include new figures and modifications in text

@@
 
-\plot{Figure6}{width=8cm}{
+\plot{Figure678_r1}{width=6cm}{
 	{Application to synthetic data. 
-	(a) Predicted data produced by the fast equivalent-layer technique \citep{siqueira-etal2017}. 
-	(b) Residuals between the simulated (Figure \ref{fig:synthetic_data}) and 
-	predicted (shown in a) data. The mean and standard deviation are $8.264 \times 10^{-7}$ and $0.0144$ mGal,
-	respectively.}
-	\label{fig:classic_fast_val}
+	(a) Residuals between the simulated data (Figure \ref{fig:synthetic_data}) and 
+	predicted data produced by the fast equivalent-layer technique \citep{siqueira-etal2017}. 
+	The mean and standard deviation are $8.264 \times 10^{-7}$ and $0.0144$ mGal, respectively.
+	(b) Residuals between the simulated data (Figure \ref{fig:synthetic_data}) and 
+	predicted data produced by our method. The mean and standard deviation are 
+	$8.264 \times 10^{-7}$ and $0.0144$ mGal, respectively.
+	(c) Difference between the predicted data produced by the fast equivalent-layer technique 
+	and our method.}
+	\label{fig:comparison_datafits}
 }
 
-\plot{Figure7}{width=8cm}{
+\plot{Figure10_r1}{width=6cm}{
 	{Application to synthetic data.
-	(a) Predicted data produced by our method. 
-	(b) Residuals between the simulated (Figure \ref{fig:synthetic_data}) and 
-	predicted (shown in a) data. The mean and standard deviation are $8.264 \times 10^{-7}$ and $0.0144$ mGal,
-	respectively.}
-	\label{fig:bccb_fast_val}
-}
-
-\plot{Figure8}{width=13.5cm}{
-	{Application to synthetic data. 
-	Difference between the predicted data produced by the fast equivalent-layer technique 
-	(Figure \ref{fig:classic_fast_val}a) and by our method (Figure \ref{fig:bccb_fast_val}a).}
-	\label{fig:delta_comparison}
-}
-
-\plot{Figure9}{width=13.5cm}{
-	{Application to synthetic data. 
-	Difference between the estimated mass distribution within the equivalent layers produced by 
-	the fast equivalent-layer technique (not shown) and by our method (not shown).}
-	\label{fig:delta_rho}
-}
-
-\plot{Figure10}{width=5cm}{
-	{Application to synthetic data. 
-	The upward-continued gravity data obtained with (a) the fast equivalent-layer technique 
-	\citep{siqueira-etal2017} and (b) our method. (c) Residuals between 
-	panels a and b with mean $-5.938 \times 10^{-18}$ mGal and standard deviation $8.701 \times 10^{-18}$ mGal. 
-	The total computation times spent by the fast equivalent-layer technique and our method 
-	are $7.62026$ and $0.00834$ seconds, respectively.}
+	(a) Gravity data produced by the synthetic sources at $z = -XXXX$ m.
+	(b) Residuals between the simulated data shown in a and the upward-continued 
+	gravity data obtained with the fast equivalent-layer technique (not shown), with 
+	mean $XXXXXX$ mGal and standard deviation $XXXXXX$ mGal. 
+	(c) Residuals between the simulated data shown in a and the upward-continued 
+	gravity data obtained with our method (not shown), with mean $XXXXX$ mGal and 
+	standard deviation $XXXXX$ mGal. 
+	The total computation times spent by the fast equivalent-layer technique and 
+	our method were $7.62026$ and $0.00834$ seconds, respectively.}
 	\label{fig:upward_comparison}
@@
 
-\plot{Figure11}{width=5cm}{
+\plot{Figure11_r1}{width=6cm}{
 	{Application to synthetic data. 
-	The downward-continued gravity data obtained with (a) the fast equivalent-layer technique 
-	\citep{siqueira-etal2017} and (b) our method. (c) Residuals between 
-	panels a and b with mean $5.914 \times 10^{-18}$ mGal and standard deviation $9.014 \times 10^{-18}$ mGal. 
-	The total computation times spent by the fast equivalent-layer technique and our method 
-	are $7.59654$ and $0.00547$ seconds, respectively.}
+	(a) Gravity data produced by the synthetic sources at $z = -XXXX$ m.
+	(b) Residuals between the simulated data shown in a and the downward-continued 
+	gravity data obtained with the fast equivalent-layer technique (not shown), with 
+	mean $XXXXXX$ mGal and standard deviation $XXXXXX$ mGal. 
+	(c) Residuals between the simulated data shown in a and the downward-continued 
+	gravity data obtained with our method (not shown), with mean $XXXXX$ mGal and 
+	standard deviation $XXXXX$ mGal. 
+	The total computation times spent by the fast equivalent-layer technique and 
+	our method were $7.59654$ and $0.00547$ seconds, respectively.}
 	\label{fig:downward_comparison}
@@
 
-\plot{Figure13}{width=8cm}{
+\plot{Figure13_r1}{width=8cm}{
 	{Application to field data over the Caraj{\'a}s Province, Brazil. 

### 2020-03-26 manuscript/synthetic_data.tex
include new figures and modifications in text

@@
 
-Figures \ref{fig:classic_fast_val}a and \ref{fig:bccb_fast_val}a show the data fits 
+Figures \ref{fig:comparison_datafits}a and \ref{fig:comparison_datafits}b show the data fits 
 obtained, respectively, by the fast equivalent-layer technique \citep{siqueira-etal2017} 
-and by our method. The corresponding residuals (Figures \ref{fig:classic_fast_val}b 
-and \ref{fig:bccb_fast_val}b), defined as the difference between the simulated (Figure \ref{fig:synthetic_data}) 
-and predicted data (Figures \ref{fig:classic_fast_val}a and \ref{fig:bccb_fast_val}a) after $N^{it} = 23$ iterations, 
-show means close to zero and standard deviations of $0.0144$ mGal, indicating that both methods produced virtually 
-the same results.
-This excellent agreement is confirmed by Figures \ref{fig:delta_comparison} and 
-\ref{fig:delta_rho}. They show that there are virtually no differences between, respectively, 
-the predicted data (Figures \ref{fig:classic_fast_val}b and \ref{fig:bccb_fast_val}b) and the 
-estimated mass distributions within the equivalent layers (not shown) yielded by 
-both methods.
+and by our method. They represent the differences between the simulated data (Figure \ref{fig:synthetic_data}) and the predicted data produced by both methods 
+(not shown) after $N^{it} = 23$ iterations. 
+As we can see, both methods produced virtually the same results.
+This excellent agreement is confirmed by Figure \ref{fig:comparison_datafits}c, which shows 
+the differences between the predicted data produced by both methods.
 
 We performed the upward and downward continuations of the simulated gravity data 
-(Figure \ref{fig:synthetic_data}) by using the estimated equivalent layers.
-The results show that the continued gravity data obtained by using the 
-fast equivalent-layer (Figures \ref{fig:upward_comparison}a and \ref{fig:downward_comparison}a) 
-are in close agreement with those produced by our method (Figures \ref{fig:upward_comparison}b and 
-\ref{fig:downward_comparison}b). The residuals (Figures \ref{fig:upward_comparison}c and \ref{fig:downward_comparison}c)
-quantify this agreement since their means and standard deviations are close to zero.
-
-The continued gravity data obtained with both methods agree with the true ones (not shown).
-The difference is that the computation time spent by our method is 
-approximately $1\,500$ times smaller than that of the fast-equivalent-layer 
+(Figure \ref{fig:synthetic_data}) by using the fast equivalent-layer technique 
+and our method.
+Figure \ref{fig:upward_comparison} shows that the upward-continued gravity data obtained 
+by using both methods are in close agreement with those produced by the synthetic 
+bodies at $z = XXXX$ m (Figure \ref{fig:upward_comparison}a).
+As we can see in Figures \ref{fig:upward_comparison}b and \ref{fig:upward_comparison}c,
+the residuals between the upward-continued and true data have their means and standard 
+deviations close to zero. 
+Figure \ref{fig:downward_comparison} shows similar results obtained by using both 
+methods to compute downward continuation of gravity data.
+It is important to stress that, although both methods are able to produce a reliable 
+upward/downward continuation of gravity data, the computation time spent by our method 
+is approximately $1\,500$ times smaller than that of the fast equivalent-layer 
 technique \citep{siqueira-etal2017}.

### 2020-04-06 manuscript/figures.tex
include parameter values used to compute the continuation

@@
 	{Application to synthetic data.
-	(a) Gravity data produced by the synthetic sources at $z = -XXXX$ m.
+	(a) Gravity data produced by the synthetic sources at $z = -300$ m.
 	(b) Residuals between the simulated data shown in a and the upward-continued 
 	gravity data obtained with the fast equivalent-layer technique (not shown), with 
-	mean $XXXXXX$ mGal and standard deviation $XXXXXX$ mGal. 
+	mean $-0.0003$ mGal and standard deviation $0.0124$ mGal. 
 	(c) Residuals between the simulated data shown in a and the upward-continued 
-	gravity data obtained with our method (not shown), with mean $XXXXX$ mGal and 
-	standard deviation $XXXXX$ mGal. 
+	gravity data obtained with our method (not shown), with mean $-0.0003$ mGal and 
+	standard deviation $0.0124$ mGal. 
 	The total computation times spent by the fast equivalent-layer technique and 
@@
 	{Application to synthetic data. 
-	(a) Gravity data produced by the synthetic sources at $z = -XXXX$ m.
+	(a) Gravity data produced by the synthetic sources at $z = -50$ m.
 	(b) Residuals between the simulated data shown in a and the downward-continued 
 	gravity data obtained with the fast equivalent-layer technique (not shown), with 
-	mean $XXXXXX$ mGal and standard deviation $XXXXXX$ mGal. 
+	mean $0.0001$ mGal and standard deviation $0.0079$ mGal. 
 	(c) Residuals between the simulated data shown in a and the downward-continued 
-	gravity data obtained with our method (not shown), with mean $XXXXX$ mGal and 
-	standard deviation $XXXXX$ mGal. 
+	gravity data obtained with our method (not shown), with mean $0.0001$ mGal and 
+	standard deviation $0.0079$ mGal. 
 	The total computation times spent by the fast equivalent-layer technique and 

### 2020-04-29 manuscript/figures.tex
include new figures for tests with synthetic data

@@
 
-\plot{Figure5}{width=13.5cm}{
+\plot{synthetic_data_med}{width=13.5cm}{
 	{Application to synthetic data. 
 	Noise-corrupted gravity data (in color map) produced by three synthetic sources: 
-	a sphere with density contrast $-0.5\, \mathrm{g/cm^3}$ and 
-	two rectangular prisms with density contrasts $0.35\, \mathrm{g/cm^3}$ (upper-left body) and 
-	$0.4\, \mathrm{g/cm^3}$ (upper-right body). The black lines represent the horizontal projection 
-	of the sources.}
+	a sphere with density contrast $-1.25\, \mathrm{g/cm^3}$ and 
+	two rectangular prisms with density contrasts $1.00\, \mathrm{g/cm^3}$ (upper-left body) and 
+	$1.30\, \mathrm{g/cm^3}$ (upper-right body). 
+	The synthetic data are contaminated with a pseudorandom Gaussian noise having null mean 
+	and standard deviation equal to $0.1 \, \mathrm{mGal}$.
+	The black lines represent the horizontal projection of the sources.}
 	\label{fig:synthetic_data}
@@
 
-\plot{Figure678_r1}{width=6cm}{
+\plot{deltas_gz_med}{width=6cm}{
 	{Application to synthetic data. 
@@
 	predicted data produced by the fast equivalent-layer technique \citep{siqueira-etal2017}. 
-	The mean and standard deviation are $8.264 \times 10^{-7}$ and $0.0144$ mGal, respectively.
 	(b) Residuals between the simulated data (Figure \ref{fig:synthetic_data}) and 
-	predicted data produced by our method. The mean and standard deviation are 
-	$8.264 \times 10^{-7}$ and $0.0144$ mGal, respectively.
+	predicted data produced by our method. 
+	The mean ($-4.493 \times 10^{-5}$ mGal) and standard deviation ($0.093$ mGal) for residuals
+	shown in a and b are exactly the same.
 	(c) Difference between the predicted data produced by the fast equivalent-layer technique 
-	and our method.}
+	and our method.
+	The computation times spent by the fast equivalent-layer technique and 
+	our method were $10.996$ and $0.428$ seconds, respectively.}
 	\label{fig:comparison_datafits}
@@
 
-\plot{Figure10_r1}{width=6cm}{
+\plot{upward_fourier_med}{width=13.5cm}{
 	{Application to synthetic data.
 	(a) Gravity data produced by the synthetic sources at $z = -300$ m.
-	(b) Residuals between the simulated data shown in a and the upward-continued 
-	gravity data obtained with the fast equivalent-layer technique (not shown), with 
-	mean $-0.0003$ mGal and standard deviation $0.0124$ mGal. 
-	(c) Residuals between the simulated data shown in a and the upward-continued 
-	gravity data obtained with our method (not shown), with mean $-0.0003$ mGal and 
-	standard deviation $0.0124$ mGal. 
-	The total computation times spent by the fast equivalent-layer technique and 
-	our method were $7.62026$ and $0.00834$ seconds, respectively.}
+	(b) Residuals between the data shown in a and the upward-continued 
+	data obtained with our method (not shown), with 
+	mean $0.003$ mGal and standard deviation $0.034$ mGal. 
+	(c) Residuals between the data shown in a and the upward-continued 
+	data obtained with the fast equivalent-layer technique (not shown), with 
+	mean $0.003$ mGal and standard deviation $0.034$ mGal.
+	(d) Residuals between the data shown in a and the upward-continued 
+	data obtained with the classical Fourier approach (not shown), with 
+	mean $-0.030$ mGal and standard deviation $0.262$ mGal.
+	The residuals shown in b, c and d have the same scale. 	 
+	The computation times spent by the fast equivalent-layer technique and 
+	our method were $8.697$ and $0.005$ seconds, respectively.}
 	\label{fig:upward_comparison}
@@
 
-\plot{Figure11_r1}{width=6cm}{
+\plot{downward_fourier_med}{width=13.5cm}{
 	{Application to synthetic data. 
 	(a) Gravity data produced by the synthetic sources at $z = -50$ m.
-	(b) Residuals between the simulated data shown in a and the downward-continued 
-	gravity data obtained with the fast equivalent-layer technique (not shown), with 
-	mean $0.0001$ mGal and standard deviation $0.0079$ mGal. 
-	(c) Residuals between the simulated data shown in a and the downward-continued 
-	gravity data obtained with our method (not shown), with mean $0.0001$ mGal and 
-	standard deviation $0.0079$ mGal. 
-	The total computation times spent by the fast equivalent-layer technique and 
-	our method were $7.59654$ and $0.00547$ seconds, respectively.}
+	(b) Residuals between the data shown in a and the downward-continued 
+	data obtained with our method (not shown), with 
+	mean $-0.001$ mGal and standard deviation $0.038$ mGal. 
+	(c) Residuals between the data shown in a and the downward-continued 
+	data obtained with the fast equivalent-layer technique (not shown), with 
+	mean $-0.001$ mGal and standard deviation $0.038$ mGal.
+	(d) Residuals between the data shown in a and the downward-continued 
+	data obtained with the classical Fourier approach (not shown), with 
+	mean $-0.030$ mGal and standard deviation $0.262$ mGal.
+	The residuals shown in b, c and d have the same scale.	 
+	The computation times spent by the fast equivalent-layer technique and 
+	our method were $8.795$ and $0.004$ seconds, respectively.}
 	\label{fig:downward_comparison}

### 2020-04-29 manuscript/synthetic_data.tex
include new figures for tests with synthetic data

@@
 
-We simulated three sources whose horizontal projections are shown in Figure \ref{fig:synthetic_data} as black lines. 
-These sources are a sphere with radius of $1\,000$ m with density contrast of 
-$-0.5\, \mathrm{g/cm^3}$ and two rectangular prisms with density contrasts of $0.35\, \mathrm{g/cm^3}$ (upper-left prism) 
-and $0.4\, \mathrm{g/cm^3}$ (upper-right prism). 
-Figure \ref{fig:synthetic_data} shows the gravity disturbance (vertical component of gravitational attraction) 
-produced by these sources. The synthetic data are contaminated with additive pseudorandom Gaussian 
-noise with zero mean and standard deviation of $0.015$ mGal.
-We created $N = 10\,000$ observation points regularly spaced on a grid of $100 \times 100$, on a 
-horizontal plane at $z_{1} = -100$ m, and set a grid of equivalent point masses, each one directly beneath each 
-observation point, located at depth $z_{0} = 300$ m.
+We have simulated three sources whose horizontal projections are shown in 
+Figure \ref{fig:synthetic_data} as black lines. 
+These sources are a sphere with density contrast  
+$-1.25\, \mathrm{g/cm^3}$ and two rectangular prisms with density contrasts 
+$1.00\, \mathrm{g/cm^3}$ (upper-left prism) 
+and $1.30\, \mathrm{g/cm^3}$ (upper-right prism). 
+Figure \ref{fig:synthetic_data} shows the gravity disturbance (vertical component of 
+gravitational attraction) produced by these sources. 
+The synthetic data are contaminated with additive pseudorandom Gaussian 
+noise with zero mean and standard deviation of $0.1 \, \mathrm{mGal}$.
+The data are computed at $N = 10\,000$ observation points that are regularly spaced on a 
+$100 \times 100$ grid, at $z_{1} = -100$ m. 
+We have set a grid of equivalent sources, each one directly beneath each 
+observation point, located at $z_{0} = 300$ m.
+
+PAREI AQUI
 
@@
 by using both methods are in close agreement with those produced by the synthetic 
-bodies at $z = XXXX$ m (Figure \ref{fig:upward_comparison}a).
+bodies at $z = -300$ m (Figure \ref{fig:upward_comparison}a).
 As we can see in Figures \ref{fig:upward_comparison}b and \ref{fig:upward_comparison}c,

### 2020-04-29 manuscript/figures.tex
discuss the new results with synthetic data

@@
 	The computation times spent by the fast equivalent-layer technique and 
-	our method were $10.996$ and $0.428$ seconds, respectively.}
+	our method were $10.416$ and $0.177$ seconds, respectively.}
 	\label{fig:comparison_datafits}
@@
 	{Application to synthetic data.
-	(a) Gravity data produced by the synthetic sources at $z = -300$ m.
+	(a) Noise-free gravity data produced by the synthetic sources at $z = -300$ m.
 	(b) Residuals between the data shown in a and the upward-continued 
@@
 	mean $-0.030$ mGal and standard deviation $0.262$ mGal.
-	The residuals shown in b, c and d have the same scale. 	 
+	Values shown in b, c and d have the same scale. 	 
 	The computation times spent by the fast equivalent-layer technique and 
@@
 	{Application to synthetic data. 
-	(a) Gravity data produced by the synthetic sources at $z = -50$ m.
+	(a) Noise-free gravity data produced by the synthetic sources at $z = -50$ m.
 	(b) Residuals between the data shown in a and the downward-continued 
@@
 	mean $-0.030$ mGal and standard deviation $0.262$ mGal.
-	The residuals shown in b, c and d have the same scale.	 
+	Values shown in b, c and d have the same scale.	 
 	The computation times spent by the fast equivalent-layer technique and 

### 2020-04-29 manuscript/synthetic_data.tex
discuss the new results with synthetic data

@@
 We have set a grid of equivalent sources, each one directly beneath each 
-observation point, located at $z_{0} = 300$ m.
-
-PAREI AQUI
+observation point, at $z_{0} = 300$ m.
 
@@
 and by our method. They represent the differences between the simulated data (Figure \ref{fig:synthetic_data}) and the predicted data produced by both methods 
-(not shown) after $N^{it} = 23$ iterations. 
-As we can see, both methods produced virtually the same results.
+(not shown) after $N^{it} = 40$ iterations. 
+As we can see, both methods produce virtually the same results.
 This excellent agreement is confirmed by Figure \ref{fig:comparison_datafits}c, which shows 
@@
 (Figure \ref{fig:synthetic_data}) by using the fast equivalent-layer technique
-and our method.
-Figure \ref{fig:upward_comparison} shows that the upward-continued gravity data obtained 
-by using both methods are in close agreement with those produced by the synthetic 
-bodies at $z = -300$ m (Figure \ref{fig:upward_comparison}a).
-As we can see in Figures \ref{fig:upward_comparison}b and \ref{fig:upward_comparison}c,
-the residuals between the upward-continued and true data have their means and standard 
-deviations close to zero. 
-Figure \ref{fig:downward_comparison} shows similar results obtained by using both 
+\citep{siqueira-etal2017}, our method and also the classical approach in the 
+Fourier domain. This approach consists in performing the upward/downward continuation 
+by directly computing the Fourier transform of the gravity data 
+\citep[e.g., ][ p. 317]{blakely1996}.
+Figure \ref{fig:upward_comparison} shows the upward-continued gravity data 
+obtained by the three methods.
+As we can see, the residuals between the true data at $z = -300$ m 
+(Figure \ref{fig:upward_comparison}a) and the upward-continued data obtained by 
+using our method (Figure \ref{fig:upward_comparison}b) and 
+the fast equivalent-layer technique (Figure \ref{fig:upward_comparison}c) are very similar 
+to each other.
+Notice that the absolute values of the residuals produced by the classical Fourier 
+approach (Figure \ref{fig:upward_comparison}d) are $\approx 10$ times greater than
+those produced by our method and the fast equivalent-layer technique 
+(Figures \ref{fig:upward_comparison}b and \ref{fig:upward_comparison}c), 
+with maximum values concentrated at the border of the simulated area.
+We opted for showing the results produced by all methods without removing the border 
+effects in order to properly compare them to each other.
+We can also notice that, differently from the results produced by the other methods, 
+that obtained with the classical Fourier approach exhibits a slight noise amplification.
+
+Figure \ref{fig:downward_comparison} shows the results obtained by using all 
 methods to compute downward continuation of gravity data.
-It is important to stress that, although both methods are able to produce a reliable 
-upward/downward continuation of gravity data, the computation time spent by our method 
-is approximately $1\,500$ times smaller than that of the fast equivalent-layer 
-technique \citep{siqueira-etal2017}.
+In this case, the maximum absolute values of the residuals produced by the 
+classical Fourier approach (Figure \ref{fig:downward_comparison}d) are 
+$\approx 20$ times greater than those produced by our method and the fast equivalent-layer technique (Figures \ref{fig:downward_comparison}b and \ref{fig:downward_comparison}c).
+It is also important to stress that the result produced by the classical Fourier approach 
+exhibits a considerable noise amplification when compared to the other methods.
+This noise amplification is a well-known problem of the downward continuation 
+produced by the classical Fourier approach \citep[e.g., ][ p. 320]{blakely1996}.
+
+Another important aspect to be pointed out about the results we obtained with 
+synthetic data (Figures \ref{fig:upward_comparison} and \ref{fig:downward_comparison})
+is the computational times spent by our method and the fast equivalent-layer technique. 
+The total computational time required by our method to estimate the physical property 
+distribution on the equivalent layer and perform the upward/downward continuation 
+is about two orders of magnitude lower than that spent by the fast equivalent-layer technique.
+This significant reduction in computational time was obtained by using a data set 
+composed of $N = 10\,000$ observation points. Considerably better results can be 
+obtained with larger data sets.
+
+Finally, we did not compare the total computational times spent by our method and the classical 
+Fourier approach, but we can affirm that the second is faster. 
+While the classical approach needs basically one DFT/IDFT of the data, 
+our method requires one DFT/IDFT per iteration. 
+On the other hand, the considerably smaller noise amplification and 
+border effect are the main advantages of our method over the classical Fourier 
+approach, especially in the downward continuation.

### 2020-04-30 manuscript/abstract.tex
include the classical FFT results in abstract

@@
 dataset, the faster and more efficient our method becomes compared to the available equivalent-layer techniques. 
-Synthetic tests demonstrate the ability of our method to satisfactorily upward- and downward-continuing the 
-gravity data. For example, while our method requires 26.8 seconds to run one million of observations, the fast 
-equivalent-layer technique required 48.3 seconds to run 22,500 observations. Test with real data from Caraj{\'a}s, 
-Brazil, shows its applicability to process very large dataset at low computational cost.
+Tests with synthetic data demonstrate the ability of our method to satisfactorily upward-
+and downward-continuing the gravity data.
+Our results show very small border effects and noise amplification compared to those 
+produced by the classical approach in the Fourier domain.
+Besides, they show that while the running time of our method is $\approx 30.9$ seconds for 
+processing $N = 1\,000\,000$ observations, the fast equivalent-layer technique spent 
+$\approx 46.8$ seconds with $N = 22\,500$.
+A test with field data from Caraj{\'a}s province, Brazil, illustrates the low computational 
+cost of our method to process a large data set composed of $N = 250\,000$ observations.
 

### 2020-04-30 manuscript/abstract.tex
include FFT results in conclusion

@@
 $\approx 46.8$ seconds with $N = 22\,500$.
-A test with field data from Caraj{\'a}s province, Brazil, illustrates the low computational 
+A test with field data from Caraj{\'a}s Province, Brazil, illustrates the low computational 
 cost of our method to process a large data set composed of $N = 250\,000$ observations.

### 2020-04-30 manuscript/conclusions.tex
include FFT results in conclusion

@@
 linear systems.
-Our approach greatly reduces the number of flops and the RAM memory necessary to estimate a 
+Compared to the fast equivalent-layer technique, 
+our approach greatly reduces the number of flops and the RAM memory necessary to estimate a 
 2D mass distribution within a planar equivalent layer that fits the observed gravity data. 
-For example, the number of flops of the fast equivalent-layer technique is reduced by two orders 
+For example, the number of flops is reduced by two orders 
 of magnitude when processing one million of observations by using our approach. 
@@
 We have successfully applied the proposed method to compute the upward/downward continuation of 
-synthetic gravity data. Application to field data over the Caraj{\'a}s Province, north of Brazil, 
-confirms the potential of our approach in processing a gravity data set with $250\,000$ observations 
-in approximately $0.2$ seconds. Further studies need to be carried out in order to apply our 
-method to process other potential-field data.
+synthetic gravity data. 
+Our results show that our method produces considerably smaller border effects and 
+noise amplification compared to the classical Fourier approach, 
+which computes the Fourier transform of the observed data.
+Application to field data over the Caraj{\'a}s Province, north of Brazil, 
+confirms the potential of our approach in processing a gravity data set with $250\,000$ observations in approximately $0.2$ seconds.
+Further studies need to be carried out in order to apply our method to process other potential-field data.

### 2020-04-30 manuscript/conclusions.tex
include application to marine gravity data at conclusion

@@
 Application to field data over the Caraj{\'a}s Province, north of Brazil, 
-confirms the potential of our approach in processing a gravity data set with $250\,000$ observations in approximately $0.2$ seconds.
-Further studies need to be carried out in order to apply our method to process other potential-field data.
+confirms the potential of our approach in processing an airborne gravity data set with $250\,000$ observations in approximately $0.2$ seconds. 
+Our method can also be applied to process large marine gravity data sets. 
+Further studies need to be carried out in order to apply our method to process 
+other potential-field data.

### 2020-04-30 manuscript/synthetic_data.tex
include application to marine gravity data at conclusion

@@
 Finally, we did not compare the total computational times spent by our method and the classical 
-Fourier approach, but we can affirm that the second is faster. 
+Fourier approach, but we can affirm that the second is smaller. 
 While the classical approach needs basically one DFT/IDFT of the data, 

### 2020-05-05 manuscript/abstract.tex
changes made by Val

@@
 
-We have developed an efficient and very fast equivalent-layer technique for gravity data processing by modifying an 
-iterative method grounded on excess mass constraint that does not require the solution of linear systems. 
-Taking advantage of the symmetric Block-Toeplitz Toeplitz-block (BTTB) structure of the sensitivity matrix, that raises 
-when regular grids of observation points and equivalent sources (point masses) are used to set up a fictitious 
-equivalent layer, we have developed an algorithm which greatly reduces the number of flops and RAM memory necessary 
-to estimate a 2D mass distribution over the equivalent layer. The structure of symmetric BTTB matrix consists of 
-the elements of the first column of the sensitivity matrix, which in turn can be embedded into a symmetric 
-Block-Circulant Circulant-Block (BCCB) matrix. Likewise, only the first column of the BCCB matrix is needed 
-to reconstruct the full sensitivity matrix completely. From the first column of BCCB matrix, its eigenvalues 
-can be calculated using the 2D Fast Fourier Transform (2D FFT), which can be used to readily 
-compute the matrix-vector product of the forward modeling in the fast equivalent-layer technique. As a result, 
-our method is efficient to process very large datasets using either fine- or mid-grid meshes. The larger the 
-dataset, the faster and more efficient our method becomes compared to the available equivalent-layer techniques. 
+We have developed an efficient and very fast equivalent-layer technique for 
+gravity data processing by modifying an iterative method grounded on excess mass 
+constraint that does not require the solution of linear systems. 
+Taking advantage of the symmetric Block-Toeplitz Toeplitz-block (BTTB) structure 
+of the sensitivity matrix, that raises when regular grids of observation points 
+and equivalent sources (point masses) are used to set up a fictitious 
+equivalent layer, we have developed an algorithm which greatly reduces the number o
+f flops and RAM memory necessary to estimate a 2D mass distribution over the 
+equivalent layer. 
+The structure of symmetric BTTB matrix consists of the elements of the first column of 
+the sensitivity matrix, which in turn can be embedded into a symmetric 
+Block-Circulant Circulant-Block (BCCB) matrix. 
+Likewise, only the first column of the BCCB matrix is needed 
+to reconstruct the full sensitivity matrix completely. From the first column of 
+BCCB matrix, its eigenvalues can be calculated using the 2D Fast Fourier Transform 
+(2D FFT), which can be used to readily compute the matrix-vector product of the 
+forward modeling in the fast equivalent-layer technique. 
+As a result, our method is efficient to process very large datasets using either 
+fine- or mid-grid meshes. 
+The larger the dataset, the faster and more efficient our method becomes compared 
+to the available equivalent-layer techniques. 
 Tests with synthetic data demonstrate the ability of our method to satisfactorily upward-
@@
 Besides, they show that while the running time of our method is $\approx 30.9$ seconds for 
-processing $N = 1\,000\,000$ observations, the fast equivalent-layer technique spent 
-$\approx 46.8$ seconds with $N = 22\,500$.
+processing $N = 1,000,000$ observations, the fast equivalent-layer technique spent 
+$\approx 46.8$ seconds with $N = 22,500$.
 A test with field data from Caraj{\'a}s Province, Brazil, illustrates the low computational 
-cost of our method to process a large data set composed of $N = 250\,000$ observations.
+cost of our method to process a large data set composed of $N = 250,000$ observations.
 

### 2020-05-05 manuscript/conclusions.tex
changes made by Val

@@
 By exploring the BTTB structure of the sensitivity matrix, we formulate the forward modeling as a 
-fast FFT convolution that requires only one column of the sensitivity matrix 
-and propose an efficient approach for optimizing the computational time of an 
+FFT convolution that requires only one column of the sensitivity matrix; 
+hence, the forward modeling uses, in practice, a single equivalent source,
+which results in a drastic reduction of the required RAM memory. 
+Next, we propose an efficient approach for optimizing the computational time of an 
 iterative method called fast equivalent-layer technique.
@@
 the full sensitivity matrix whereas our method takes only $61.035$ Megabytes.
-This drastic reduction comes from the fact that, by using only one column of the sensitivity matrix,
-our method is able to compute the gravity data produced by the whole equivalent layer with a single 
-point mass via FFT.
-We have successfully applied the proposed method to compute the upward/downward continuation of 
+We have successfully applied the proposed method to compute the upward/downward continuations of 
 synthetic gravity data. 
@@
 Application to field data over the Caraj{\'a}s Province, north of Brazil, 
-confirms the potential of our approach in processing an airborne gravity data set with $250\,000$ observations in approximately $0.2$ seconds. 
+confirms the potential of our approach in processing an airborne gravity data set with $250,000$ observations in approximately $0.2$ seconds. 
 Our method can also be applied to process large marine gravity data sets. 

### 2020-05-05 manuscript/field_data.tex
changes made by Val

@@
 This data set was divided in two different areas, collected in different times, having samples spacing of 
-$7.65$ m and $15.21$ m along the lines, totalizing  $5\,492\,551$ observation points at a fixed height 
+$7.65$ m and $15.21$ m along the lines, totalizing  $5,492,551$ observation points at a fixed height 
 of $900$ m ($z_{1} = -900$ m). 
 The gravity data were interpolated (Figure \ref{fig:carajas_real_data}) into a regularly spaced grid of 
-$500 \times 500$ observation points ($N = 250\,000$) with a grid spacing of $716.9311$ m north-south and 
-$781.7387$ m east-west.
+$500 \times 500$ observation points ($N = 250,000$) with a grid spacing of 
+$\approx 717$ m north-south and $\approx 782$ m east-west.
 
@@
 By using the estimated mass distribution (not shown), we performed an upward-continuation of the 
-observed gravity data to a horizontal plane located $5\,000$ m above. Figure \ref{fig:up2000_carajas_500x500} 
+observed gravity data to a horizontal plane located $5,000$ m above. Figure \ref{fig:up2000_carajas_500x500} 
 shows a very consistent upward-continued gravity data, with a clear attenuation of the short 
-wavelengths. By using our approach, the processing of the $250\,000$ observations took only 
+wavelengths. By using our approach, the processing of the $250,000$ observations took only 
 $0.216$ seconds.

### 2020-05-05 manuscript/figures.tex
changes made by Val

@@
 	associated with the fast equivalent-layer technique \citep{siqueira-etal2017} and our method, 
-	for $N$ varying from $5\,000$ to $1\,000\,000$. All values are computed with $N^{it} = 50$ 
+	for $N$ varying from $5,000$ to $1,000,000$. All values are computed with $N^{it} = 50$ 
 	iterations and $\kappa = 5$.}
@@
 	$1.30\, \mathrm{g/cm^3}$ (upper-right body). 
-	The synthetic data are contaminated with a pseudorandom Gaussian noise having null mean 
-	and standard deviation equal to $0.1 \, \mathrm{mGal}$.
 	The black lines represent the horizontal projection of the sources.}
@@
 	{Application to synthetic data. 
-	(a) Residuals between the simulated data (Figure \ref{fig:synthetic_data}) and 
-	predicted data produced by the fast equivalent-layer technique \citep{siqueira-etal2017}. 
-	(b) Residuals between the simulated data (Figure \ref{fig:synthetic_data}) and 
-	predicted data produced by our method. 
+	Residuals between the simulated data (Figure \ref{fig:synthetic_data}) and 
+	predicted data produced by: (a) the fast equivalent-layer technique \citep{siqueira-etal2017} and (b) our method. 
 	The mean ($-4.493 \times 10^{-5}$ mGal) and standard deviation ($0.093$ mGal) for residuals
@@
 	(a) Noise-free gravity data produced by the synthetic sources at $z = -300$ m.
-	(b) Residuals between the data shown in a and the upward-continued 
-	data obtained with our method (not shown), with 
-	mean $0.003$ mGal and standard deviation $0.034$ mGal. 
-	(c) Residuals between the data shown in a and the upward-continued 
-	data obtained with the fast equivalent-layer technique (not shown), with 
-	mean $0.003$ mGal and standard deviation $0.034$ mGal.
-	(d) Residuals between the data shown in a and the upward-continued 
-	data obtained with the classical Fourier approach (not shown), with 
+	Residuals between the data shown in a and the upward-continued 
+	data obtained by: (b) our method (not shown), with 
+	mean $0.003$ mGal and standard deviation $0.034$ mGal, 
+	(c) the fast equivalent-layer technique (not shown), with 
+	mean $0.003$ mGal and standard deviation $0.034$ mGal and 
+	(d) the classical Fourier approach (not shown), with 
 	mean $-0.030$ mGal and standard deviation $0.262$ mGal.
@@
 	(a) Noise-free gravity data produced by the synthetic sources at $z = -50$ m.
-	(b) Residuals between the data shown in a and the downward-continued 
-	data obtained with our method (not shown), with 
-	mean $-0.001$ mGal and standard deviation $0.038$ mGal. 
-	(c) Residuals between the data shown in a and the downward-continued 
-	data obtained with the fast equivalent-layer technique (not shown), with 
-	mean $-0.001$ mGal and standard deviation $0.038$ mGal.
-	(d) Residuals between the data shown in a and the downward-continued 
-	data obtained with the classical Fourier approach (not shown), with 
+	Residuals between the data shown in a and the downward-continued 
+	data obtained by: (b) our method (not shown), with 
+	mean $-0.001$ mGal and standard deviation $0.038$ mGal, 
+	(c) the fast equivalent-layer technique (not shown), with 
+	mean $-0.001$ mGal and standard deviation $0.038$ mGal and 
+	(d) the classical Fourier approach (not shown), with 
 	mean $-0.030$ mGal and standard deviation $0.262$ mGal.
@@
 	{Application to field data over the Caraj{\'a}s Province, Brazil. 
-	The upward-continued gravity data obtained with our method $5\,000$ m above the 
+	The upward-continued gravity data obtained with our method $5,000$ m above the 
 	observed data (Figure \ref{fig:carajas_real_data}).

### 2020-05-05 manuscript/synthetic_data.tex
changes made by Val

@@
 noise with zero mean and standard deviation of $0.1 \, \mathrm{mGal}$.
-The data are computed at $N = 10\,000$ observation points that are regularly spaced on a 
+The data are computed at $N = 10,000$ observation points that are regularly spaced on a 
 $100 \times 100$ grid, at $z_{1} = -100$ m. 
@@
 
-Figures \ref{fig:comparison_datafits}a and \ref{fig:comparison_datafits}b show the data fits 
+Figure \ref{fig:comparison_datafits}a and \ref{fig:comparison_datafits}b show the data fits 
 obtained, respectively, by the fast equivalent-layer technique \citep{siqueira-etal2017} 
@@
 those produced by our method and the fast equivalent-layer technique 
-(Figures \ref{fig:upward_comparison}b and \ref{fig:upward_comparison}c), 
+(Figure \ref{fig:upward_comparison}b and \ref{fig:upward_comparison}c), 
 with maximum values concentrated at the border of the simulated area.
-We opted for showing the results produced by all methods without removing the border 
-effects in order to properly compare them to each other.
-We can also notice that, differently from the results produced by the other methods, 
-that obtained with the classical Fourier approach exhibits a slight noise amplification.
+Differently from the results yield by our method (Figure \ref{fig:upward_comparison}b) 
+and the fast equivalent-layer technique (Figure \ref{fig:upward_comparison}c), 
+that obtained with the classical Fourier approach exhibits a slight noise 
+amplification (Figure \ref{fig:upward_comparison}d).
 
@@
 classical Fourier approach (Figure \ref{fig:downward_comparison}d) are 
-$\approx 20$ times greater than those produced by our method and the fast equivalent-layer technique (Figures \ref{fig:downward_comparison}b and \ref{fig:downward_comparison}c).
-It is also important to stress that the result produced by the classical Fourier approach 
-exhibits a considerable noise amplification when compared to the other methods.
+$\approx 20$ times greater than those produced by our method and the fast equivalent-layer technique (Figure \ref{fig:downward_comparison}b and \ref{fig:downward_comparison}c).
 This noise amplification is a well-known problem of the downward continuation 
@@
 
-Another important aspect to be pointed out about the results we obtained with 
-synthetic data (Figures \ref{fig:upward_comparison} and \ref{fig:downward_comparison})
+We opted for showing all the results (Figures \ref{fig:upward_comparison} and 
+\ref{fig:downward_comparison}) produced by all methods without removing the border 
+effects in order to properly compare them to each other.
+We also stress that no padding function to expand the data was used in applying 
+our method, the fast equivalent-layer technique or the Fourier approach.
+Another important aspect to be pointed out about these results 
 is the computational times spent by our method and the fast equivalent-layer technique. 
-The total computational time required by our method to estimate the physical property 
-distribution on the equivalent layer and perform the upward/downward continuation 
+The total computational time required by our method to estimate the physical-property 
+distribution on the equivalent layer and to perform the upward/downward continuations 
 is about two orders of magnitude lower than that spent by the fast equivalent-layer technique.
 This significant reduction in computational time was obtained by using a data set 
-composed of $N = 10\,000$ observation points. Considerably better results can be 
+composed of $N = 10,000$ observation points. Considerably better results can be 
 obtained with larger data sets.
 
-Finally, we did not compare the total computational times spent by our method and the classical 
-Fourier approach, but we can affirm that the second is smaller. 
-While the classical approach needs basically one DFT/IDFT of the data, 
-our method requires one DFT/IDFT per iteration. 
-On the other hand, the considerably smaller noise amplification and 
+Finally, we did not compare the total computational times spent by our method 
+and the classical Fourier approach, but we can affirm that the second is smaller. 
+Because  the Fourier approach requires only one DFT/IDFT of the data, whereas 
+our method requires one DFT/IDFT per iteration, it is computationally faster 
+than our method.
+However, the considerably smaller noise amplification and practically nonexistent 
 border effect are the main advantages of our method over the classical Fourier 

### 2020-05-06 manuscript/figures.tex
changes in figures 5 to 11

@@
 	predicted data produced by: (a) the fast equivalent-layer technique \citep{siqueira-etal2017} and (b) our method. 
-	The mean ($-4.493 \times 10^{-5}$ mGal) and standard deviation ($0.093$ mGal) for residuals
-	shown in a and b are exactly the same.
-	(c) Difference between the predicted data produced by the fast equivalent-layer technique 
-	and our method.
+	The mean ($-4.493 \times 10^{-5}$ mGal) and standard deviation ($0.093$ mGal) for residuals shown in a and b are exactly the same.
+	(c) Difference between a and b.
 	The computation times spent by the fast equivalent-layer technique and 
@@
 	mean $-0.030$ mGal and standard deviation $0.262$ mGal. 
-	Values shown in b, c and d have the same scale. 	 
 	The computation times spent by the fast equivalent-layer technique and 
@@
 	mean $-0.030$ mGal and standard deviation $0.262$ mGal.
-	Values shown in b, c and d have the same scale.	 
 	The computation times spent by the fast equivalent-layer technique and 
@@
 
-\plot{Figure12}{width=13.5cm}{
+\plot{carajas_real_data}{width=13.5cm}{
 	{Application to field data over the Caraj{\'a}s Province, Brazil. 
 	Observed gravity data on a regular grid of $500 \times 500$ points, totaling $N = 250,000$ 
-	observations. The inset shows the study area (blue rectangle) which covers the southeast part of 
+	observations. The inset shows the study area (red rectangle) which covers the southeast part of 
 	the state of Par{\'a}, north of Brazil.}
@@
 
-\plot{Figure13_r1}{width=8cm}{
+\plot{carajas_gz_predito}{width=8cm}{
 	{Application to field data over the Caraj{\'a}s Province, Brazil. 
@@
 
-\plot{Figure14}{width=13.5cm}{
+\plot{up5000_carajas_500x500}{width=13.5cm}{
 	{Application to field data over the Caraj{\'a}s Province, Brazil. 

### 2020-08-07 manuscript/abstract.tex

--- 2020-04-30 manuscript/abstract.tex
@@
 
-We have developed an efficient and very fast equivalent-layer technique for gravity data processing by modifying an 
-iterative method grounded on excess mass constraint that does not require the solution of linear systems. 
-Taking advantage of the symmetric Block-Toeplitz Toeplitz-block (BTTB) structure of the sensitivity matrix, that raises 
-when regular grids of observation points and equivalent sources (point masses) are used to set up a fictitious 
-equivalent layer, we have developed an algorithm which greatly reduces the number of flops and RAM memory necessary 
-to estimate a 2D mass distribution over the equivalent layer. The structure of symmetric BTTB matrix consists of 
-the elements of the first column of the sensitivity matrix, which in turn can be embedded into a symmetric 
-Block-Circulant Circulant-Block (BCCB) matrix. Likewise, only the first column of the BCCB matrix is needed 
-to reconstruct the full sensitivity matrix completely. From the first column of BCCB matrix, its eigenvalues 
-can be calculated using the 2D Fast Fourier Transform (2D FFT), which can be used to readily 
-compute the matrix-vector product of the forward modeling in the fast equivalent-layer technique. As a result, 
-our method is efficient to process very large datasets using either fine- or mid-grid meshes. The larger the 
-dataset, the faster and more efficient our method becomes compared to the available equivalent-layer techniques. 
+We have developed an efficient and very fast equivalent-layer technique for 
+gravity data processing by modifying an iterative method grounded on excess mass 
+constraint that does not require the solution of linear systems. 
+Taking advantage of the symmetric Block-Toeplitz Toeplitz-block (BTTB) structure 
+of the sensitivity matrix, that raises when regular grids of observation points 
+and equivalent sources (point masses) are used to set up a fictitious 
+equivalent layer, we have developed an algorithm which greatly reduces the number 
+of flops and RAM memory necessary to estimate a 2D mass distribution over the 
+equivalent layer. 
+The structure of symmetric BTTB matrix consists of the elements of the first column of 
+the sensitivity matrix, which in turn can be embedded into a symmetric 
+Block-Circulant Circulant-Block (BCCB) matrix. 
+Likewise, only the first column of the BCCB matrix is needed 
+to reconstruct the full sensitivity matrix completely. From the first column of 
+BCCB matrix, its eigenvalues can be calculated using the 2D Fast Fourier Transform 
+(2D FFT), which can be used to readily compute the matrix-vector product of the 
+forward modeling in the fast equivalent-layer technique. 
+As a result, our method is efficient to process very large datasets. 
 Tests with synthetic data demonstrate the ability of our method to satisfactorily upward-

### 2020-08-07 manuscript/introduction.tex

@@
 upward/downward continuation \citep[e.g.,][]{emilia1973, hansen-miyazaki1984, 
-li-oldenburg2010}, reduction to the pole \citep[e.g.,]{silva1986, leao-silva1989, guspi-novara2009, 
+li-oldenburg2010}, reduction to the pole \citep[e.g.,][]{silva1986, leao-silva1989, guspi-novara2009, 
 oliveirajr-etal2013}, combining multiple data sets \citep[e.g.,][]{boggs-dransfield2004} and 
@@
 The method consists in estimating an equivalent layer which exactly reproduces the potential-field data 
-within a small data window. The data window is shifted over the whole gridded data set in a procedure similar 
+within a small data window. The data window is shifted over the whole gridded data in a procedure similar 
 to a discrete convolution. 
@@
 interpolating surface that fits the chosen subset also automatically fits all remaining data. 
-The equivalent data set is obtained by iteratively introducing the potential-field observation with the 
+The equivalent-data set is obtained by iteratively introducing the potential-field observation with the 
 greatest residual in the preceding iteration. By applying to the interpolation problem, the method is 
@@
 \citet{oliveirajr-etal2013}, the computational efficiency of their method relies on the fact that the total 
-number of polynomial coefficients needed to describe the physical property distribution within the 
+number of polynomial coefficients needed to describe the physical-property distribution within the 
 equivalent layer is generally much smaller than the number of equivalent sources, leading to a very smaller 
@@
 the equivalent layer without solving linear systems.
-The method presented by \citet{cordell1992} and after generalized by \citet{guspi-novara2009} updates 
+The method presented by \citet{cordell1992}, and later generalized by \citet{guspi-novara2009}, updates 
 the physical property of the sources, which are located below each potential-field data, using a 
@@
 \citep{zhang-etal2016} and 3D magnetic modeling \citep{qiang_etal2019}.
-By using this property, we propose an efficient algorithm based on fast FFT convolution 
+By using this property, we propose an efficient algorithm based on FFT convolution 
 \citep[e.g.,][ p. 207]{vanloan1992} for computing the forward problem at each iteration of 

### 2020-08-07 manuscript/methodology.tex

@@
 the point $(x_{i}, y_{i}, z_{i})$, $i = 1, ..., N$, of a local Cartesian
-system with $x$ axis pointing to north, the $y$ axis pointing to east and 
-the $z$ axis pointing downward.
+system with $x$-axis pointing to north, the $y$-axis pointing to east and 
+the $z$-axis pointing downward.
 Let us consider an equivalent layer composed by a set of $N$ point masses 
@@
 Considering the limitation of 16 Gb of RAM memory in our system, we could run the 
-\citeauthor{siqueira-etal2017}'s~(\citeyear{siqueira-etal2017}) method only up to 22 500 observation points;
+\citeauthor{siqueira-etal2017}'s~(\citeyear{siqueira-etal2017}) method only up to $22,500$  observation points;
 Otherwise, it is costly and can be prohibitive in terms of RAM 
@@
 Consider that the observed data are located on an $N_{x} \times N_{y}$ regular grid of
-points regularly spaced by $\Delta x$ and $\Delta y$ along the $x$ and $y$ directions,
+points regularly spaced by $\Delta x$ and $\Delta y$ along the $x$- and $y$-directions,
 respectively, on a horizontal plane defined by the constant vertical coordinate $z_{1} < z_{0}$. 
@@
 
-\item[\textbf{(3)}] Rearrange $\mathbf{c}_{0}$ along the rows of matrix $\mathbf{G}$ and
-use the 2D FFT to compute matrix $\mathbf{L}$ (equation \ref{eq:DTF_G});
+\item[\textbf{(3)}] Rearrange $\mathbf{c}_{0}$ along the rows of matrix $\mathbf{G}$
+(equation \ref{eq:DFT_G}) and use the 2D FFT to compute matrix $\mathbf{L}$ 
+(equation \ref{eq:DFT_G});
 
@@
 and uses it to compute the complex $2Q \times 2P$ matrix $\mathbf{L}$ 
-(equation \ref{eq:DTF_G}) via 2D FFT, which is stored during all iterations.
+(equation \ref{eq:DFT_G}) via 2D FFT, which is stored during all iterations.
 Table \ref{tab:RAM-usage} shows the RAM memory usage needed to store the 
@@
 $\mathbf{L}$. These quantities were computed for different numbers of observations $N$. 
-Notice that $N = 1\,000\,000$ observations require nearly $7.6$ Terabytes of RAM memory 
+Notice that $N = 1,000,000$ observations require nearly $7.6$ Terabytes of RAM memory 
 to store the whole matrix $\mathbf{A}$.
@@
 The computational efficiency of our approach is significantly superior to that of the 
-fast equivalent-layer technique for a number of observations $N$ greater than $10\,000$. 
-We could not perform this comparison with a number of observations greater than $22\,500$
+fast equivalent-layer technique for a number of observations $N$ greater than $10,000$. 
+We could not perform this comparison with a number of observations greater than $22,500$
 due to limitations of our PC in storing the full matrix $\mathbf{A}$.
@@
 These results shows that, while the running time of our method is $\approx 30.9$ seconds for 
-$N = 1\,000\,000$, the fast equivalent-layer technique spends $\approx 46.8$ seconds for $N = 22\,500$.
+$N = 1,000,000$, the fast equivalent-layer technique spends $\approx 46.8$ seconds for $N = 22,500$.

### 2020-08-07 manuscript/tables.tex
include parameter values used to compute the continuation
changes made by Val

@@
 matrix $\mathbf{A}$ (equations \ref{eq:predicted-data-vector} and \ref{eq:BTTB_A}), a single 
-column of $\mathbf{A}$ and the full matrix $\mathbf{L}$ (equation \ref{eq:DTF_G}). The quantities 
+column of $\mathbf{A}$ and the full matrix $\mathbf{L}$ (equation \ref{eq:DFT_G}). The quantities 
 are represented in megabyte (MB), for different total number of observations $N$. We consider 
@@
 			\hline 
-			$100$ & 0.0763 & 0.0000763 & 0.0006104\\
+			$100$ & 0.0763 & 0.000763 & 0.00610\\
 			\hline
@@
 			\hline
-			$2\,500$ & 48 & 0.0191 & 0.1528\\
+			$2,500$ & 48 & 0.0191 & 0.1528\\
 			\hline
-			$10\,000$ & 763 & 0.00763 & 0.6104\\
+			$10,000$ & 763 & 0.0763 & 0.6104\\
 			\hline
-			$40\,000$ & 12\,207 & 0.305 & 2.4416 \\
+			$40,000$ & 12,207 & 0.305 & 2.4416 \\
 			\hline
-			$250\,000$ & 476\,837 & 1.907 & 15.3 \\
+			$250,000$ & 476,837 & 1.907 & 15.3 \\
 			\hline
-			$500\,000$ & 1\,907\,349 & 3.815 & 30.518 \\
+			$500,000$ & 1,907,349 & 3.815 & 30.518 \\
 			\hline
-			$1\,000\,000$ & 7\,629\,395 & 7.629 & 61.035 \\
+			$1,000,000$ & 7,629,395 & 7.629 & 61.035 \\
 			\hline

### 2020-08-07 manuscript/synthetic_data.tex
include EIC and reviewer #3 comments/corrections

@@
 to each other.
-Notice that the absolute values of the residuals produced by the classical Fourier 
+Note that the absolute values of the residuals produced by the classical Fourier 
 approach (Figure \ref{fig:upward_comparison}d) are $\approx 10$ times greater than
@@
 our method, the fast equivalent-layer technique or the Fourier approach.
+It is important to stress that the Fourier method shows heavy border effects at large 
+altitudes if no padding is made \citep{baniamerian_etal2017}. 
 Another important aspect to be pointed out about these results 
-is the computational times spent by our method and the fast equivalent-layer technique. 
-The total computational time required by our method to estimate the physical-property 
+is that the computational time required by our method to estimate the physical-property 
 distribution on the equivalent layer and to perform the upward/downward continuations 
@@
 Because  the Fourier approach requires only one DFT/IDFT of the data, whereas 
-our method requires one DFT/IDFT per iteration, it is computationally faster 
+our method requires one DFT/IDFT per iteration, the former is computationally faster 
 than our method.
 However, the considerably smaller noise amplification and practically nonexistent 
-border effect are the main advantages of our method over the classical Fourier 
+border effects obtained without applying any padding function to expand 
+the data are the main advantages of our method over the classical Fourier 
 approach, especially in the downward continuation.

### 2020-08-07 manuscript/conclusions.tex
include EIC corrections

@@
 for processing gravity data has a BTTB structure when the following conditions are
-satisfied: (i) the observed data are disposed at a regular grid on a horizontal plane
-and (ii) the equivalent sources (point masses) are placed at a constant depth, one directly 
+satisfied: (1) the observed data are disposed at a regular grid on a horizontal plane
+and (2) the equivalent sources (point masses) are placed at a constant depth, one directly 
 beneath each observation point.
@@
 of magnitude when processing one million of observations by using our approach. 
-Traditionally, such amount of data impractically requires $7.6$ Terabytes of RAM memory to handle 
-the full sensitivity matrix whereas our method takes only $61.035$ Megabytes.
-We have successfully applied the proposed method to compute the upward/downward continuations of 
+Traditionally, such amount of data impractically requires $7.6$ TB of memory to handle 
+the full sensitivity matrix whereas our method takes only $61.035$ MB.
+We successfully applied the proposed method to compute the upward/downward continuations of 
 synthetic gravity data. 
-Our results show that our method produces considerably smaller border effects and 
+The results show that our method produces considerably smaller border effects and 
 noise amplification compared to the classical Fourier approach, 
@@
 Application to field data over the Caraj{\'a}s Province, north of Brazil, 
-confirms the potential of our approach in processing an airborne gravity data set with $250,000$ observations in approximately $0.2$ seconds. 
+confirms the potential of our approach in processing an airborne gravity data set with $250,000$ observations in approximately $0.2$ s. 
 Our method can also be applied to process large marine gravity data sets. 
-Further studies need to be carried out in order to apply our method to process 
+Further studies need to be carried out to apply our method to process 
 other potential-field data.

### 2020-08-08 manuscript/figures.tex
fix figure 10
fix problems in references

@@
 \plot{Figure3}{width=10cm}{
-	{Comparison between the running time of the fast equivalent-layer technique \citep{siqueira-etal2017} and our method.
+	{Comparison between the runtime of the fast equivalent-layer technique \citep{siqueira-etal2017} and our method.
 	The values were obtained for $N^{it} = 50$ iterations.}
@@
 \plot{Figure4}{width=10cm}{
-	{Running time of our method for a number of observations $N$ up to $25$ millions. 
+	{Runtime of our method for a number of observations $N$ up to $25$ millions. 
 	The values were obtained for $N^{it} = 50$ iterations.}
@@
 	The computation times spent by the fast equivalent-layer technique and 
-	our method were $10.416$ and $0.177$ seconds, respectively.}
+	our method were $10.416$ s and $0.177$ s, respectively.}
 	\label{fig:comparison_datafits}
@@
 	The computation times spent by the fast equivalent-layer technique and 
-	our method were $8.697$ and $0.005$ seconds, respectively.}
+	our method were $8.697$ s and $0.005$ s, respectively.}
 	\label{fig:upward_comparison}
@@
 	The computation times spent by the fast equivalent-layer technique and 
-	our method were $8.795$ and $0.004$ seconds, respectively.}
+	our method were $8.795$ s and $0.004$ s, respectively.}
 	\label{fig:downward_comparison}
@@
 
-\plot{carajas_gz_predito}{width=8cm}{
+\plot{Carajas_gz_predito}{width=8cm}{
 	{Application to field data over the Caraj{\'a}s Province, Brazil. 
@@
 	observed data (Figure \ref{fig:carajas_real_data}).
-	The total computation time for processing of the $250,000$ observations was $0.216$ seconds.}
+	The total computation time for processing of the $250,000$ observations was $0.216$ s.}
 	\label{fig:up2000_carajas_500x500}

### 2020-08-09 manuscript/figures.tex
replot all figures with width 4.33 inches / 12000 dpi

@@
 
-\plot{synthetic_data_med}{width=13.5cm}{
+\plot{Figure5}{width=13.5cm}{
 	{Application to synthetic data. 
@@
 
-\plot{deltas_gz_med}{width=6cm}{
+\plot{Figure6}{width=6cm}{
 	{Application to synthetic data. 
@@
 
-\plot{upward_fourier_med}{width=13.5cm}{
+\plot{Figure7}{width=13.5cm}{
 	{Application to synthetic data.
@@
 
-\plot{downward_fourier_med}{width=13.5cm}{
+\plot{Figure8}{width=13.5cm}{
 	{Application to synthetic data. 
@@
 
-\plot{carajas_real_data}{width=13.5cm}{
+\plot{Figure9}{width=13.5cm}{
 	{Application to field data over the Caraj{\'a}s Province, Brazil. 
@@
 
-\plot{Carajas_gz_predito}{width=8cm}{
+\plot{Figure10}{width=8cm}{
 	{Application to field data over the Caraj{\'a}s Province, Brazil. 
@@
 
-\plot{up5000_carajas_500x500}{width=13.5cm}{
+\plot{Figure11}{width=13.5cm}{
 	{Application to field data over the Caraj{\'a}s Province, Brazil. 

### 2020-08-09 manuscript/figures.tex
fix figs 3 and 4, include aknowledgements, update makefile

@@
 
-\plot{Figure2}{width=\textwidth}{
+\plot{Figure2}{width=10cm}{
 	{Comparison between the number of flops (equations \ref{eq:float_fast_eqlayer} and \ref{eq:float_bccb}) 

### 2020-08-15 manuscript/abstract.tex

@@
 
-We have developed an efficient and very fast equivalent-layer technique for 
-gravity data processing by modifying an iterative method grounded on excess mass 
+We develop an efficient and very fast equivalent-layer technique for 
+gravity data processing by modifying an iterative method grounded on an excess mass 
 constraint that does not require the solution of linear systems. 
-Taking advantage of the symmetric Block-Toeplitz Toeplitz-block (BTTB) structure 
-of the sensitivity matrix, that raises when regular grids of observation points 
+Taking advantage of the symmetric Block-Toeplitz Toeplitz-Block (BTTB) structure 
+of the sensitivity matrix, that arises when regular grids of observation points 
 and equivalent sources (point masses) are used to set up a fictitious 
-equivalent layer, we have developed an algorithm which greatly reduces the number 
-of flops and RAM memory necessary to estimate a 2D mass distribution over the 
+equivalent layer, we develop an algorithm that greatly reduces the computational 
+complexity and RAM memory necessary to estimate a 2D mass distribution over the 
 equivalent layer. 
@@
 forward modeling in the fast equivalent-layer technique. 
-As a result, our method is efficient to process very large datasets. 
+As a result, our method is efficient for processing very large datasets. 
 Tests with synthetic data demonstrate the ability of our method to satisfactorily upward-
-and downward-continuing the gravity data.
+and downward-continuing gravity data.
 Our results show very small border effects and noise amplification compared to those 

### 2020-08-15 manuscript/introduction.tex

@@
 The equivalent layer is a well-known technique for processing potential-field data in applied geophysics 
-since the 60's. 
-It comes from potential theory as a mathematical solution of the Laplace's equation, in the region above the
+since the 1960's. 
+It comes from potential theory as a mathematical solution of Laplace's equation, in the region above the
 sources, by using the Dirichlet boundary condition \citep{kellogg1929}.
@@
 problems is still its high computational cost for processing large-data sets. 
-This problem propelled several studies to improve the computational efficiency of the 
+This problem propelled several studies aimed at improving the computational efficiency of the 
 equivalent layer technique. 
-\citet{leao-silva1989} developed a fast method for processing a regular grid of potential-field data.
+\citet{leao-silva1989} develop a fast method for processing a regular grid of potential-field data.
 The method consists in estimating an equivalent layer which exactly reproduces the potential-field data 
@@
 just one large linear system for the entire equivalent layer. 
-\citet{mendonca-silva1994} also followed the strategy of solving successive small linear systems for 
+\citet{mendonca-silva1994} also follow the strategy of solving successive small linear systems for 
 constructing an equivalent layer. Their method is based on the  equivalent-data concept, which 
@@
 interpolating surface that fits the chosen subset also automatically fits all remaining data. 
-The equivalent-data set is obtained by iteratively introducing the potential-field observation with the 
+The equivalent data set is obtained by iteratively introducing the potential-field observation with the 
 greatest residual in the preceding iteration. By applying to the interpolation problem, the method is 
@@
 
-By following a different strategy, \citet{li-oldenburg2010} developed a rapid method that transforms the 
+By following a different strategy, \citet{li-oldenburg2010} develop a rapid method that transforms the 
 dense sensitivity matrix associated with the linear system into a sparse one by using a wavelet technique. 
@@
 formulation using a dense matrix. 
-\citet{barnes-lumley2011} followed a similar strategy and transformed the dense sensitivity matrix 
+\citet{barnes-lumley2011} follow a similar strategy and transformed the dense sensitivity matrix 
 into a sparse one. However, differently from \citet{li-oldenburg2010}, their method operates in the space 
@@
 by solving the transformed linear system with a weighted-least-squares conjugate-gradient algorithm.
-Notice that, instead of constructing the equivalent layer by solving successive small linear systems, 
+Note that, instead of constructing the equivalent layer by solving successive small linear systems, 
 these last two methods first transform the large linear system into a sparse one and then take advantage of 
@@
 
-\citet{oliveirajr-etal2013} developed a fast method based on the reparameterization of the physical-property 
+\citet{oliveirajr-etal2013} develop a fast method based on the reparameterization of the physical-property 
 distribution within the equivalent layer. Those authors divided the equivalent layer into a regular grid of 
@@
 procedure that removes the maximum residual between the observed and predicted data.
-\citet{xia-sprowl1991} and \citet{xia-etal1993} developed 
+\citet{xia-sprowl1991} and \citet{xia-etal1993} develop 
 fast iterative schemes for updating the physical-property distribution
 within the equivalent layer in the wavenumber and space domains, respectively.
-Grounded on excess mass constraint, \cite{siqueira-etal2017} developed an iterative scheme 
+Grounded on excess mass constraint, \cite{siqueira-etal2017} develop an iterative scheme 
 starting with a mass distribution within the equivalent layer that is proportional to observed gravity data.
@@
 of point masses has a very well-defined structure called Block-Toeplitz Toeplitz-Block (BTTB) for 
-the case in which (i) the observed gravity data is located on a regularly spaced grid at constant 
-height and (ii) there is one point mass directly beneath each observation point.
+the case in which (1) the observed gravity data is located on a regularly spaced grid at constant 
+height and (2) there is one point mass directly beneath each observation point.
 This technique have been successfully used in potential-field methods for 

### 2020-08-15 manuscript/methodology.tex

@@
 transforms from $\mathrm{m/s^2}$ to mGal.
-Notice that the sensitivity matrix depends on the $i$th coordinate of the observation 
+Note that the sensitivity matrix depends on the $i$th coordinate of the observation 
 and the $j$th coordinate of the equivalent source. For convenience, we designate 
@@
 
-\citet{siqueira-etal2017} developed an iterative least-squares method to estimate the mass 
+\citet{siqueira-etal2017} develop an iterative least-squares method to estimate the mass 
 distribution over the equivalent layer based on the excess of mass and the positive correlation 
@@
 classical equivalent layer approach (equation \ref{eq:estimated-p-parameter-space}) if the 
-dataset is greater than at least 200 observation points, even using a large number of iterations.
+dataset is greater than 200 observation points, even using a large number of iterations.
 
@@
 the matrix-vector product 
-$\tensor{A} \hat{\mathbf{p}}^{k} = \mathbf{d}(\hat{\mathbf{p}}^{k})$ must be calculated to get a 
+$\tensor{A} \hat{\mathbf{p}}^{k} = \mathbf{d}(\hat{\mathbf{p}}^{k})$ must be calculated 
+to obtain a 
 new residual $\mathbf{d^0} - \tensor{A} \hat{\mathbf{p}}^{k}$, which represents a bottleneck. 
@@
 Consider that the observed data are located on an $N_{x} \times N_{y}$ regular grid of
-points regularly spaced by $\Delta x$ and $\Delta y$ along the $x$- and $y$-directions,
+points spaced by $\Delta x$ and $\Delta y$ along the $x$- and $y$-directions,
 respectively, on a horizontal plane defined by the constant vertical coordinate $z_{1} < z_{0}$. 
@@
 and $l(i)$ and $l(j)$ are defined by equation \ref{eq:l-y-oriented}.
-Notice that, for each element $a_{ij}$ (equation \ref{eq:aij-regular-grids}),
+Note that, for each element $a_{ij}$ (equation \ref{eq:aij-regular-grids}),
 defined by matrix indices $i$ and $j$, there is a corresponding block element
@@
 
-It is important noting that different matrix indices $i$ or $j$ produce the same 
+It is important to note that different matrix indices $i$ or $j$ produce the same 
 absolute values for the grid indices $k$ (equations \ref{eq:k-x-oriented} or 
@@
 technique \citep{siqueira-etal2017} accounts for most of its total computation time 
-and can cause RAM memory shortage when large data sets are used.
+and can cause RAM memory shortage when processing large data sets.
 This computational load can be drastically reduced by exploring the well-defined structure of 
@@
 is computed once and stored during the entire iterative process.
-On the other hand, our method computes only one column of $\mathbf{A}$ 
+On the other hand, our method calculates only one column of $\mathbf{A}$ 
 and uses it to compute the complex $2Q \times 2P$ matrix $\mathbf{L}$ 
@@
 $\mathbf{L}$. These quantities were computed for different numbers of observations $N$. 
-Notice that $N = 1,000,000$ observations require nearly $7.6$ Terabytes of RAM memory 
+Notice that $N = 1,000,000$ observations require nearly $7.6$ TB of memory 
 to store the whole matrix $\mathbf{A}$.
 
-Figure \ref{fig:time_fast_eqlayer_bccb} compares the running time of the fast equivalent-layer technique 
+Figure \ref{fig:time_fast_eqlayer_bccb} compares the runtime of the fast equivalent-layer technique 
 \citep{siqueira-etal2017} and of our method, considering a constant number of iterations $N^{it} = 50$. 
@@
 up to  $25$ millions. 
-These results shows that, while the running time of our method is $\approx 30.9$ seconds for 
-$N = 1,000,000$, the fast equivalent-layer technique spends $\approx 46.8$ seconds for $N = 22,500$.
+These results shows that, while the running time of our method is $\approx 30.9$ s for 
+$N = 1,000,000$, the fast equivalent-layer technique spends $\approx 46.8$ s for $N = 22,500$.
